---
title: "Renal Analysis Severe Malaria"
author: "James Watson"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = T, cache.comments = FALSE,
                      echo = F, include = TRUE, 
                      fig.width = 8, fig.height = 8,
                      fig.pos = 'H',dev = 'png', dpi = 300)

library(brms)
library(tidybayes)
library(tidyverse)
library(grid)
library(gridExtra)
library(rstanarm)
library(scales)
```



Plotting params

```{r}
source('setup_parameters.R')
RUN_STAN_MODELS=F
```


## Load analysis data

Analysis dataset
```{r}
load('RData/Analysis_dataset.RData')

sm_data_analysis = sm_data_analysis %>%
  mutate(Age_Categories3 = cut(Age,breaks=c(0,5,15,100)),
         Age_Categories3 = fct_recode(
           Age_Categories3,
           "<5"     = "(0,5]",
           "[5,15]" = "(5,15]",
           ">15"    = "(15,100]"
         ))

# Estimate baseline creatinine using the Schwartz formula
baseline_creatinine_schwartz <- function(height_cm) {
  egfr_normal = 120
  creat <- 0.413 * (height_cm / egfr_normal) * 88.4
  return(creat)
}

baseline_creatinine_pottel = function(age_years){
  egfr_normal = 120
  Q_age <- 0.243 * (age_years ^ 0.244)        # age-based reference value
  creat_mg_dl <- 107.3 * Q_age / egfr_normal  # baseline creatinine (mg/dL)
  creat_umol  <- creat_mg_dl * 88.4           # convert to µmol/L
  return(creat_umol)
}

baseline_creatinine_cockroftgault = function(age_years, weight_kg){
  creat_umol = 88.4*((140-age_years)*weight_kg)/(120*72)
  return(creat_umol)
}
baseline_creatinine_pottel(5)
baseline_creatinine_schwartz(100)
baseline_creatinine_cockroftgault(age_years = 40, weight_kg = 75)
```


# Calculate expected Cr based on age, weight

## Baseline creatinines

We look at the data from the TRACT study, Malawi site where we have follow-up creatinines and BUN in patients who survived and were not lost to follow-up. We compare these values to the expected values from the Pottel equation.

```{r baseline_creatinine_BUN, fig.width=12, fig.height=9}
baseline_data = sm_data %>% filter(!is.na(Baseline_Creatinine) | !is.na(`Baseline_Urea Nitrogen`)) %>%
  mutate(Baseline_Pottel = baseline_creatinine_pottel(Age),
         Baseline_Schwartz = baseline_creatinine_schwartz(Height_imputed),
         Outliers = Baseline_Creatinine > Baseline_Pottel+15) 

# p1= baseline_data %>%
#   ggplot(aes(x=Weight_kg, y = Baseline_Creatinine))+geom_point()+
#   theme_bw()+xlab('Weight (kg)')+ylab('Normal creatinine (umol/L)')+
#   geom_smooth(se = F)

p1= baseline_data %>%
  ggplot(aes(x=Age, y = Baseline_Creatinine))+geom_point(shape = 3)+
  stat_function(
    fun = baseline_creatinine_pottel,
    color = "red",
    linewidth = 1.2)+
  theme_bw(base_size = 16)+ xlab('Age (years)')+ylab('Normal creatinine (umol/L)')+
  geom_smooth(se = F)
p2=sm_data %>% filter(!is.na(Baseline_Creatinine)) %>%
  ggplot(aes(x=Height_cm, y = Baseline_Creatinine))+geom_point(shape = 3)+
  theme_bw(base_size = 16)+
  xlab('Height (cm)')+ylab('Normal creatinine (umol/L)')+geom_smooth(se=F)+
  stat_function(
    fun = baseline_creatinine_schwartz,
    color = "red",
    linewidth = 1.2
  )
# p4=sm_data %>% filter(!is.na(Baseline_Creatinine)) %>%
#   ggplot(aes(x=wfaz, y = Baseline_Creatinine))+geom_point()+
#   theme_bw()+
#   xlab('Weight-for-age z-score')+ylab('Normal creatinine (umol/L)')+
#   geom_smooth(se = F)


summary(lm(Baseline_Creatinine~Height_cm, data = sm_data))
summary(lm(Baseline_Creatinine~wfaz, data = sm_data))
summary(lm(Baseline_Creatinine~Age, data = sm_data))
summary(lm(Baseline_Creatinine~Weight_kg, data = sm_data))
eGFR=120 # Using Pottel equation

##########** BUN BASELINES **** ######
p3=sm_data %>% filter(!is.na(`Baseline_Urea Nitrogen`)) %>%
  ggplot(aes(x=Age, y = `Baseline_Urea Nitrogen`))+geom_point(shape = 3)+
  theme_bw(base_size = 16)+
  xlab('Age (years)')+ylab('Normal BUN (mmol/L)')+geom_smooth(se = F)


p4=sm_data %>% filter(!is.na(`Baseline_Urea Nitrogen`)) %>%
  ggplot(aes(x=Height_imputed, y = `Baseline_Urea Nitrogen`))+geom_point(shape = 3)+
  theme_bw(base_size = 18)+
  xlab('Height (cm)')+ylab('Normal BUN (mmol/L)')+geom_smooth(se = F)
mod_bun_baseline_age = mgcv::gam(b_bun ~ s(sqrt(Age),k=5), data=baseline_data %>% mutate(b_bun=`Baseline_Urea Nitrogen`))
round(predict(mod_bun_baseline_age, newdata = data.frame(Age=c(0.5,3,10))),1)


grid.arrange(p1,p2,p3,p4)

# 
# xx = sm_data %>% filter(!is.na(Baseline_Creatinine) | !is.na(`Baseline_Urea Nitrogen`))
# xx$MUAC = xx$`Mid-Upper Arm Circumference_cm`
# xx$Baseline_BUN = xx$`Baseline_Urea Nitrogen`
# mod_creat_baseline=stan_lm(Baseline_Creatinine ~ Age+Height_cm+Weight_kg,
#                            data = xx, prior = R2(0.5))
# plot(mod_creat_baseline,pars = c('Age','Height_cm','Weight_kg'))
# summary(mod_creat_baseline)
# 
# mod_bun_baseline=stan_lm(Baseline_BUN ~ scale(Age) + scale(Weight_kg)+scale(Height_cm)+SEX+scale(MUAC),
#                          data = xx, prior = R2(0.5))
# plot(mod_bun_baseline,pars = c('scale(Age)','scale(Weight_kg)','scale(Height_cm)','SEXM','scale(MUAC)'))
# summary(mod_bun_baseline)
baseline_data = baseline_data %>% filter(!Outliers) %>%
  mutate(Residual_Pottel = Baseline_Creatinine-Baseline_Pottel,
         Residual_Schwartz = Baseline_Creatinine-Baseline_Schwartz)
ylims = range(c(baseline_data$Residual_Pottel, baseline_data$Residual_Schwartz))+c(-1,1)

p1= baseline_data %>% 
  ggplot(aes(x=Baseline_Creatinine, y = Residual_Pottel))+geom_point()+
  scale_x_log10()+
  theme_bw(base_size = 18)+ xlab('Baseline creatinine (umol/L)')+ylab('Residual: Baseline vs Pottel predicted')+
  geom_smooth(se = F, method = lm)+ylim(ylims[1], ylims[2])+geom_hline(yintercept = c(-10,10))

p2= baseline_data %>% 
  ggplot(aes(x=Baseline_Creatinine, y = Residual_Schwartz))+geom_point()+
  scale_x_log10()+
  theme_bw(base_size = 18)+ xlab('Baseline creatinine (umol/L)')+ylab('Residual: Baseline vs Schwartz predicted')+
  geom_smooth(se = F, method = lm)+ylim(ylims[1], ylims[2])+geom_hline(yintercept = c(-10,10))
grid.arrange(p1,p2,nrow=1)
```

Pottel age-based equation, where eGFR=107.3/(SCr/Q)
assuming a normal GFR of 120 mL/min per 1.73 m2 (height-independent). 
Q=0.0270*age+0.2329.

We estimate the Pottel baseline creatinine (age only for children)
We estimate the adult baseline creatine using Cockcroft-Gault equation
We also estimate across the age spectrum using the Full Age equation (Pottel 2016)


```{r}
sm_data_analysis = sm_data_analysis %>%
  mutate(
    # Height_m = Height_imputed/100,
    # Height_m = ifelse(Height_m<0.8, 0.8, Height_m),
    # Q2=3.94-13.4*Height_m+17.6*Height_m^2-9.84*Height_m^3+2.04*Height_m^4,
    Baseline_Pottel = baseline_creatinine_pottel(age_years = Age),
    Baseline_CG = baseline_creatinine_cockroftgault(age_years = Age, weight_kg = Weight_kg),
    Baseline_Cr_Pottel_CG = ifelse(Age<15, Baseline_Pottel, Baseline_CG),
    # Baseline_Cr_Height = ifelse(Age<15, (107.3/120)*Q2*88.4, bSrc_adults),
    Baseline_Cr_Schwartz = baseline_creatinine_schwartz(Height_imputed))

sm_data_analysis %>% ggplot(aes(x=Age, y = Baseline_Pottel))+geom_point()+
  geom_point(aes(x=Age, y=Baseline_Cr_Schwartz, colour = 'red',alpha=.5))+
  theme_bw()

# # Make the Fold Change values for each patients
sm_data_analysis = sm_data_analysis %>%
  mutate(Creatinine_fold_change_Pottel_CG = log2(Creatinine_umol_L/Baseline_Cr_Pottel_CG),
         Creatinine_fold_change_Schwartz = log2(Creatinine_umol_L/Baseline_Cr_Schwartz))

p1=sm_data_analysis %>% ggplot(aes(x=Age, y=Baseline_Cr_Pottel_CG))+geom_point()+
  theme_bw()+ylim(0,120)+ylab('Pottel/CG predicted baseline Creatinine (umol/L)')
p2=sm_data_analysis %>% ggplot(aes(x=Age, y=Baseline_Cr_Schwartz))+geom_point()+
  theme_bw()+ylim(0,120)+ylab('Schwartz predicted baseline Creatinine (umol/L)')

grid.arrange(p1,p2)
```


TRACT predicted baseline creatinine for children. Age predicted values look better than the height predicted values.

```{r tract_baseline_Cr}
p1=sm_data_analysis %>% filter(!is.na(Baseline_Creatinine)) %>%
  ggplot(aes(x=Baseline_Creatinine, y = Baseline_Cr_Pottel_CG))+geom_point()+
  xlab('Observed baseline Creatinine in TRACT')+
  ylab('Age (Pottel equation) predicted baseline Creatinine (umol/L)')+
  theme_bw()+geom_abline(intercept = 0, slope = 1)

p2=sm_data_analysis %>% filter(!is.na(Baseline_Creatinine)) %>%
  ggplot(aes(x= Baseline_Creatinine, y= Baseline_Cr_Schwartz))+geom_point()+
  theme_bw()+geom_abline(intercept = 0, slope = 1)+
  ylab('Height (Schwartz equation) predicted baseline Creatinine (umol/L)')+
  xlab('Observed baseline Creatinine in TRACT')
grid.arrange(p1,p2)
```


is there heterogeneity by age/weight?

```{r creatinine_fold_change}
sm_data_creat = sm_data_analysis %>% filter(!is.na(Creatinine_umol_L))
p0 = sm_data_creat %>% ggplot(aes(x=Creatinine_fold_change_Pottel_CG, y = Creatinine_fold_change_Schwartz, colour = Age_category))+
  geom_point()+theme_bw()+xlab('Creatinine fold change (Potel or CG)')+
  ylab('Creatinine fold change (Schwartz)')+
  geom_abline(slope = 1, intercept = 0)

p1=sm_data_creat %>% ggplot(aes(x=Age, y = Creatinine_fold_change_Pottel_CG, colour = STUDY))+
  geom_point()+theme_bw(base_size = 14)+
  geom_smooth(aes(group=NA))+ylab('Creatinine fold change (Potel or CG)')
p2=sm_data_creat %>% ggplot(aes(x=Age, y = Creatinine_fold_change_Schwartz, colour = STUDY))+
  geom_point()+theme_bw(base_size = 14)+
  geom_smooth(aes(group=NA))+ylab('Creatinine fold change (Schwartz)')

grid.arrange(p0,p1,p2)
```


```{r age_vs_Cr_foldchange, fig.width=10, fig.height=7}
sm_data_creat %>% 
  ggplot(aes(x=Age, y = 2^Creatinine_fold_change_Pottel_CG))+
  geom_point()+ scale_y_log10()+scale_x_sqrt()+
  geom_smooth()+theme_bw(base_size = 12)+
  ylab('Creatinine fold change (Pottel/CG)')+
  facet_wrap(~STUDY,nrow=2)
```



When using the creatinine fold change as a predictor of death there is still some residual heterogeneity. This seems to be entirely driven by the TRACT study data.

```{r mgcv_tensor_product_creatinine}
par(las=1)
# Fit GAM with age-varying spline for fold-change creatinine
m = mgcv::gam(
  Died_28D ~ s(Age, k = 6) + 
    te(Creatinine_fold_change_Pottel_CG, Age, bs = c("cr", "cr"), k = c(10, 6))+
    Coma + s(SITE_STUDY,bs='re'),
  data   = sm_data_creat,
  family = binomial(link = "logit"),
  method = "REML",
  select = TRUE
)
mgcv::vis.gam(m, view = c("Creatinine_fold_change_Pottel_CG", "Age"),
              plot.type = "contour", color = "topo", too.far = 0.05,
              main='Predicted baseline using Age/Weight')

# Fit GAM with age-varying spline for fold-change creatinine - without TRACT
m_rmTRACT = mgcv::gam(
  Died_28D ~ s(Age, k = 6) + 
    te(Creatinine_fold_change_Pottel_CG, Age, bs = c("cr", "cr"), k = c(10, 6))+
    Coma + s(SITE_STUDY,bs='re'),
  data   = sm_data_creat %>% filter(STUDY != 'TRACT'),
  family = binomial(link = "logit"),
  method = "REML",
  select = TRUE
)
mgcv::vis.gam(m_rmTRACT, view = c("Creatinine_fold_change_Pottel_CG", "Age"),
              plot.type = "contour", color = "topo", too.far = 0.05,
              main='Predicted baseline using Age/Weight\nwithout TRACT')
```


```{r}
#tensor product models
Creat_tensor_prod_Pottel_CG = as.formula(bf(Died_28D ~ t2(Creatinine_fold_change_Pottel_CG,Age,bs = c("cr", "cr"))+Coma+s(year_admission)+(1|STUDY/SITE_STUDY)))
Creat_tensor_prod_Schwartz = as.formula(bf(Died_28D ~ t2(Creatinine_fold_change_Schwartz,Age,bs = c("cr", "cr"))+Coma+s(year_admission)+(1|STUDY/SITE_STUDY)))
Creat_tensor_prod_abs = as.formula(bf(Died_28D ~ t2(log10(Creatinine_umol_L),Age,bs = c("cr", "cr"))+Coma+s(year_admission)+(1|STUDY/SITE_STUDY)))

#smooth spline models with categorical age
Creat_fs_mod_Pottel_CG = as.formula(bf(Died_28D ~ s(Creatinine_fold_change_Pottel_CG, bs = "cr", k = 10) + s(Creatinine_fold_change_Pottel_CG,Age_category,bs = "fs", m = 1, k = 10)+Age_category+Coma+s(year_admission)+(1|STUDY/SITE_STUDY)))
Creat_fs_mod_Schwartz = as.formula(bf(Died_28D ~ s(Creatinine_fold_change_Schwartz, bs = "cr", k = 10) + s(Creatinine_fold_change_Schwartz,Age_category,bs = "fs", m = 1, k = 10)+Age_category+Coma+s(year_admission)+(1|STUDY/SITE_STUDY)))

f_path = 'RData/creatinine_models.RData'
if(!file.exists(f_path) | RUN_STAN_MODELS){
  niter = 2000
  Creat_tensor_prod_fit_Pottel_CG = 
    brm(formula = Creat_tensor_prod_Pottel_CG,
        data = sm_data_creat ,
        family = bernoulli(), cores = 4, seed = 17,
        iter = niter, warmup = niter/2, thin = 1, refresh = 250)
  
  Creat_tensor_prod_fit_Schwartz = 
    brm(formula = Creat_tensor_prod_Schwartz,
        data = sm_data_creat ,
        family = bernoulli(), cores = 4, seed = 17,
        iter = niter, warmup = niter/2, thin = 1, refresh = 250)  
  Creat_tensor_prod_fit_abs = 
    brm(formula = Creat_tensor_prod_abs,
        data = sm_data_creat ,
        family = bernoulli(), cores = 4, seed = 17,
        iter = niter, warmup = niter/2, thin = 1, refresh = 250)  
  Creat_fs_mod_fit_Pottel_CG = 
    brm(formula = Creat_fs_mod_Pottel_CG,
        data = sm_data_creat ,
        family = bernoulli(), cores = 4, seed = 17,
        iter = niter, warmup = niter/2, thin = 1, refresh = 250)  
  Creat_fs_mod_fit_Schwartz = 
    brm(formula = Creat_fs_mod_Schwartz,
        data = sm_data_creat ,
        family = bernoulli(), cores = 4, seed = 17,
        iter = niter, warmup = niter/2, thin = 1, refresh = 250)  
  
  save(Creat_tensor_prod_fit_Pottel_CG, Creat_tensor_prod_fit_Schwartz, Creat_tensor_prod_fit_abs, Creat_fs_mod_fit_Pottel_CG, Creat_fs_mod_fit_Schwartz, file = f_path)
} else {
  load(f_path)
}
```


summary of model fits

```{r}
summary(Creat_tensor_prod_fit_Pottel_CG)
summary(Creat_tensor_prod_fit_Schwartz)
summary(Creat_tensor_prod_fit_abs)
summary(Creat_fs_mod_fit_Pottel_CG)
summary(Creat_fs_mod_fit_Schwartz)
```


```{r nested_random_effects}
ranef_Pottel_CG = as.data.frame(ranef(Creat_tensor_prod_fit_Pottel_CG)$STUDY[,,'Intercept'])
ranef_Pottel_CG$Study = rownames(ranef_Pottel_CG)
ranef_Pottel_CG %>% ggplot(aes(x =Study, y = Estimate ))+geom_point()+geom_errorbar(aes(ymin = Q2.5, ymax = Q97.5))+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+
  ylab('Log odds ratio for death')


ranef_Pottel_CG = as.data.frame(ranef(Creat_tensor_prod_fit_Pottel_CG)$`STUDY:SITE_STUDY`[,,'Intercept'])
ranef_Pottel_CG$Site = rownames(ranef_Pottel_CG)
ranef_Pottel_CG %>% ggplot(aes(x =Site, y = Estimate ))+geom_point()+geom_errorbar(aes(ymin = Q2.5, ymax = Q97.5))+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+
  ylab('Log odds ratio for death')

```


Make some test data

```{r}
k_vals=40
test_data = expand_grid(Creatinine_fold_change_Pottel_CG = seq(0,3,length.out=k_vals),
                        Age = c(2, 5, 20),
                        year_admission=2020,
                        Coma=unique(sm_data_creat$Coma),
                        STUDY=unique(sm_data_creat$STUDY), 
                        SITE_STUDY = -1) %>% arrange(Creatinine_fold_change_Pottel_CG) %>%
  mutate(Creatinine_fold_change_Schwartz=Creatinine_fold_change_Pottel_CG,
         Age_category=cut(Age,breaks = c(0, 2, 5, 15, 100)),
         Age_category = recode(Age_category,
                               "(0,2]" = "≤2",
                               "(2,5]" = "(2,5]",
                               "(5,15]" = "(5,15]",
                               "(15,100]" = ">15"))

test_data2 = expand_grid(Creatinine_umol_L = 10^seq(1,3,length.out=k_vals),
                         Age = c(2, 5, 20),
                         year_admission=2020,
                         Coma=unique(sm_data_creat$Coma),
                         STUDY=unique(sm_data_creat$STUDY), 
                         SITE_STUDY = -1) %>% arrange(Creatinine_umol_L) %>%
  mutate(Age_category=cut(Age,breaks = c(0, 2, 5, 15, 100)),
         Age_category = recode(Age_category,
                               "(0,2]" = "≤2",
                               "(2,5]" = "(2,5]",
                               "(5,15]" = "(5,15]",
                               "(15,100]" = ">15"))

# remove non plausible combinations
# test_data = test_data %>%
#   filter(!(Age_category==">15" & Weight_category=='(0,25]'),
#          !(Age_category %in% c("(2,5]","≤2") & Weight_category=='(25,100]')) %>%
#   mutate(Age_category=factor(Age_category, levels = c("≤2","(2,5]","(5,15]",">15")))

# get model predictions
# Full model
preds_Pottel_CG = posterior_epred(Creat_tensor_prod_fit_Pottel_CG, newdata = test_data, re_formula = NA)
preds_Schwartz = posterior_epred(Creat_tensor_prod_fit_Schwartz, newdata = test_data, re_formula = NA)
test_data$Creat_tensor_prod_pred_Pottel_CG = colMeans(preds_Pottel_CG)
test_data$Creat_tensor_prod_pred_Schwartz = colMeans(preds_Schwartz)
test_data$Creat_tensor_prod_threshold_Pottel_CG = apply(preds_Pottel_CG, 2, f_threshold,mortality_threshold=mortality_threshold)
test_data$Creat_tensor_prod_threshold_Schwartz = apply(preds_Schwartz, 2, f_threshold,mortality_threshold=mortality_threshold)

# Full model - abs cr values
preds = posterior_epred(Creat_tensor_prod_fit_abs, newdata = test_data2, re_formula = NA)
test_data2$Creat_abs_tensor_prod_pred = colMeans(preds)
test_data2$Creat_abs_tensor_prod_threshold = apply(preds, 2, f_threshold,mortality_threshold=mortality_threshold)

preds_Pottel_CG = posterior_epred(Creat_fs_mod_fit_Pottel_CG, newdata = test_data, re_formula = NA)
preds_Schwartz = posterior_epred(Creat_fs_mod_fit_Schwartz, newdata = test_data, re_formula = NA)
test_data$Creat_fs_pred_Pottel_CG = colMeans(preds_Pottel_CG)
test_data$Creat_fs_pred_Schwartz = colMeans(preds_Schwartz)
test_data$Creat_fs_threshold_Pottel_CG = apply(preds_Pottel_CG, 2, f_threshold,mortality_threshold=mortality_threshold)
test_data$Creat_fs_threshold_Schwartz = apply(preds_Schwartz, 2, f_threshold,mortality_threshold=mortality_threshold)
```


# The effect of year
```{r temporal_trend}
test_data_year = expand_grid(Creatinine_fold_change_Pottel_CG = c(1,3),
                             Age = 2,
                             year_admission=1990:2020,
                             Coma='No',
                             STUDY=unique(sm_data_creat$STUDY), 
                             SITE_STUDY = -1) %>% arrange(Creatinine_fold_change_Pottel_CG) %>%
  mutate(Creatinine_fold_change_Schwartz=Creatinine_fold_change_Pottel_CG,
         Age_category=cut(Age,breaks = c(0, 2, 5, 15, 100)),
         Age_category = recode(Age_category,
                               "(0,2]" = "≤2",
                               "(2,5]" = "(2,5]",
                               "(5,15]" = "(5,15]",
                               "(15,100]" = ">15"))


preds_Pottel_CG = posterior_epred(Creat_tensor_prod_fit_Pottel_CG, newdata = test_data_year, re_formula = NA)
test_data_year$pred_year = colMeans(preds_Pottel_CG)
test_data_year %>% ggplot(aes(x=year_admission, y = pred_year, 
                              group = Creatinine_fold_change_Pottel_CG))+
  geom_line(linewidth = 1.5)+ ylab('Predicted mortality')+
  theme_bw(base_size = 15)+
  xlab('Year')+
  geom_hline(yintercept = mortality_threshold)+
  scale_colour_manual(
    name = "Age (years)",
    values = c("2" = "orange", "5" = "darkred", "20" = 'purple') )+
  theme(legend.position = c(0.02, 0.98),  # (x, y) in [0,1] relative to plot area
        legend.justification = c(0, 1))   # align legend's bottom-left to that point
```



Using absolute creatinine

```{r creatinine_abs_tensor}
p0c=test_data2 %>% mutate(Age = as.factor(Age))%>%
  ggplot(aes(x = Creatinine_umol_L, y = Creat_abs_tensor_prod_pred, 
             colour =Age,linetype = Coma,
             group = interaction(Age,Coma)))+
  geom_line(linewidth = 1.5)+ ylab('Predicted mortality')+
  scale_x_log10() +
  annotation_logticks(sides='b')+theme_bw(base_size = 15)+
  xlab('Creatinine (umol/L)')+
  geom_hline(yintercept = mortality_threshold)+
  scale_colour_manual(
    name = "Age (years)",
    values = c("2" = "orange", "5" = "darkred", "20" = 'purple') )+
  theme(legend.position = c(0.02, 0.98),  # (x, y) in [0,1] relative to plot area
        legend.justification = c(0, 1))   # align legend's bottom-left to that point


p0c_threshold=test_data2 %>%mutate(Age = as.factor(Age))%>%
  ggplot(aes(x = Creatinine_umol_L, y = Creat_abs_tensor_prod_threshold, 
             colour = Age,linetype = Coma,
             group = interaction(Age,Coma)))+
  geom_line(linewidth = 1.5)+ylab('Probability mortality > 5%')+
  scale_x_log10() + theme_bw(base_size = 15)+
  annotation_logticks(sides='b')+
  scale_colour_manual(
    name = "Age (years)",
    values = c("2" = "orange", "5" = "darkred", "20" = 'purple') )+
  xlab('Creatinine (umol/L)')+
  geom_hline(yintercept = .9)+
  theme(legend.position = 'none')
grid.arrange(p0c, p0c_threshold)
```


Using creatinine fold change based on age and weight (adults) - Pottel equation and CG equation

```{r creatinine_fc_baseline_mortality_Pottel}
p1c=test_data %>% mutate(Age = as.factor(Age))%>%
  ggplot(aes(x = 2^Creatinine_fold_change_Pottel_CG, y = Creat_tensor_prod_pred_Pottel_CG, 
             colour =Age,linetype = Coma,
             group = interaction(Age,Coma)))+
  geom_line(linewidth = 1.5)+ ylab('Predicted mortality')+
  scale_x_log10() +
  annotation_logticks(sides='b')+theme_bw(base_size = 15)+
  xlab('Creatinine fold change')+
  geom_hline(yintercept = mortality_threshold)+
  scale_colour_manual(
    name = "Age (years)",
    values = c("2" = "orange", "5" = "darkred", "20" = 'purple') )+
  theme(legend.position = c(0.02, 0.98),  # (x, y) in [0,1] relative to plot area
        legend.justification = c(0, 1))   # align legend's bottom-left to that point


p1c_threshold=test_data %>%mutate(Age = as.factor(Age))%>%
  ggplot(aes(x = 2^Creatinine_fold_change_Pottel_CG, y = Creat_tensor_prod_threshold_Pottel_CG, 
             colour = Age,linetype = Coma,
             group = interaction(Age,Coma)))+
  geom_line(linewidth = 1.5)+ylab('Probability mortality > 5%')+
  scale_x_log10() + theme_bw(base_size = 15)+
  annotation_logticks(sides='b')+
  scale_colour_manual(
    name = "Age (years)",
    values = c("2" = "orange", "5" = "darkred", "20" = 'purple') )+
  xlab('Creatinine fold change')+
  geom_hline(yintercept = .9)+geom_vline(xintercept = 3)+
  theme(legend.position = 'none')



p2c=test_data %>% mutate(Age = as.factor(Age))%>%
  ggplot(aes(x = 2^Creatinine_fold_change_Pottel_CG, y = Creat_fs_pred_Pottel_CG, 
             colour =Age,linetype = Coma,
             group = interaction(Age,Coma)))+
  geom_line(linewidth = 1.5)+ ylab('Predicted mortality')+
  scale_x_log10() +
  annotation_logticks(sides='b')+theme_bw(base_size = 15)+
  xlab('Creatinine fold change')+
  geom_hline(yintercept = mortality_threshold)+
  labs(colour='Age')+
  theme(legend.position = c(0.02, 0.98),  # (x, y) in [0,1] relative to plot area
        legend.justification = c(0, 1))   # align legend's bottom-left to that point


p2c_threshold=test_data %>%mutate(Age = as.factor(Age))%>%
  ggplot(aes(x = 2^Creatinine_fold_change_Pottel_CG, y = Creat_fs_threshold_Pottel_CG, 
             colour = Age,linetype = Coma,
             group = interaction(Age,Coma)))+
  geom_line(linewidth = 1.5)+ylab('Probability mortality > 5%')+
  scale_x_log10() + theme_bw(base_size = 15)+
  annotation_logticks(sides='b')+
  xlab('Creatinine fold change')+
  geom_hline(yintercept = .9)+geom_vline(xintercept = 3)+
  theme(legend.position = 'none')

grid.arrange(p1c, p1c_threshold, p2c, p2c_threshold)


sm_data_analysis %>% filter(!is.na(Creatinine_umol_L)) %>%
  ggplot(aes(x=Age, y = Creatinine_umol_L, colour = Creatinine_fold_change_Pottel_CG>log2(3)))+
  geom_point()+scale_y_log10(breaks = breaks, minor_breaks = minor_breaks)+
  geom_hline(yintercept = 265,aes(linetype = 'dashed'))+
  scale_x_sqrt()+theme_bw(base_size = 14)+
  geom_vline(xintercept = c(2, 5,15))
```


Combined figure with BUN for paper

```{r combined_fig, fig.width=12, fig.height=10}
load('RData/bun_figures.RData')
# combined fig 3
grid.arrange(p0c+geom_vline(xintercept = c(265, 150)), p1c+theme(legend.position = 'none'), 
             p1+theme(legend.position = 'none'),p0c_threshold, 
             p1c_threshold, p1_threshold, nrow=2)
```



Using creatinine fold change based on Height

```{r creatinine_fc_baseline_mortality_Schwartz}
p1c=test_data %>% mutate(Age = as.factor(Age))%>%
  ggplot(aes(x = 2^Creatinine_fold_change_Schwartz, y = Creat_tensor_prod_pred_Schwartz, 
             colour =Age,linetype = Coma,
             group = interaction(Age,Coma)))+
  geom_line(linewidth = 1.5)+ ylab('Predicted mortality')+
  scale_x_log10() +
  annotation_logticks(sides='b')+theme_bw(base_size = 15)+
  xlab('Creatinine fold change (Schwartz)')+
  geom_hline(yintercept = mortality_threshold)+
  scale_colour_manual(
    name = "Age (years)",
    values = c("2" = "orange", "5" = "darkred", "20" = 'purple') )+
  theme(legend.position = c(0.02, 0.98),  # (x, y) in [0,1] relative to plot area
        legend.justification = c(0, 1))   # align legend's bottom-left to that point


p1c_threshold=test_data %>%mutate(Age = as.factor(Age))%>%
  ggplot(aes(x = 2^Creatinine_fold_change_Schwartz, y = Creat_tensor_prod_threshold_Schwartz, 
             colour = Age,linetype = Coma,
             group = interaction(Age,Coma)))+
  geom_line(linewidth = 1.5)+ylab('Probability mortality > 5%')+
  scale_x_log10() + theme_bw(base_size = 15)+
  annotation_logticks(sides='b')+
  scale_colour_manual(
    name = "Age (years)",
    values = c("2" = "orange", "5" = "darkred", "20" = 'purple') )+
  xlab('Creatinine fold change (Schwartz)')+
  geom_hline(yintercept = .9)+geom_vline(xintercept = 3)+
  theme(legend.position = 'none')



p2c=test_data %>% mutate(Age = as.factor(Age))%>%
  ggplot(aes(x = 2^Creatinine_fold_change_Schwartz, y = Creat_fs_pred_Schwartz, 
             colour =Age,linetype = Coma,
             group = interaction(Age,Coma)))+
  geom_line(linewidth = 1.5)+ ylab('Predicted mortality')+
  scale_x_log10() +
  annotation_logticks(sides='b')+theme_bw(base_size = 15)+
  xlab('Creatinine fold change (Schwartz)')+
  geom_hline(yintercept = mortality_threshold)+
  labs(colour='Age')+
  theme(legend.position = c(0.02, 0.98),  # (x, y) in [0,1] relative to plot area
        legend.justification = c(0, 1))   # align legend's bottom-left to that point


p2c_threshold=test_data %>%mutate(Age = as.factor(Age))%>%
  ggplot(aes(x = 2^Creatinine_fold_change_Schwartz, y = Creat_fs_threshold_Schwartz, 
             colour = Age,linetype = Coma,
             group = interaction(Age,Coma)))+
  geom_line(linewidth = 1.5)+ylab('Probability mortality > 5%')+
  scale_x_log10() + theme_bw(base_size = 15)+
  annotation_logticks(sides='b')+
  xlab('Creatinine fold change (Schwartz)')+
  geom_hline(yintercept = .9)+geom_vline(xintercept = 3)+
  theme(legend.position = 'none')

grid.arrange(p1c, p1c_threshold, p2c, p2c_threshold)


sm_data_analysis %>% filter(!is.na(Creatinine_umol_L)) %>%
  ggplot(aes(x=Age, y = Creatinine_umol_L, colour = Creatinine_fold_change_Schwartz>log2(3)))+
  geom_point()+scale_y_log10(breaks = breaks, minor_breaks = minor_breaks)+
  geom_hline(yintercept = 265,aes(linetype = 'dashed'))+
  scale_x_sqrt()+theme_bw(base_size = 14)+
  geom_vline(xintercept = c(2, 5,15))


sm_data_analysis %>% filter(!is.na(Creatinine_umol_L)) %>%
  ggplot(aes(x=Age, y = Creatinine_fold_change_Pottel_CG- Creatinine_fold_change_Schwartz))+
  geom_point()+
  scale_x_sqrt()+theme_bw(base_size = 14)+
  geom_vline(xintercept = c(2, 5,15))
```



Model using age as categorical variable

```{r creatinine_fs_mod}
p1=test_data %>%
  ggplot(aes(x = 2^Creatinine_fold_change_Pottel_CG, y = Creat_fs_pred_Pottel_CG, 
             colour = Age_category,linetype = Coma,
             group = interaction(Age_category,Coma)))+
  geom_line()+ ylab('Predicted mortality')+
  scale_x_log10() +
  annotation_logticks(sides='b')+theme_bw(base_size = 15)+
  xlab('Creatinine fold change')+
  geom_hline(yintercept = mortality_threshold)+
  labs(colour='Age category')

p1_threshold=test_data %>%
  ggplot(aes(x = 2^Creatinine_fold_change_Pottel_CG, y = Creat_fs_threshold_Pottel_CG, 
             colour = Age_category,linetype = Coma,
             group = interaction(Age_category,Coma)))+
  geom_line()+ylab('Probability mortality > 5%')+
  scale_x_log10() +
  geom_hline(yintercept = .9)+geom_vline(xintercept = 3)+
  annotation_logticks(sides='b')+theme_bw(base_size = 15)+
  xlab('Creatinine fold change')
grid.arrange(p1, p1_threshold)
```



### Compare creatinine fold change with BUN

Compare BUN versus Creatinine when both measured

```{r}
sm_both = sm_data_analysis %>% filter(!is.na(Creatinine_umol_L), !is.na(BUN))

# children under 5
mod_both_U5 = lme4::glmer(Died_28D~Age+scale(log10(BUN))+scale(Creatinine_fold_change_Pottel_CG)+Coma+(1|SITE_STUDY),
                          family='binomial', data = sm_both %>% filter(Age<=5))
summary(mod_both_U5)

# children 5-15
mod_both_5_15 = lme4::glmer(Died_28D~Age+scale(log10(BUN))+scale(Creatinine_fold_change_Pottel_CG)+Coma+(1|SITE_STUDY),
                            family='binomial', data = sm_both %>% filter(Age>5, Age<15))
summary(mod_both_5_15)

# Adults
mod_both_adults = lme4::glmer(Died_28D~Age+scale(log10(BUN))+scale(Creatinine_fold_change_Pottel_CG)+Coma+(1|SITE_STUDY),
                              family='binomial', data = sm_both %>% filter(Age>15))
summary(mod_both_adults)

### Fit interaction model
mod_both_all = lme4::glmer(Died_28D~Age_category*scale(log10(BUN))+Age_category*scale(Creatinine_fold_change_Pottel_CG)+Coma+(1|SITE_STUDY),
                           family='binomial', data = sm_both)
summary(mod_both_all)

sm_both = sm_both %>% mutate(BUN_log2 = log2(BUN))

f_brm_mod_both = as.formula(bf(Died_28D ~ BUN_log2*Age_category +
                                 Creatinine_fold_change_Pottel_CG*Age_category+
                                 Coma+(1|STUDY/SITE_STUDY)))

f_brm_mod_both_scaled = as.formula(bf(Died_28D ~ scale(BUN_log2)*Age_category +
                                        scale(Creatinine_fold_change_Pottel_CG)*Age_category+
                                        Coma+(1|STUDY/SITE_STUDY)))
f_path = 'RData/Model_both.RData'
if(RUN_STAN_MODELS | !file.exists(f_path)){
  mod_both_stan = brm(formula = f_brm_mod_both,
                      data = sm_both,
                      family = bernoulli(), cores = 4, seed = 17,
                      iter = 2000, warmup = 1000, thin = 1, refresh = 250)
  
  mod_both_stan_scaled = brm(formula = f_brm_mod_both_scaled,
                             data = sm_both,
                             family = bernoulli(), cores = 4, seed = 17,
                             iter = 2000, warmup = 1000, thin = 1, refresh = 250)
  
  save(mod_both_stan,mod_both_stan_scaled, file = f_path)
} else{
  load(f_path)
}
brms::mcmc_plot(mod_both_stan)
```


```{r marginal_effects_children_adults, fig.width=10, fig.height=6}
posterior <- brms::as_draws_df(mod_both_stan)
marginal_effects <- posterior %>%
  mutate(
    BUN_adults = b_BUN_log2+`b_BUN_log2:Age_category>15`,
    BUN_U2 = b_BUN_log2,
    BUN_25 = b_BUN_log2+`b_BUN_log2:Age_category25`,
    BUN_515 = b_BUN_log2+`b_BUN_log2:Age_category515`,
    Creat_adults = b_Creatinine_fold_change_Pottel_CG+`b_Age_category>15:Creatinine_fold_change_Pottel_CG`,
    Creat_U2 = b_Creatinine_fold_change_Pottel_CG,
    Creat_25 = b_Creatinine_fold_change_Pottel_CG+`b_Age_category25:Creatinine_fold_change_Pottel_CG`,
    Creat_515 = b_Creatinine_fold_change_Pottel_CG+`b_Age_category515:Creatinine_fold_change_Pottel_CG`
  ) %>%
  select(BUN_U2, BUN_25, BUN_515,BUN_adults, 
         Creat_U2, Creat_25, Creat_515, Creat_adults) %>%
  pivot_longer(cols = everything(),
               names_to = c("Biomarker", "Age_category"),
               names_pattern = "(.*)_(.*)",
               values_to = "estimate")

summary_df <- marginal_effects %>%
  group_by(Biomarker, Age_category) %>%
  summarise(
    OR = median(exp(estimate)),
    lower = quantile(exp(estimate), 0.025),
    upper = quantile(exp(estimate), 0.975),
    .groups = "drop"
  ) %>%
  mutate(
    Age_category2 = 
      case_when(
        Age_category=='25' ~ '[2-5]',
        Age_category=='515'~ '(5-15]',
        Age_category=='U2' ~ '<2',
        Age_category=='adults' ~ '>15'), 
    Biomarker = case_when(
      Biomarker=='BUN' ~ 'BUN\n(absolute value)',
      Biomarker=='Creat' ~ 'Creatinine\n(Fold change from baseline)'),
    Age_category = factor(Age_category2, levels = c('<2','[2-5]','(5-15]','>15'))
  ) %>% arrange(Biomarker, Age_category)

ggplot(summary_df, aes(x = Age_category2, y = OR, ymin = lower, ymax = upper, color = Biomarker)) +
  geom_pointrange(position = position_dodge(width = 0.5)) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_y_log10()+
  coord_flip() +
  labs(
    y = "Odds ratio for death per doubling",
    x='',
    title = "Multivariable model of death",
    color = "Biomarker"
  ) +
  theme_bw(base_size = 16)
```


scaled predictors:

```{r marginal_effects_children_adults_scaled, fig.width=10, fig.height=6}
posterior <- brms::as_draws_df(mod_both_stan_scaled)
marginal_effects <- posterior %>%
  mutate(
    BUN_adults = b_scaleBUN_log2+`b_scaleBUN_log2:Age_category>15`,
    BUN_U2 = b_scaleBUN_log2,
    BUN_25 = b_scaleBUN_log2+`b_scaleBUN_log2:Age_category25`,
    BUN_515 = b_scaleBUN_log2+`b_scaleBUN_log2:Age_category515`,
    Creat_adults = b_scaleCreatinine_fold_change_Pottel_CG+`b_Age_category>15:scaleCreatinine_fold_change_Pottel_CG`,
    Creat_U2 = b_scaleCreatinine_fold_change_Pottel_CG,
    Creat_25 = b_scaleCreatinine_fold_change_Pottel_CG+`b_Age_category25:scaleCreatinine_fold_change_Pottel_CG`,
    Creat_515 = b_scaleCreatinine_fold_change_Pottel_CG+`b_Age_category515:scaleCreatinine_fold_change_Pottel_CG`
  ) %>%
  select(BUN_adults, BUN_U2, BUN_25, BUN_515,
         Creat_adults, Creat_U2, Creat_25, Creat_515) %>%
  pivot_longer(cols = everything(),
               names_to = c("Biomarker", "Age_category"),
               names_pattern = "(.*)_(.*)",
               values_to = "estimate")

summary_df <- marginal_effects %>%
  group_by(Biomarker, Age_category) %>%
  summarise(
    OR = median(exp(estimate)),
    lower = quantile(exp(estimate), 0.025),
    upper = quantile(exp(estimate), 0.975),
    .groups = "drop"
  ) %>%
  mutate(
    Age_category2 = 
      case_when(
        Age_category=='25' ~ '[2-5]',
        Age_category=='515'~ '(5-15]',
        Age_category=='U2' ~ '<2',
        Age_category=='adults' ~ '>15'), 
    Biomarker = case_when(
      Biomarker=='BUN' ~ 'BUN\n(absolute value)',
      Biomarker=='Creat' ~ 'Creatinine\n(Fold change from baseline)'),
    Age_category = factor(Age_category2, levels = c('<2','[2-5]','(5-15]','>15'))
  ) %>% arrange(Biomarker, Age_category)

ggplot(summary_df, aes(x = Age_category2, y = OR, ymin = lower, ymax = upper, color = Biomarker)) +
  geom_pointrange(position = position_dodge(width = 0.5)) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_y_log10()+
  coord_flip() +
  labs(
    y = "Odds ratio for death per standard deviation change",
    x='',
    title = "Multivariable model of death",
    color = "Biomarker"
  ) +
  theme_bw(base_size = 16)
```







Plasma HRP2
```{r PfHRP2_BUN_Creatinine, fig.width=12, fig.height=9}
library(quantreg)
library(splines)

hrp2_data = sm_data_analysis %>% filter(!is.na(PfHRP2_ng_ml), !is.na(Creatinine_umol_L) | !is.na(BUN)) %>%
  mutate(log10_hrp2 = log10(PfHRP2_ng_ml+1),
         log10_BUN = log10(BUN))
fit_rq <- rq(
  Creatinine_fold_change_Pottel_CG ~ bs(log10_hrp2, df = 5, degree = 3), 
  tau = 0.9, 
  data = hrp2_data
)
summary(fit_rq)
xs=seq(0,4.8, length.out=100)
ys=predict(fit_rq, newdata = data.frame(log10_hrp2=xs))
f_90_perc_hrp2 = approxfun(x = 10^xs, y = 2^ys)

p1=hrp2_data%>%filter(!is.na(Creatinine_umol_L))%>%
  ggplot(aes(x=PfHRP2_ng_ml+1, y = 2^Creatinine_fold_change_Pottel_CG, colour = STUDY))+
  geom_jitter(width = .05,height = .01,alpha=0.3)+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks,
                labels = trans_format("log10", math_format(10^.x)))+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks,
                labels = trans_format("log10", math_format(10^.x)))+
  theme_bw(base_size = 14)+geom_smooth(se = F, aes(group=NA))+
  stat_function(
    fun = f_90_perc_hrp2,
    color = "black",
    linewidth = 1.2, linetype='dashed')+
  xlab('Plasma PfHRP2 (ng/ml)')+ylab('Creatinine fold change')+
  theme(legend.position = c(0.02, 0.98),  # (x, y) in [0,1] relative to plot area
        legend.justification = c(0, 1))   # align legend's bottom-left to that point


fit_rq <- rq(
  log10_BUN ~ bs(log10_hrp2, df = 5, degree = 3), 
  tau = 0.9, 
  data = hrp2_data
)
summary(fit_rq)
xs=seq(0,4.8, length.out=100)
ys=predict(fit_rq, newdata = data.frame(log10_hrp2=xs))
f_90_perc_hrp2 = approxfun(x = 10^xs, y = 10^ys)


p2=sm_data_analysis %>% filter(!is.na(PfHRP2_ng_ml), !is.na(BUN)) %>%
  ggplot(aes(x=PfHRP2_ng_ml+1, y = BUN, colour = STUDY))+
  geom_jitter(width = .05,height = .01,alpha=0.3)+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks,
                labels = trans_format("log10", math_format(10^.x)))+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks,
                labels = trans_format("log10", math_format(10^.x)))+
  theme_bw(base_size = 14)+geom_smooth(se = F, aes(group=NA))+
  stat_function(
    fun = f_90_perc_hrp2,
    color = "black",
    linewidth = 1.2, linetype='dashed')+
  xlab('Plasma PfHRP2 (ng/ml)')+ylab('BUN (mmol/L)')+
  theme(legend.position = c(0.02, 0.98),  # (x, y) in [0,1] relative to plot area
        legend.justification = c(0, 1))   # align legend's bottom-left to that point

grid.arrange(p1,p2,nrow=1)

writeLines(sprintf('There are %s patients with plasma hrp2 and creatinine measured',
                   nrow( sm_data_analysis %>% filter(!is.na(PfHRP2_ng_ml), !is.na(Creatinine_umol_L)) )))

writeLines(sprintf('There are %s patients with plasma hrp2 and BUN measured',
                   nrow( sm_data_analysis %>% filter(!is.na(PfHRP2_ng_ml), !is.na(BUN)) )))
```



## Comparison with a lactate threshold of 5 mmol/L

```{r lactate}
mod_bun = lme4::glmer(Died ~ Age_category + Coma + as.numeric(BUN>10) + (1|STUDY), family = 'binomial', data = sm_data_analysis %>% filter(Age<15))
# summary(mod_bun)
fixef(mod_bun)

mod_creat = lme4::glmer(Died ~ Age_category + Coma + as.numeric(Creatinine_fold_change_Pottel_CG>log2(3)) + (1|STUDY), family = 'binomial', data = sm_data_analysis%>% filter(Age<15))
# summary(mod_creat)
fixef(mod_creat)

mod_lactate = lme4::glmer(Died ~ Age_category + Coma + as.numeric(Lactate>5) + (1|STUDY), family = 'binomial', data = sm_data_analysis%>% filter(Age<15))
# summary(mod_lactate)
fixef(mod_lactate)

mod_lactate2 = lme4::glmer(Died ~ Age_category + Coma + as.numeric(Lactate>5)*as.numeric(Hb<5) + (1|STUDY), family = 'binomial', data = sm_data_analysis%>% filter(Age<15))
summary(mod_lactate2)
fixef(mod_lactate2)

mod_both = lme4::glmer(Died ~ Age + Coma + as.numeric(Lactate>5) + as.numeric(BUN>10) + as.numeric(Creatinine_fold_change_Pottel_CG>log2(3))+ (1|STUDY), family = 'binomial', data = sm_data_analysis%>% filter(Age<5))
summary(mod_both)
fixef(mod_both)

mod_both = mgcv::gam(Died ~ s(Age) + Coma + as.numeric(Lactate>5) + as.numeric(BUN>10) + as.numeric(Creatinine_fold_change_Schwartz>log2(3))+ s(SITE_STUDY, bs='re'), 
                     family = 'binomial', data = sm_data_analysis%>% filter(Age<5))
summary(mod_both)
```


### KEMRI temporal trends

```{r kemri_temporal_trend}
KEMRI = sm_data_analysis %>% 
  filter(!is.na(Creatinine_umol_L), !is.na(date_admission), 
         STUDY=='Kilifi cohort')
KEMRI %>% ggplot(aes(x=year_admission, y = Creatinine_fold_change_Pottel_CG)) +
  geom_point()+
  geom_smooth()+theme_bw()+
  xlab('Date of admission')+
  ylab('Creatinine fold change from predicted baseline')
```




### Interactions?

```{r}
# mod=lme4::glmer(Died_28D ~ Coma*(BUN>10) + (1|STUDY), family='binomial',
#                 data= sm_data_analysis %>% filter(Continent=='AFRICA'))
# summary(mod)
# 
# mod=lme4::glmer(Died_28D ~ Coma*(Creatinine_fold_change>1) + (1|STUDY), family='binomial',
#                 data= sm_data_analysis %>% filter(Continent=='AFRICA'))
# summary(mod)
# 
# namazzi = sm_data %>% filter(STUDY=='Namazzi 2022')
# mod=glm(Died_28D ~ Coma*(BUN>10), family='binomial', data= namazzi)
# summary(mod)
# 
# mod=glm(Died_28D ~ Coma*Creatinine_umol_L, family='binomial', data= namazzi)
# summary(mod)
```



## BUN/Creatinine vs Potassium 

```{r bun_Cr_potassium, fig.width=10, fig.height=8}
xx = sm_data_analysis %>% filter(Potassium_mmol_L<=10, !is.na(Potassium_mmol_L)) %>%
  mutate(Hyperkalaemia=Potassium_mmol_L>6.5,
         Uraemia = BUN>10,
         Acidotic = ifelse(!is.na(Bicarbonate_mEq_L), Bicarbonate_mEq_L<10, `Base Excess_mmol_L`< -15))
table(xx$Acidotic)

table(xx$Age_category[xx$Uraemia & xx$Acidotic & xx$Hyperkalaemia])
table(xx$Age_category[!xx$Uraemia & xx$Acidotic & xx$Hyperkalaemia])
table(xx$Age_category[xx$Uraemia & !xx$Acidotic & xx$Hyperkalaemia])

table(Hyperkalaemia=xx$Hyperkalaemia, xx$Age<15)
aggregate(Hyperkalaemia ~ Age_category, data = xx, FUN = mean)

hyperkalaemia_dat = xx %>% filter(Potassium_mmol_L>6.5, Age<15)
table(hyperkalaemia_dat$Died_28D)
p0 = xx %>%
  ggplot(aes(x=Age, y = Potassium_mmol_L, colour = Age_Categories3))+
  geom_point(alpha=0.3, shape=4)+
  ylab('Potassium (mmol/L)')+xlab('Age (years)')+
  scale_x_sqrt()+
  theme_bw(base_size = 16)+geom_hline(yintercept = c(3.5, 6.5),color='black')+
  geom_smooth(se = F, aes(group = NA, colour = 'black'))+
  scale_colour_manual(
    name = "Age (years)",
    values = c("<5" = "orange", "[5,15]" = "darkred", ">15" = 'purple') 
  )+theme(legend.position = 'inside',legend.position.inside=c(0.85,.85),
          legend.text = element_text(size = 5),
          legend.title = element_text(size = 6))

p1= xx %>%
  ggplot(aes(x=BUN, y = Potassium_mmol_L, colour = Age_Categories3))+
  geom_point(alpha=0.3, shape=4)+
  ylab('Potassium (mmol/L)')+xlab('BUN (mmol/L)')+
  scale_x_log10(breaks=breaks, minor_breaks=minor_breaks)+
  theme_bw(base_size = 16)+geom_hline(yintercept = c(3.5, 6.5),color='black')+
  geom_smooth(se = F)+
  scale_colour_manual(
    name = "Age (years)",
    values = c("<5" = "orange", "[5,15]" = "darkred", ">15" = 'purple'))

p2=xx %>% 
  ggplot(aes(x=2^Creatinine_fold_change_Pottel_CG, y = Potassium_mmol_L, colour = Age_Categories3))+
  geom_point(alpha=0.3, shape=4)+geom_vline(xintercept = 3)+
  ylab('Potassium (mmol/L)')+xlab('Creatinine fold change')+
  scale_x_log10(breaks=breaks, minor_breaks=minor_breaks)+
  theme_bw(base_size = 16)+
  geom_hline(yintercept = c(3.5, 6.5),color='black')+
  geom_smooth(se = F)+
  scale_colour_manual(
    name = "Age (years)",
    values = c("<5" = "orange", "[5,15]" = "darkred", ">15" = 'purple') 
  )

p3= xx %>% filter(Bicarbonate_mEq_L<32) %>%
  ggplot(aes(x=Bicarbonate_mEq_L, y = Potassium_mmol_L, colour = Age_Categories3))+
  geom_point(alpha=0.3, shape=4)+
  xlab('Bicarbonate (mEq/L)')+ylab('Potassium (mmol/L)')+
  theme_bw(base_size = 16)+geom_hline(yintercept = c(3.5, 6.5),color='black')+
  geom_smooth(se = F)+
  scale_colour_manual(
    name = "Age (years)",
    values = c("<5" = "orange", "[5,15]" = "darkred", ">15" = 'purple') 
  )

p4= xx %>%
  ggplot(aes(x=Lactate, y = Potassium_mmol_L, colour = Age_Categories3))+
  geom_point(alpha=0.3, shape=4)+
  xlab('Lactate (mmol/L)')+ylab('Potassium (mmol/L)')+
  theme_bw(base_size = 16)+geom_hline(yintercept = c(3.5, 6.5),color='black')+
  geom_smooth(se = F)+
  scale_colour_manual(
    name = "Age (years)",
    values = c("<5" = "orange", "[5,15]" = "darkred", ">15" = 'purple') 
  )


xx_complete = xx %>% filter(!is.na(Acidotic), !is.na(Uraemia)) 
tab <- xx_complete %>%
  group_by(Acidotic, Uraemia) %>%                 
  summarise(
    n     = n(),                  
    propz = mean(Hyperkalaemia, na.rm = TRUE),
    .groups = "drop"
  )

p5=ggplot(tab, aes(x = Acidotic, y = Uraemia, fill = propz)) +
  geom_tile(color = "grey80") +
  geom_text(
    aes(label = sprintf("%1.1f%%\n(n = %d)", 100 * propz, n)),
    size = 3
  ) +
  scale_fill_gradient(
    name   = "Hyperkalaemia",
    low    = "white",
    high   = "red",
    limits = c(0, .2),     
    labels = percent_format(accuracy = 1)
  ) +
  coord_equal() +
  labs(x = "Bicarbonate < 10 mEq/L", y = "BUN>10 mmol/L") +
  theme_bw(base_size = 16)


grid.arrange(p0+theme(legend.position = 'none'),
             p1+theme(legend.position = 'none'),
             p2+theme(legend.position = 'none'),
             p3+theme(legend.position = 'none'),
             p4+theme(legend.position = 'none'),
             p5)

mod_hyperkalaemia1 = lme4::glmer(Hyperkalaemia ~ Age_Categories3+(Creatinine_fold_change_Pottel_CG>log2(3))+(Bicarbonate_mEq_L<10)+(1|SITEID), family = 'binomial', data = xx %>% filter(Age<15))
summary(mod_hyperkalaemia1)

mod_hyperkalaemia2 = lme4::glmer(Hyperkalaemia ~ Age_Categories3+(BUN>10)+(Bicarbonate_mEq_L<10)+(1|SITEID), family = 'binomial', data = xx %>% filter(Age<15))
summary(mod_hyperkalaemia2)

mod_hyperkalaemia_death = lme4::glmer(Died_28D ~ Hyperkalaemia+ (Creatinine_fold_change_Pottel_CG>log2(3))+(Age<2)+(Bicarbonate_imputed<10)+(1|STUDY/SITE_STUDY), family = 'binomial', data = xx %>% filter(Age<15, Bicarbonate_imputed<30))
summary(mod_hyperkalaemia_death)

grid.arrange(p0, p1,p2)

aggregate(Hyperkalaemia ~ Age_category + (BUN>10) + (Bicarbonate_imputed<10), data = xx, function(x) round(100*mean(x),1))


p0= xx %>%
  ggplot(aes(x=Age, y = Potassium_mmol_L))+
  geom_point(alpha=0.3, shape=4)+
  ylab('Potassium (mmol/L)')+xlab('Age (years)')+
  scale_x_sqrt()+
  theme_bw(base_size = 16)+geom_hline(yintercept = c(3.5, 6.5),color='red')+
  geom_smooth(se = F)
p1= xx %>%
  ggplot(aes(x=BUN, y = Potassium_mmol_L))+
  geom_point(alpha=0.3, shape=4)+
  ylab('Potassium (mmol/L)')+xlab('BUN (mmol/L)')+
  scale_x_log10(breaks=breaks, minor_breaks=minor_breaks)+
  theme_bw(base_size = 18)+geom_hline(yintercept = c(3.5, 6.5),color='red')+
  geom_smooth(se = F)
p2= xx %>%
  ggplot(aes(x=Bicarbonate_imputed, y = Potassium_mmol_L))+
  geom_point(alpha=0.3, shape=4)+
  xlab('Bicarbonate (mEq/L)')+ylab('Potassium (mmol/L)')+
  theme_bw(base_size = 18)+geom_hline(yintercept = c(3.5, 6.5),color='red')+
  geom_smooth(se = F)
p3= xx %>%
  ggplot(aes(x=Lactate_imputed, y = Potassium_mmol_L))+
  geom_point(alpha=0.3, shape=4)+
  xlab('Lactate (mmol/L)')+ylab('Potassium (mmol/L)')+
  theme_bw(base_size = 18)+geom_hline(yintercept = c(3.5, 6.5),color='red')+
  geom_smooth(se = F)
grid.arrange(p0,p1,p2,p3)
```



# Age based thresholds

```{r}
df <- data.frame(
  age       = c(0, 2, 6, 10, 14, 17, 100),
  threshold = c(60, 80, 105, 130, 160, 180, 180)
)
sm_data_creat = sm_data_creat %>% 
  mutate(creatinine_threshold = case_when(
    Age<2 ~ 60,
    Age>=2 & Age<6 ~ 80,
    Age>=6 & Age<10 ~ 100,
    Age>=10 & Age<15 ~ 130,
    Age>=15 ~ 180
  ))
sm_data_creat %>% 
  ggplot(aes(x=Age, y = Creatinine_umol_L, 
             colour = Creatinine_fold_change_Pottel_CG>log2(3)))+
  geom_point(alpha=0.5)+
  geom_step(color = "#185FA5", linewidth = 1, 
            aes(x=Age, y = creatinine_threshold)) +
  scale_y_log10()+scale_x_sqrt()+theme_minimal()+
  xlab('Age (years)')+ylab('Creatinine (umol/L)')
```

