---
title: "Renal Analysis Severe Malaria"
author: "James Watson"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = T, cache.comments = FALSE,
                      echo = F, include = TRUE, 
                      fig.width = 8, fig.height = 8,
                      fig.pos = 'H',dev = 'png', dpi = 300)

library(brms)
library(tidybayes)
library(tidyverse)
library(grid)
library(gridExtra)
library(rstanarm)
library(scales)
```



Plotting params

```{r}
source('setup_parameters.R')
RUN_STAN_MODELS=F
```


## Load analysis data

Analysis dataset
```{r}
load('RData/Analysis_dataset.RData')

sm_data_analysis = sm_data_analysis %>%
  mutate(Age_Categories3 = cut(Age,breaks=c(0,5,15,100)),
         Age_Categories3 = fct_recode(
           Age_Categories3,
           "<5"     = "(0,5]",
           "[5,15]" = "(5,15]",
           ">15"    = "(15,100]"
         ))

# Estimate baseline creatinine using the Schwartz formula
baseline_creatinine_schwartz <- function(height_cm) {
  egfr_normal = 120
  creat <- 0.413 * (height_cm / egfr_normal) * 88.4
  return(creat)
}

baseline_creatinine_pottel = function(age_years){
  egfr_normal = 120
  Q_age <- 0.243 * (age_years ^ 0.244)        # age-based reference value
  creat_mg_dl <- 107.3 * Q_age / egfr_normal  # baseline creatinine (mg/dL)
  creat_umol  <- creat_mg_dl * 88.4           # convert to µmol/L
  return(creat_umol)
}

baseline_creatinine_cockroftgault = function(age_years, weight_kg){
  creat_umol = 88.4*((140-age_years)*weight_kg)/(120*72)
  return(creat_umol)
}
baseline_creatinine_pottel(5)
baseline_creatinine_schwartz(100)
baseline_creatinine_cockroftgault(age_years = 40, weight_kg = 75)
```


# Calculate expected Cr based on age, weight

## Baseline creatinines

We look at the data from the TRACT study, Malawi site where we have follow-up creatinines and BUN in patients who survived and were not lost to follow-up. We compare these values to the expected values from the Pottel equation.

```{r baseline_creatinine_BUN, fig.width=12, fig.height=9}
baseline_data = sm_data %>% filter(!is.na(Baseline_Creatinine) | !is.na(`Baseline_Urea Nitrogen`)) %>%
  mutate(Baseline_Pottel = baseline_creatinine_pottel(Age),
         Baseline_Schwartz = baseline_creatinine_schwartz(Height_imputed),
         Outliers = Baseline_Creatinine > Baseline_Pottel+15) 

# p1= baseline_data %>%
#   ggplot(aes(x=Weight_kg, y = Baseline_Creatinine))+geom_point()+
#   theme_bw()+xlab('Weight (kg)')+ylab('Normal creatinine (umol/L)')+
#   geom_smooth(se = F)

p1= baseline_data %>%
  ggplot(aes(x=Age, y = Baseline_Creatinine))+geom_point(shape = 3)+
  stat_function(
    fun = baseline_creatinine_pottel,
    color = "red",
    linewidth = 1.2)+
  theme_bw(base_size = 16)+ xlab('Age (years)')+ylab('Normal creatinine (umol/L)')+
  geom_smooth(se = F)
p2=sm_data %>% filter(!is.na(Baseline_Creatinine)) %>%
  ggplot(aes(x=Height_cm, y = Baseline_Creatinine))+geom_point(shape = 3)+
  theme_bw(base_size = 16)+
  xlab('Height (cm)')+ylab('Normal creatinine (umol/L)')+geom_smooth(se=F)+
  stat_function(
    fun = baseline_creatinine_schwartz,
    color = "red",
    linewidth = 1.2
  )
# p4=sm_data %>% filter(!is.na(Baseline_Creatinine)) %>%
#   ggplot(aes(x=wfaz, y = Baseline_Creatinine))+geom_point()+
#   theme_bw()+
#   xlab('Weight-for-age z-score')+ylab('Normal creatinine (umol/L)')+
#   geom_smooth(se = F)


summary(lm(Baseline_Creatinine~Height_cm, data = sm_data))
summary(lm(Baseline_Creatinine~wfaz, data = sm_data))
summary(lm(Baseline_Creatinine~Age, data = sm_data))
summary(lm(Baseline_Creatinine~Weight_kg, data = sm_data))
eGFR=120 # Using Pottel equation

##########** BUN BASELINES **** ######
p3=sm_data %>% filter(!is.na(`Baseline_Urea Nitrogen`)) %>%
  ggplot(aes(x=Age, y = `Baseline_Urea Nitrogen`))+geom_point(shape = 3)+
  theme_bw(base_size = 16)+
  xlab('Age (years)')+ylab('Normal BUN (mmol/L)')+geom_smooth(se = F)


p4=sm_data %>% filter(!is.na(`Baseline_Urea Nitrogen`)) %>%
  ggplot(aes(x=Height_imputed, y = `Baseline_Urea Nitrogen`))+geom_point(shape = 3)+
  theme_bw(base_size = 18)+
  xlab('Height (cm)')+ylab('Normal BUN (mmol/L)')+geom_smooth(se = F)
mod_bun_baseline_age = mgcv::gam(b_bun ~ s(sqrt(Age),k=5), data=baseline_data %>% mutate(b_bun=`Baseline_Urea Nitrogen`))
round(predict(mod_bun_baseline_age, newdata = data.frame(Age=c(0.5,3,10))),1)


grid.arrange(p1,p2,p3,p4)

# 
# xx = sm_data %>% filter(!is.na(Baseline_Creatinine) | !is.na(`Baseline_Urea Nitrogen`))
# xx$MUAC = xx$`Mid-Upper Arm Circumference_cm`
# xx$Baseline_BUN = xx$`Baseline_Urea Nitrogen`
# mod_creat_baseline=stan_lm(Baseline_Creatinine ~ Age+Height_cm+Weight_kg,
#                            data = xx, prior = R2(0.5))
# plot(mod_creat_baseline,pars = c('Age','Height_cm','Weight_kg'))
# summary(mod_creat_baseline)
# 
# mod_bun_baseline=stan_lm(Baseline_BUN ~ scale(Age) + scale(Weight_kg)+scale(Height_cm)+SEX+scale(MUAC),
#                          data = xx, prior = R2(0.5))
# plot(mod_bun_baseline,pars = c('scale(Age)','scale(Weight_kg)','scale(Height_cm)','SEXM','scale(MUAC)'))
# summary(mod_bun_baseline)
baseline_data = baseline_data %>% filter(!Outliers) %>%
  mutate(Residual_Pottel = Baseline_Creatinine-Baseline_Pottel,
         Residual_Schwartz = Baseline_Creatinine-Baseline_Schwartz)
ylims = range(c(baseline_data$Residual_Pottel, baseline_data$Residual_Schwartz))+c(-1,1)

p1= baseline_data %>% 
  ggplot(aes(x=Baseline_Creatinine, y = Residual_Pottel))+geom_point()+
  scale_x_log10()+
  theme_bw(base_size = 18)+ xlab('Baseline creatinine (umol/L)')+ylab('Residual: Baseline vs Pottel predicted')+
  geom_smooth(se = F, method = lm)+ylim(ylims[1], ylims[2])+geom_hline(yintercept = c(-10,10))

p2= baseline_data %>% 
  ggplot(aes(x=Baseline_Creatinine, y = Residual_Schwartz))+geom_point()+
  scale_x_log10()+
  theme_bw(base_size = 18)+ xlab('Baseline creatinine (umol/L)')+ylab('Residual: Baseline vs Schwartz predicted')+
  geom_smooth(se = F, method = lm)+ylim(ylims[1], ylims[2])+geom_hline(yintercept = c(-10,10))
grid.arrange(p1,p2,nrow=1)
```

Pottel age-based equation, where eGFR=107.3/(SCr/Q)
assuming a normal GFR of 120 mL/min per 1.73 m2 (height-independent). 
Q=0.0270*age+0.2329.

We estimate the Pottel baseline creatinine (age only for children)
We estimate the adult baseline creatine using Cockcroft-Gault equation
We also estimate across the age spectrum using the Full Age equation (Pottel 2016)


```{r}
sm_data_analysis = sm_data_analysis %>%
  mutate(
    # Height_m = Height_imputed/100,
    # Height_m = ifelse(Height_m<0.8, 0.8, Height_m),
    # Q2=3.94-13.4*Height_m+17.6*Height_m^2-9.84*Height_m^3+2.04*Height_m^4,
    Baseline_Pottel = baseline_creatinine_pottel(age_years = Age),
    Baseline_CG = baseline_creatinine_cockroftgault(age_years = Age, weight_kg = Weight_kg),
    Baseline_Cr_Pottel_CG = ifelse(Age<15, Baseline_Pottel, Baseline_CG),
    # Baseline_Cr_Height = ifelse(Age<15, (107.3/120)*Q2*88.4, bSrc_adults),
    Baseline_Cr_Schwartz = baseline_creatinine_schwartz(Height_imputed))

sm_data_analysis %>% ggplot(aes(x=Age, y = Baseline_Pottel))+geom_point()+
  geom_point(aes(x=Age, y=Baseline_Cr_Schwartz, colour = 'red',alpha=.5))+
  theme_bw()

# # Make the Fold Change values for each patients
sm_data_analysis = sm_data_analysis %>%
  mutate(Creatinine_fold_change_Pottel_CG = log2(Creatinine_umol_L/Baseline_Cr_Pottel_CG),
         Creatinine_fold_change_Schwartz = log2(Creatinine_umol_L/Baseline_Cr_Schwartz))

p1=sm_data_analysis %>% ggplot(aes(x=Age, y=Baseline_Cr_Pottel_CG))+geom_point()+
  theme_bw()+ylim(0,120)+ylab('Pottel/CG predicted baseline Creatinine (umol/L)')
p2=sm_data_analysis %>% ggplot(aes(x=Age, y=Baseline_Cr_Schwartz))+geom_point()+
  theme_bw()+ylim(0,120)+ylab('Schwartz predicted baseline Creatinine (umol/L)')

grid.arrange(p1,p2)
```


TRACT predicted baseline creatinine for children. Age predicted values look better than the height predicted values.

```{r tract_baseline_Cr}
p1=sm_data_analysis %>% filter(!is.na(Baseline_Creatinine)) %>%
  ggplot(aes(x=Baseline_Creatinine, y = Baseline_Cr_Pottel_CG))+geom_point()+
  xlab('Observed baseline Creatinine in TRACT')+
  ylab('Age (Pottel equation) predicted baseline Creatinine (umol/L)')+
  theme_bw()+geom_abline(intercept = 0, slope = 1)

p2=sm_data_analysis %>% filter(!is.na(Baseline_Creatinine)) %>%
  ggplot(aes(x= Baseline_Creatinine, y= Baseline_Cr_Schwartz))+geom_point()+
  theme_bw()+geom_abline(intercept = 0, slope = 1)+
  ylab('Height (Schwartz equation) predicted baseline Creatinine (umol/L)')+
  xlab('Observed baseline Creatinine in TRACT')
grid.arrange(p1,p2)
```


is there heterogeneity by age/weight?

```{r creatinine_fold_change}
sm_data_creat = sm_data_analysis %>% filter(!is.na(Creatinine_umol_L))
p0 = sm_data_creat %>% ggplot(aes(x=Creatinine_fold_change_Pottel_CG, y = Creatinine_fold_change_Schwartz, colour = Age_category))+
  geom_point()+theme_bw()+xlab('Creatinine fold change (Potel or CG)')+
  ylab('Creatinine fold change (Schwartz)')+
  geom_abline(slope = 1, intercept = 0)

p1=sm_data_creat %>% ggplot(aes(x=Age, y = Creatinine_fold_change_Pottel_CG, colour = STUDY))+
  geom_point()+theme_bw(base_size = 14)+
  geom_smooth(aes(group=NA))+ylab('Creatinine fold change (Potel or CG)')
p2=sm_data_creat %>% ggplot(aes(x=Age, y = Creatinine_fold_change_Schwartz, colour = STUDY))+
  geom_point()+theme_bw(base_size = 14)+
  geom_smooth(aes(group=NA))+ylab('Creatinine fold change (Schwartz)')

grid.arrange(p0,p1,p2)
```


```{r age_vs_Cr_foldchange, fig.width=10, fig.height=7}
sm_data_creat %>% 
  ggplot(aes(x=Age, y = 2^Creatinine_fold_change_Pottel_CG))+
  geom_point()+ scale_y_log10()+scale_x_sqrt()+
  geom_smooth()+theme_bw(base_size = 12)+
  ylab('Creatinine fold change (Pottel/CG)')+
  facet_wrap(~STUDY,nrow=2)
```



When using the creatinine fold change as a predictor of death there is still some residual heterogeneity. This seems to be entirely driven by the TRACT study data.

```{r mgcv_tensor_product_creatinine}
par(las=1)
# Fit GAM with age-varying spline for fold-change creatinine
m = mgcv::gam(
  Died_28D ~ s(Age, k = 6) + 
    te(Creatinine_fold_change_Pottel_CG, Age, bs = c("cr", "cr"), k = c(10, 6))+
    Coma + s(SITE_STUDY,bs='re'),
  data   = sm_data_creat,
  family = binomial(link = "logit"),
  method = "REML",
  select = TRUE
)
mgcv::vis.gam(m, view = c("Creatinine_fold_change_Pottel_CG", "Age"),
              plot.type = "contour", color = "topo", too.far = 0.05,
              main='Predicted baseline using Age/Weight')

# Fit GAM with age-varying spline for fold-change creatinine - without TRACT
m_rmTRACT = mgcv::gam(
  Died_28D ~ s(Age, k = 6) + 
    te(Creatinine_fold_change_Pottel_CG, Age, bs = c("cr", "cr"), k = c(10, 6))+
    Coma + s(SITE_STUDY,bs='re'),
  data   = sm_data_creat %>% filter(STUDY != 'TRACT'),
  family = binomial(link = "logit"),
  method = "REML",
  select = TRUE
)
mgcv::vis.gam(m_rmTRACT, view = c("Creatinine_fold_change_Pottel_CG", "Age"),
              plot.type = "contour", color = "topo", too.far = 0.05,
              main='Predicted baseline using Age/Weight\nwithout TRACT')
```


```{r}
#tensor product models
Creat_tensor_prod_Pottel_CG = as.formula(bf(Died_28D ~ t2(Creatinine_fold_change_Pottel_CG,Age,bs = c("cr", "cr"))+Coma+s(year_admission)+(1|STUDY/SITE_STUDY)))
Creat_tensor_prod_Schwartz = as.formula(bf(Died_28D ~ t2(Creatinine_fold_change_Schwartz,Age,bs = c("cr", "cr"))+Coma+s(year_admission)+(1|STUDY/SITE_STUDY)))
Creat_tensor_prod_abs = as.formula(bf(Died_28D ~ t2(log10(Creatinine_umol_L),Age,bs = c("cr", "cr"))+Coma+s(year_admission)+(1|STUDY/SITE_STUDY)))

#smooth spline models with categorical age
Creat_fs_mod_Pottel_CG = as.formula(bf(Died_28D ~ s(Creatinine_fold_change_Pottel_CG, bs = "cr", k = 10) + s(Creatinine_fold_change_Pottel_CG,Age_category,bs = "fs", m = 1, k = 10)+Age_category+Coma+s(year_admission)+(1|STUDY/SITE_STUDY)))
Creat_fs_mod_Schwartz = as.formula(bf(Died_28D ~ s(Creatinine_fold_change_Schwartz, bs = "cr", k = 10) + s(Creatinine_fold_change_Schwartz,Age_category,bs = "fs", m = 1, k = 10)+Age_category+Coma+s(year_admission)+(1|STUDY/SITE_STUDY)))

f_path = 'RData/creatinine_models.RData'
if(!file.exists(f_path) | RUN_STAN_MODELS){
  niter = 2000
  Creat_tensor_prod_fit_Pottel_CG = 
    brm(formula = Creat_tensor_prod_Pottel_CG,
        data = sm_data_creat ,
        family = bernoulli(), cores = 4, seed = 17,
        iter = niter, warmup = niter/2, thin = 1, refresh = 250)
  
  Creat_tensor_prod_fit_Schwartz = 
    brm(formula = Creat_tensor_prod_Schwartz,
        data = sm_data_creat ,
        family = bernoulli(), cores = 4, seed = 17,
        iter = niter, warmup = niter/2, thin = 1, refresh = 250)  
  Creat_tensor_prod_fit_abs = 
    brm(formula = Creat_tensor_prod_abs,
        data = sm_data_creat ,
        family = bernoulli(), cores = 4, seed = 17,
        iter = niter, warmup = niter/2, thin = 1, refresh = 250)  
  Creat_fs_mod_fit_Pottel_CG = 
    brm(formula = Creat_fs_mod_Pottel_CG,
        data = sm_data_creat ,
        family = bernoulli(), cores = 4, seed = 17,
        iter = niter, warmup = niter/2, thin = 1, refresh = 250)  
  Creat_fs_mod_fit_Schwartz = 
    brm(formula = Creat_fs_mod_Schwartz,
        data = sm_data_creat ,
        family = bernoulli(), cores = 4, seed = 17,
        iter = niter, warmup = niter/2, thin = 1, refresh = 250)  
  
  save(Creat_tensor_prod_fit_Pottel_CG, Creat_tensor_prod_fit_Schwartz, Creat_tensor_prod_fit_abs, Creat_fs_mod_fit_Pottel_CG, Creat_fs_mod_fit_Schwartz, file = f_path)
} else {
  load(f_path)
}
```


summary of model fits

```{r}
summary(Creat_tensor_prod_fit_Pottel_CG)
summary(Creat_tensor_prod_fit_Schwartz)
summary(Creat_tensor_prod_fit_abs)
summary(Creat_fs_mod_fit_Pottel_CG)
summary(Creat_fs_mod_fit_Schwartz)
```


Estimated study-level and study-site-level random intercepts for the two primary
models, shown side by side. The urea model is loaded here, so `BUN_Analysis.Rmd`
must have been run first.

```{r nested_random_effects_setup}
load('RData/bun_models.RData')   # BUN_tensor_prod_fit, for the right-hand panels

RANEF_MODELS = list(
  'Creatinine fold change' = Creat_tensor_prod_fit_Pottel_CG,
  'Urea (BUN)'             = BUN_tensor_prod_fit)

f_get_ranef = function(mod, model_name, group){
  re = as.data.frame(ranef(mod)[[group]][, , 'Intercept'])
  re$level = rownames(re)
  re$model = model_name
  re
}

# Nested factor levels have the form "<STUDY>_<SITE>_<STUDY>"; strip the
# redundant study name from both ends to leave a readable "STUDY: SITE" label.
f_site_label = function(level, studies){
  vapply(level, function(l){
    st   = studies[startsWith(l, paste0(studies, '_'))]
    st   = st[which.max(nchar(st))]          # longest matching study prefix
    site = sub(paste0('^\\Q', st, '\\E_'), '', l)
    site = sub(paste0('_\\Q', st, '\\E$'), '', site)
    paste0(st, ': ', site)
  }, character(1), USE.NAMES = FALSE)
}

# The two models contain different studies and sites, so the y variable is made
# unique per panel and the panel prefix is stripped again at the axis.
f_order_within = function(d){
  d %>%
    mutate(y = paste(model, label, sep = '___')) %>%
    arrange(model, Estimate) %>%
    mutate(y = factor(y, levels = unique(y)))
}
f_strip_panel = function(x) sub('^.*?___', '', x)

f_caterpillar = function(d, colour_by_study = FALSE){
  aes_base = if(colour_by_study){
    aes(x = Estimate, y = y, colour = STUDY)
  } else {
    aes(x = Estimate, y = y)
  }
  p = ggplot(d, aes_base) +
    geom_vline(xintercept = 0, linetype = 'dashed', colour = 'grey40') +
    geom_errorbar(aes(xmin = Q2.5, xmax = Q97.5), orientation = 'y', width = 0) +
    geom_point(size = 2.2) +
    facet_wrap(~model, scales = 'free_y') +
    scale_y_discrete(labels = f_strip_panel) +
    theme_bw(base_size = 13) +
    theme(panel.grid.major.y = element_blank(), legend.position = 'none') +
    xlab('Log odds ratio for death (deviation from overall intercept)') +
    ylab('')
  if(colour_by_study) p = p + scale_colour_viridis_d(end = 0.9)
  p
}
```

Study-level random intercepts (S18 Fig):

```{r nested_random_effects_study, fig.width=9, fig.height=4.5}
ranef_study = bind_rows(lapply(names(RANEF_MODELS), function(nm)
  f_get_ranef(RANEF_MODELS[[nm]], nm, 'STUDY'))) %>%
  rename(label = level) %>%
  f_order_within()

f_caterpillar(ranef_study)
```

Study-site-level random intercepts, nested within study (S19 Fig):

```{r nested_random_effects_site, fig.width=10, fig.height=8}
ranef_site = bind_rows(lapply(names(RANEF_MODELS), function(nm){
  mod     = RANEF_MODELS[[nm]]
  studies = unique(as.character(mod$data$STUDY))
  d       = f_get_ranef(mod, nm, 'STUDY:SITE_STUDY')
  d$label = f_site_label(d$level, studies)
  d$STUDY = sub(':.*$', '', d$label)
  d
})) %>%
  f_order_within()

f_caterpillar(ranef_site, colour_by_study = TRUE)

# Range of site-level effects quoted in the Results
ranef_site %>% group_by(model) %>%
  summarise(n_site   = n(),
            min_lor  = round(min(Estimate), 2),
            max_lor  = round(max(Estimate), 2),
            OR_range = round(exp(max(Estimate) - min(Estimate)), 1),
            .groups  = 'drop')
```

### Variance components for all fitted models

Posterior summaries of the nested random-effect standard deviations, with the
corresponding median odds ratios (S1 Table). The median odds ratio is the median
of the distribution of odds ratios obtained by comparing two patients with
identical covariates drawn from two randomly chosen clusters, ranking the
higher-risk patient against the lower-risk one; for a random-intercept standard
deviation $\sigma$ it equals $\exp\{\sqrt{2}\,\sigma\,\Phi^{-1}(0.75)\}$. The
"both levels" column applies that formula to the total standard deviation, and so
compares two sites in two different studies.

```{r variance_components}
f_vc_ci  = function(x) sprintf('%.2f (%.2f-%.2f)', median(x), quantile(x, .025), quantile(x, .975))
f_mor    = function(sigma) exp(sqrt(2) * sigma * qnorm(0.75))

f_variance_components = function(m){
  d  = brms::as_draws_df(m)
  s1 = d[['sd_STUDY__Intercept']]
  s2 = d[['sd_STUDY:SITE_STUDY__Intercept']]
  st = sqrt(s1^2 + s2^2)
  data.frame(n         = nobs(m),
             n_study   = length(unique(m$data$STUDY)),
             n_site    = length(unique(m$data$SITE_STUDY)),
             sd_study  = f_vc_ci(s1),
             sd_site   = f_vc_ci(s2),
             sd_total  = f_vc_ci(st),
             ICC       = f_vc_ci((s1^2 + s2^2) / (s1^2 + s2^2 + pi^2/3)),
             MOR_study = f_vc_ci(f_mor(s1)),
             MOR_site  = f_vc_ci(f_mor(s2)),
             MOR_both  = f_vc_ci(f_mor(st)))
}

VC_MODELS = list(
  `Cr FC (Pottel/CG), t2` = Creat_tensor_prod_fit_Pottel_CG,
  `Cr FC (Schwartz), t2`  = Creat_tensor_prod_fit_Schwartz,
  `Cr absolute, t2`       = Creat_tensor_prod_fit_abs,
  `Cr FC (Pottel/CG), fs` = Creat_fs_mod_fit_Pottel_CG,
  `Cr FC (Schwartz), fs`  = Creat_fs_mod_fit_Schwartz,
  `BUN, t2`               = BUN_tensor_prod_fit,
  `BUN, fs`               = BUN_fs_mod_fit)

do.call(rbind, lapply(VC_MODELS, f_variance_components))
```


Make some test data

```{r}
k_vals=40
test_data = expand_grid(Creatinine_fold_change_Pottel_CG = seq(0,3,length.out=k_vals),
                        Age = c(2, 5, 20),
                        year_admission=2020,
                        Coma=unique(sm_data_creat$Coma),
                        STUDY=unique(sm_data_creat$STUDY), 
                        SITE_STUDY = -1) %>% arrange(Creatinine_fold_change_Pottel_CG) %>%
  mutate(Creatinine_fold_change_Schwartz=Creatinine_fold_change_Pottel_CG,
         Age_category=cut(Age,breaks = c(0, 2, 5, 15, 100)),
         Age_category = recode(Age_category,
                               "(0,2]" = "≤2",
                               "(2,5]" = "(2,5]",
                               "(5,15]" = "(5,15]",
                               "(15,100]" = ">15"))

test_data2 = expand_grid(Creatinine_umol_L = 10^seq(1,3,length.out=k_vals),
                         Age = c(2, 5, 20),
                         year_admission=2020,
                         Coma=unique(sm_data_creat$Coma),
                         STUDY=unique(sm_data_creat$STUDY), 
                         SITE_STUDY = -1) %>% arrange(Creatinine_umol_L) %>%
  mutate(Age_category=cut(Age,breaks = c(0, 2, 5, 15, 100)),
         Age_category = recode(Age_category,
                               "(0,2]" = "≤2",
                               "(2,5]" = "(2,5]",
                               "(5,15]" = "(5,15]",
                               "(15,100]" = ">15"))

# remove non plausible combinations
# test_data = test_data %>%
#   filter(!(Age_category==">15" & Weight_category=='(0,25]'),
#          !(Age_category %in% c("(2,5]","≤2") & Weight_category=='(25,100]')) %>%
#   mutate(Age_category=factor(Age_category, levels = c("≤2","(2,5]","(5,15]",">15")))

# get model predictions
# Full model
preds_Pottel_CG = posterior_epred(Creat_tensor_prod_fit_Pottel_CG, newdata = test_data, re_formula = NA)
preds_Schwartz = posterior_epred(Creat_tensor_prod_fit_Schwartz, newdata = test_data, re_formula = NA)
test_data$Creat_tensor_prod_pred_Pottel_CG = colMeans(preds_Pottel_CG)
test_data$Creat_tensor_prod_pred_Schwartz = colMeans(preds_Schwartz)
test_data$Creat_tensor_prod_threshold_Pottel_CG = apply(preds_Pottel_CG, 2, f_threshold,mortality_threshold=mortality_threshold)
test_data$Creat_tensor_prod_threshold_Schwartz = apply(preds_Schwartz, 2, f_threshold,mortality_threshold=mortality_threshold)

# Full model - abs cr values
preds = posterior_epred(Creat_tensor_prod_fit_abs, newdata = test_data2, re_formula = NA)
test_data2$Creat_abs_tensor_prod_pred = colMeans(preds)
test_data2$Creat_abs_tensor_prod_threshold = apply(preds, 2, f_threshold,mortality_threshold=mortality_threshold)

preds_Pottel_CG = posterior_epred(Creat_fs_mod_fit_Pottel_CG, newdata = test_data, re_formula = NA)
preds_Schwartz = posterior_epred(Creat_fs_mod_fit_Schwartz, newdata = test_data, re_formula = NA)
test_data$Creat_fs_pred_Pottel_CG = colMeans(preds_Pottel_CG)
test_data$Creat_fs_pred_Schwartz = colMeans(preds_Schwartz)
test_data$Creat_fs_threshold_Pottel_CG = apply(preds_Pottel_CG, 2, f_threshold,mortality_threshold=mortality_threshold)
test_data$Creat_fs_threshold_Schwartz = apply(preds_Schwartz, 2, f_threshold,mortality_threshold=mortality_threshold)
```


# The effect of year
```{r temporal_trend}
test_data_year = expand_grid(Creatinine_fold_change_Pottel_CG = c(1,3),
                             Age = 2,
                             year_admission=1990:2020,
                             Coma='No',
                             STUDY=unique(sm_data_creat$STUDY), 
                             SITE_STUDY = -1) %>% arrange(Creatinine_fold_change_Pottel_CG) %>%
  mutate(Creatinine_fold_change_Schwartz=Creatinine_fold_change_Pottel_CG,
         Age_category=cut(Age,breaks = c(0, 2, 5, 15, 100)),
         Age_category = recode(Age_category,
                               "(0,2]" = "≤2",
                               "(2,5]" = "(2,5]",
                               "(5,15]" = "(5,15]",
                               "(15,100]" = ">15"))


preds_Pottel_CG = posterior_epred(Creat_tensor_prod_fit_Pottel_CG, newdata = test_data_year, re_formula = NA)
test_data_year$pred_year = colMeans(preds_Pottel_CG)
test_data_year %>% ggplot(aes(x=year_admission, y = pred_year, 
                              group = Creatinine_fold_change_Pottel_CG))+
  geom_line(linewidth = 1.5)+ ylab('Predicted mortality')+
  theme_bw(base_size = 15)+
  xlab('Year')+
  geom_hline(yintercept = mortality_threshold)+
  scale_colour_manual(
    name = "Age (years)",
    values = c("2" = "orange", "5" = "darkred", "20" = 'purple') )+
  theme(legend.position = c(0.02, 0.98),  # (x, y) in [0,1] relative to plot area
        legend.justification = c(0, 1))   # align legend's bottom-left to that point
```



Using absolute creatinine

```{r creatinine_abs_tensor}
p0c=test_data2 %>% mutate(Age = as.factor(Age))%>%
  ggplot(aes(x = Creatinine_umol_L, y = Creat_abs_tensor_prod_pred, 
             colour =Age,linetype = Coma,
             group = interaction(Age,Coma)))+
  geom_line(linewidth = 1.5)+ ylab('Predicted mortality')+
  scale_x_log10() +
  annotation_logticks(sides='b')+theme_bw(base_size = 15)+
  xlab('Creatinine (umol/L)')+
  geom_hline(yintercept = mortality_threshold)+
  scale_colour_manual(
    name = "Age (years)",
    values = c("2" = "orange", "5" = "darkred", "20" = 'purple') )+
  theme(legend.position = c(0.02, 0.98),  # (x, y) in [0,1] relative to plot area
        legend.justification = c(0, 1))   # align legend's bottom-left to that point


p0c_threshold=test_data2 %>%mutate(Age = as.factor(Age))%>%
  ggplot(aes(x = Creatinine_umol_L, y = Creat_abs_tensor_prod_threshold, 
             colour = Age,linetype = Coma,
             group = interaction(Age,Coma)))+
  geom_line(linewidth = 1.5)+ylab('Probability mortality > 5%')+
  scale_x_log10() + theme_bw(base_size = 15)+
  annotation_logticks(sides='b')+
  scale_colour_manual(
    name = "Age (years)",
    values = c("2" = "orange", "5" = "darkred", "20" = 'purple') )+
  xlab('Creatinine (umol/L)')+
  geom_hline(yintercept = .9)+
  theme(legend.position = 'none')
grid.arrange(p0c, p0c_threshold)
```


Using creatinine fold change based on age and weight (adults) - Pottel equation and CG equation

```{r creatinine_fc_baseline_mortality_Pottel}
p1c=test_data %>% mutate(Age = as.factor(Age))%>%
  ggplot(aes(x = 2^Creatinine_fold_change_Pottel_CG, y = Creat_tensor_prod_pred_Pottel_CG, 
             colour =Age,linetype = Coma,
             group = interaction(Age,Coma)))+
  geom_line(linewidth = 1.5)+ ylab('Predicted mortality')+
  scale_x_log10() +
  annotation_logticks(sides='b')+theme_bw(base_size = 15)+
  xlab('Creatinine fold change')+
  geom_hline(yintercept = mortality_threshold)+
  scale_colour_manual(
    name = "Age (years)",
    values = c("2" = "orange", "5" = "darkred", "20" = 'purple') )+
  theme(legend.position = c(0.02, 0.98),  # (x, y) in [0,1] relative to plot area
        legend.justification = c(0, 1))   # align legend's bottom-left to that point


p1c_threshold=test_data %>%mutate(Age = as.factor(Age))%>%
  ggplot(aes(x = 2^Creatinine_fold_change_Pottel_CG, y = Creat_tensor_prod_threshold_Pottel_CG, 
             colour = Age,linetype = Coma,
             group = interaction(Age,Coma)))+
  geom_line(linewidth = 1.5)+ylab('Probability mortality > 5%')+
  scale_x_log10() + theme_bw(base_size = 15)+
  annotation_logticks(sides='b')+
  scale_colour_manual(
    name = "Age (years)",
    values = c("2" = "orange", "5" = "darkred", "20" = 'purple') )+
  xlab('Creatinine fold change')+
  geom_hline(yintercept = .9)+geom_vline(xintercept = 3)+
  theme(legend.position = 'none')



p2c=test_data %>% mutate(Age = as.factor(Age))%>%
  ggplot(aes(x = 2^Creatinine_fold_change_Pottel_CG, y = Creat_fs_pred_Pottel_CG, 
             colour =Age,linetype = Coma,
             group = interaction(Age,Coma)))+
  geom_line(linewidth = 1.5)+ ylab('Predicted mortality')+
  scale_x_log10() +
  annotation_logticks(sides='b')+theme_bw(base_size = 15)+
  xlab('Creatinine fold change')+
  geom_hline(yintercept = mortality_threshold)+
  labs(colour='Age')+
  theme(legend.position = c(0.02, 0.98),  # (x, y) in [0,1] relative to plot area
        legend.justification = c(0, 1))   # align legend's bottom-left to that point


p2c_threshold=test_data %>%mutate(Age = as.factor(Age))%>%
  ggplot(aes(x = 2^Creatinine_fold_change_Pottel_CG, y = Creat_fs_threshold_Pottel_CG, 
             colour = Age,linetype = Coma,
             group = interaction(Age,Coma)))+
  geom_line(linewidth = 1.5)+ylab('Probability mortality > 5%')+
  scale_x_log10() + theme_bw(base_size = 15)+
  annotation_logticks(sides='b')+
  xlab('Creatinine fold change')+
  geom_hline(yintercept = .9)+geom_vline(xintercept = 3)+
  theme(legend.position = 'none')

grid.arrange(p1c, p1c_threshold, p2c, p2c_threshold)


sm_data_analysis %>% filter(!is.na(Creatinine_umol_L)) %>%
  ggplot(aes(x=Age, y = Creatinine_umol_L, colour = Creatinine_fold_change_Pottel_CG>log2(3)))+
  geom_point()+scale_y_log10(breaks = breaks, minor_breaks = minor_breaks)+
  geom_hline(yintercept = 265,aes(linetype = 'dashed'))+
  scale_x_sqrt()+theme_bw(base_size = 14)+
  geom_vline(xintercept = c(2, 5,15))
```


Combined figure with BUN for paper

```{r combined_fig, fig.width=12, fig.height=10}
load('RData/bun_figures.RData')
# combined fig 3
grid.arrange(p0c+geom_vline(xintercept = c(265, 150)), p1c+theme(legend.position = 'none'), 
             p1+theme(legend.position = 'none'),p0c_threshold, 
             p1c_threshold, p1_threshold, nrow=2)
```



Using creatinine fold change based on Height

```{r creatinine_fc_baseline_mortality_Schwartz}
p1c=test_data %>% mutate(Age = as.factor(Age))%>%
  ggplot(aes(x = 2^Creatinine_fold_change_Schwartz, y = Creat_tensor_prod_pred_Schwartz, 
             colour =Age,linetype = Coma,
             group = interaction(Age,Coma)))+
  geom_line(linewidth = 1.5)+ ylab('Predicted mortality')+
  scale_x_log10() +
  annotation_logticks(sides='b')+theme_bw(base_size = 15)+
  xlab('Creatinine fold change (Schwartz)')+
  geom_hline(yintercept = mortality_threshold)+
  scale_colour_manual(
    name = "Age (years)",
    values = c("2" = "orange", "5" = "darkred", "20" = 'purple') )+
  theme(legend.position = c(0.02, 0.98),  # (x, y) in [0,1] relative to plot area
        legend.justification = c(0, 1))   # align legend's bottom-left to that point


p1c_threshold=test_data %>%mutate(Age = as.factor(Age))%>%
  ggplot(aes(x = 2^Creatinine_fold_change_Schwartz, y = Creat_tensor_prod_threshold_Schwartz, 
             colour = Age,linetype = Coma,
             group = interaction(Age,Coma)))+
  geom_line(linewidth = 1.5)+ylab('Probability mortality > 5%')+
  scale_x_log10() + theme_bw(base_size = 15)+
  annotation_logticks(sides='b')+
  scale_colour_manual(
    name = "Age (years)",
    values = c("2" = "orange", "5" = "darkred", "20" = 'purple') )+
  xlab('Creatinine fold change (Schwartz)')+
  geom_hline(yintercept = .9)+geom_vline(xintercept = 3)+
  theme(legend.position = 'none')



p2c=test_data %>% mutate(Age = as.factor(Age))%>%
  ggplot(aes(x = 2^Creatinine_fold_change_Schwartz, y = Creat_fs_pred_Schwartz, 
             colour =Age,linetype = Coma,
             group = interaction(Age,Coma)))+
  geom_line(linewidth = 1.5)+ ylab('Predicted mortality')+
  scale_x_log10() +
  annotation_logticks(sides='b')+theme_bw(base_size = 15)+
  xlab('Creatinine fold change (Schwartz)')+
  geom_hline(yintercept = mortality_threshold)+
  labs(colour='Age')+
  theme(legend.position = c(0.02, 0.98),  # (x, y) in [0,1] relative to plot area
        legend.justification = c(0, 1))   # align legend's bottom-left to that point


p2c_threshold=test_data %>%mutate(Age = as.factor(Age))%>%
  ggplot(aes(x = 2^Creatinine_fold_change_Schwartz, y = Creat_fs_threshold_Schwartz, 
             colour = Age,linetype = Coma,
             group = interaction(Age,Coma)))+
  geom_line(linewidth = 1.5)+ylab('Probability mortality > 5%')+
  scale_x_log10() + theme_bw(base_size = 15)+
  annotation_logticks(sides='b')+
  xlab('Creatinine fold change (Schwartz)')+
  geom_hline(yintercept = .9)+geom_vline(xintercept = 3)+
  theme(legend.position = 'none')

grid.arrange(p1c, p1c_threshold, p2c, p2c_threshold)


sm_data_analysis %>% filter(!is.na(Creatinine_umol_L)) %>%
  ggplot(aes(x=Age, y = Creatinine_umol_L, colour = Creatinine_fold_change_Schwartz>log2(3)))+
  geom_point()+scale_y_log10(breaks = breaks, minor_breaks = minor_breaks)+
  geom_hline(yintercept = 265,aes(linetype = 'dashed'))+
  scale_x_sqrt()+theme_bw(base_size = 14)+
  geom_vline(xintercept = c(2, 5,15))


sm_data_analysis %>% filter(!is.na(Creatinine_umol_L)) %>%
  ggplot(aes(x=Age, y = Creatinine_fold_change_Pottel_CG- Creatinine_fold_change_Schwartz))+
  geom_point()+
  scale_x_sqrt()+theme_bw(base_size = 14)+
  geom_vline(xintercept = c(2, 5,15))
```



Model using age as categorical variable

```{r creatinine_fs_mod}
p1=test_data %>%
  ggplot(aes(x = 2^Creatinine_fold_change_Pottel_CG, y = Creat_fs_pred_Pottel_CG, 
             colour = Age_category,linetype = Coma,
             group = interaction(Age_category,Coma)))+
  geom_line()+ ylab('Predicted mortality')+
  scale_x_log10() +
  annotation_logticks(sides='b')+theme_bw(base_size = 15)+
  xlab('Creatinine fold change')+
  geom_hline(yintercept = mortality_threshold)+
  labs(colour='Age category')

p1_threshold=test_data %>%
  ggplot(aes(x = 2^Creatinine_fold_change_Pottel_CG, y = Creat_fs_threshold_Pottel_CG, 
             colour = Age_category,linetype = Coma,
             group = interaction(Age_category,Coma)))+
  geom_line()+ylab('Probability mortality > 5%')+
  scale_x_log10() +
  geom_hline(yintercept = .9)+geom_vline(xintercept = 3)+
  annotation_logticks(sides='b')+theme_bw(base_size = 15)+
  xlab('Creatinine fold change')
grid.arrange(p1, p1_threshold)
```



### Compare creatinine fold change with BUN

Compare BUN versus Creatinine when both measured

```{r}
sm_both = sm_data_analysis %>% filter(!is.na(Creatinine_umol_L), !is.na(BUN))

# children under 5
mod_both_U5 = lme4::glmer(Died_28D~Age+scale(log10(BUN))+scale(Creatinine_fold_change_Pottel_CG)+Coma+(1|SITE_STUDY),
                          family='binomial', data = sm_both %>% filter(Age<=5))
summary(mod_both_U5)

# children 5-15
mod_both_5_15 = lme4::glmer(Died_28D~Age+scale(log10(BUN))+scale(Creatinine_fold_change_Pottel_CG)+Coma+(1|SITE_STUDY),
                            family='binomial', data = sm_both %>% filter(Age>5, Age<15))
summary(mod_both_5_15)

# Adults
mod_both_adults = lme4::glmer(Died_28D~Age+scale(log10(BUN))+scale(Creatinine_fold_change_Pottel_CG)+Coma+(1|SITE_STUDY),
                              family='binomial', data = sm_both %>% filter(Age>15))
summary(mod_both_adults)

### Fit interaction model
mod_both_all = lme4::glmer(Died_28D~Age_category*scale(log10(BUN))+Age_category*scale(Creatinine_fold_change_Pottel_CG)+Coma+(1|SITE_STUDY),
                           family='binomial', data = sm_both)
summary(mod_both_all)

sm_both = sm_both %>% mutate(BUN_log2 = log2(BUN))

f_brm_mod_both = as.formula(bf(Died_28D ~ BUN_log2*Age_category +
                                 Creatinine_fold_change_Pottel_CG*Age_category+
                                 Coma+(1|STUDY/SITE_STUDY)))

f_brm_mod_both_scaled = as.formula(bf(Died_28D ~ scale(BUN_log2)*Age_category +
                                        scale(Creatinine_fold_change_Pottel_CG)*Age_category+
                                        Coma+(1|STUDY/SITE_STUDY)))
f_path = 'RData/Model_both.RData'
if(RUN_STAN_MODELS | !file.exists(f_path)){
  mod_both_stan = brm(formula = f_brm_mod_both,
                      data = sm_both,
                      family = bernoulli(), cores = 4, seed = 17,
                      iter = 2000, warmup = 1000, thin = 1, refresh = 250)
  
  mod_both_stan_scaled = brm(formula = f_brm_mod_both_scaled,
                             data = sm_both,
                             family = bernoulli(), cores = 4, seed = 17,
                             iter = 2000, warmup = 1000, thin = 1, refresh = 250)
  
  save(mod_both_stan,mod_both_stan_scaled, file = f_path)
} else{
  load(f_path)
}
brms::mcmc_plot(mod_both_stan)
```


```{r marginal_effects_children_adults, fig.width=10, fig.height=6}
posterior <- brms::as_draws_df(mod_both_stan)
marginal_effects <- posterior %>%
  mutate(
    BUN_adults = b_BUN_log2+`b_BUN_log2:Age_category>15`,
    BUN_U2 = b_BUN_log2,
    BUN_25 = b_BUN_log2+`b_BUN_log2:Age_category25`,
    BUN_515 = b_BUN_log2+`b_BUN_log2:Age_category515`,
    Creat_adults = b_Creatinine_fold_change_Pottel_CG+`b_Age_category>15:Creatinine_fold_change_Pottel_CG`,
    Creat_U2 = b_Creatinine_fold_change_Pottel_CG,
    Creat_25 = b_Creatinine_fold_change_Pottel_CG+`b_Age_category25:Creatinine_fold_change_Pottel_CG`,
    Creat_515 = b_Creatinine_fold_change_Pottel_CG+`b_Age_category515:Creatinine_fold_change_Pottel_CG`
  ) %>%
  select(BUN_U2, BUN_25, BUN_515,BUN_adults, 
         Creat_U2, Creat_25, Creat_515, Creat_adults) %>%
  pivot_longer(cols = everything(),
               names_to = c("Biomarker", "Age_category"),
               names_pattern = "(.*)_(.*)",
               values_to = "estimate")

summary_df <- marginal_effects %>%
  group_by(Biomarker, Age_category) %>%
  summarise(
    OR = median(exp(estimate)),
    lower = quantile(exp(estimate), 0.025),
    upper = quantile(exp(estimate), 0.975),
    .groups = "drop"
  ) %>%
  mutate(
    Age_category2 =
      case_when(
        Age_category=='25' ~ '[2-5]',
        Age_category=='515'~ '(5-15]',
        Age_category=='U2' ~ '<2',
        Age_category=='adults' ~ '>15'),
    Biomarker = case_when(
      Biomarker=='BUN' ~ 'BUN\n(absolute value)',
      Biomarker=='Creat' ~ 'Creatinine\n(Fold change from baseline)'),
    # The plot maps x = Age_category2, so the ordering has to be set on that
    # column. coord_flip() draws the first factor level at the bottom of the
    # panel, hence levels run oldest to youngest to put <2 years at the top.
    Age_category2 = factor(Age_category2, levels = c('>15','(5-15]','[2-5]','<2')),
    Age_category = Age_category2
  ) %>% arrange(Biomarker, desc(Age_category2))

ggplot(summary_df, aes(x = Age_category2, y = OR, ymin = lower, ymax = upper, color = Biomarker)) +
  geom_pointrange(position = position_dodge(width = 0.5)) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_y_log10()+
  coord_flip() +
  labs(
    y = "Odds ratio for death per doubling",
    x='',
    title = "Multivariable model of death",
    color = "Biomarker"
  ) +
  theme_bw(base_size = 16)
```


scaled predictors:

```{r marginal_effects_children_adults_scaled, fig.width=10, fig.height=6}
posterior <- brms::as_draws_df(mod_both_stan_scaled)
marginal_effects <- posterior %>%
  mutate(
    BUN_adults = b_scaleBUN_log2+`b_scaleBUN_log2:Age_category>15`,
    BUN_U2 = b_scaleBUN_log2,
    BUN_25 = b_scaleBUN_log2+`b_scaleBUN_log2:Age_category25`,
    BUN_515 = b_scaleBUN_log2+`b_scaleBUN_log2:Age_category515`,
    Creat_adults = b_scaleCreatinine_fold_change_Pottel_CG+`b_Age_category>15:scaleCreatinine_fold_change_Pottel_CG`,
    Creat_U2 = b_scaleCreatinine_fold_change_Pottel_CG,
    Creat_25 = b_scaleCreatinine_fold_change_Pottel_CG+`b_Age_category25:scaleCreatinine_fold_change_Pottel_CG`,
    Creat_515 = b_scaleCreatinine_fold_change_Pottel_CG+`b_Age_category515:scaleCreatinine_fold_change_Pottel_CG`
  ) %>%
  select(BUN_adults, BUN_U2, BUN_25, BUN_515,
         Creat_adults, Creat_U2, Creat_25, Creat_515) %>%
  pivot_longer(cols = everything(),
               names_to = c("Biomarker", "Age_category"),
               names_pattern = "(.*)_(.*)",
               values_to = "estimate")

summary_df <- marginal_effects %>%
  group_by(Biomarker, Age_category) %>%
  summarise(
    OR = median(exp(estimate)),
    lower = quantile(exp(estimate), 0.025),
    upper = quantile(exp(estimate), 0.975),
    .groups = "drop"
  ) %>%
  mutate(
    Age_category2 =
      case_when(
        Age_category=='25' ~ '[2-5]',
        Age_category=='515'~ '(5-15]',
        Age_category=='U2' ~ '<2',
        Age_category=='adults' ~ '>15'),
    Biomarker = case_when(
      Biomarker=='BUN' ~ 'BUN\n(absolute value)',
      Biomarker=='Creat' ~ 'Creatinine\n(Fold change from baseline)'),
    # The plot maps x = Age_category2, so the ordering has to be set on that
    # column. coord_flip() draws the first factor level at the bottom of the
    # panel, hence levels run oldest to youngest to put <2 years at the top.
    Age_category2 = factor(Age_category2, levels = c('>15','(5-15]','[2-5]','<2')),
    Age_category = Age_category2
  ) %>% arrange(Biomarker, desc(Age_category2))

ggplot(summary_df, aes(x = Age_category2, y = OR, ymin = lower, ymax = upper, color = Biomarker)) +
  geom_pointrange(position = position_dodge(width = 0.5)) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_y_log10()+
  coord_flip() +
  labs(
    y = "Odds ratio for death per standard deviation change",
    x='',
    title = "Multivariable model of death",
    color = "Biomarker"
  ) +
  theme_bw(base_size = 16)
```



### ROC curves

```{r roc_creatinine_bun}
library(pROC)
# Fit ROC for each predictor
roc_creat   <- roc(sm_both$Died_28D, sm_both$Creatinine_fold_change_Pottel_CG)
roc_bun     <- roc(sm_both$Died_28D, sm_both$BUN)
roc_lactate <- roc(sm_both$Died_28D, sm_both$Lactate)

# Combined plot
ggroc(list(
  Creatinine = roc_creat,
  BUN        = roc_bun,
  Lactate    = roc_lactate
), linewidth = 1) +
  geom_abline(intercept = 1, slope = 1, linetype = "dashed", color = "gray60") +
  scale_colour_manual(
    name   = NULL,
    values = c(
      Creatinine = "#185FA5",
      BUN        = "#D85A30",
      Lactate    = "#1D9E75"
    ),
    labels = c(
      paste0("Creatinine fold-change (AUC = ", round(auc(roc_creat),   3), ")"),
      paste0("BUN (AUC = ",                    round(auc(roc_bun),     3), ")"),
      paste0("Lactate (AUC = ",                round(auc(roc_lactate), 3), ")")
    )
  ) +
  labs(x = "1 - Specificity", y = "Sensitivity",
       title = "ROC curves for 28-day mortality") +
  theme_minimal(base_size = 13) +
  theme(legend.position = c(0.72, 0.18),
        legend.background = element_rect(fill = "white", color = NA))
```






## Plasma HRP2
```{r PfHRP2_BUN_Creatinine, fig.width=12, fig.height=9}
library(quantreg)
library(splines)

hrp2_data = sm_data_analysis %>% filter(!is.na(PfHRP2_ng_ml), !is.na(Creatinine_umol_L) | !is.na(BUN)) %>%
  mutate(log10_hrp2 = log10(PfHRP2_ng_ml+1),
         log10_BUN = log10(BUN))
fit_rq <- rq(
  Creatinine_fold_change_Pottel_CG ~ bs(log10_hrp2, df = 5, degree = 3), 
  tau = 0.9, 
  data = hrp2_data
)
summary(fit_rq)
xs=seq(0,4.8, length.out=100)
ys=predict(fit_rq, newdata = data.frame(log10_hrp2=xs))
f_90_perc_hrp2 = approxfun(x = 10^xs, y = 2^ys)

p1=hrp2_data%>%filter(!is.na(Creatinine_umol_L))%>%
  ggplot(aes(x=PfHRP2_ng_ml+1, y = 2^Creatinine_fold_change_Pottel_CG, colour = STUDY))+
  geom_jitter(width = .05,height = .01,alpha=0.3)+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks,
                labels = trans_format("log10", math_format(10^.x)))+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks,
                labels = trans_format("log10", math_format(10^.x)))+
  theme_bw(base_size = 14)+geom_smooth(se = F, aes(group=NA))+
  stat_function(
    fun = f_90_perc_hrp2,
    color = "black",
    linewidth = 1.2, linetype='dashed')+
  xlab('Plasma PfHRP2 (ng/ml)')+ylab('Creatinine fold change')+
  theme(legend.position = c(0.02, 0.98),  # (x, y) in [0,1] relative to plot area
        legend.justification = c(0, 1))   # align legend's bottom-left to that point


fit_rq <- rq(
  log10_BUN ~ bs(log10_hrp2, df = 5, degree = 3), 
  tau = 0.9, 
  data = hrp2_data
)
summary(fit_rq)
xs=seq(0,4.8, length.out=100)
ys=predict(fit_rq, newdata = data.frame(log10_hrp2=xs))
f_90_perc_hrp2 = approxfun(x = 10^xs, y = 10^ys)


p2=sm_data_analysis %>% filter(!is.na(PfHRP2_ng_ml), !is.na(BUN)) %>%
  ggplot(aes(x=PfHRP2_ng_ml+1, y = BUN, colour = STUDY))+
  geom_jitter(width = .05,height = .01,alpha=0.3)+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks,
                labels = trans_format("log10", math_format(10^.x)))+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks,
                labels = trans_format("log10", math_format(10^.x)))+
  theme_bw(base_size = 14)+geom_smooth(se = F, aes(group=NA))+
  stat_function(
    fun = f_90_perc_hrp2,
    color = "black",
    linewidth = 1.2, linetype='dashed')+
  xlab('Plasma PfHRP2 (ng/ml)')+ylab('BUN (mmol/L)')+
  theme(legend.position = c(0.02, 0.98),  # (x, y) in [0,1] relative to plot area
        legend.justification = c(0, 1))   # align legend's bottom-left to that point

grid.arrange(p1,p2,nrow=1)

writeLines(sprintf('There are %s patients with plasma hrp2 and creatinine measured',
                   nrow( sm_data_analysis %>% filter(!is.na(PfHRP2_ng_ml), !is.na(Creatinine_umol_L)) )))

writeLines(sprintf('There are %s patients with plasma hrp2 and BUN measured',
                   nrow( sm_data_analysis %>% filter(!is.na(PfHRP2_ng_ml), !is.na(BUN)) )))
```



## Comparison with a lactate threshold of 5 mmol/L

```{r lactate}
# Helper: adjusted odds ratios with 95% CIs on the Wald scale, for reporting
f_or_table = function(mod, terms = NULL){
  co = lme4::fixef(mod)
  ci = suppressMessages(confint(mod, method = 'Wald'))
  keep = if(is.null(terms)) names(co)[names(co) != '(Intercept)'] else terms
  out = data.frame(
    term = keep,
    OR    = round(exp(co[keep]), 2),
    lower = round(exp(ci[keep, 1]), 2),
    upper = round(exp(ci[keep, 2]), 2),
    row.names = NULL)
  out$OR_CI = sprintf('%.2f (%.2f-%.2f)', out$OR, out$lower, out$upper)
  out
}

mod_bun = lme4::glmer(Died_28D ~ Age_category + Coma + as.numeric(BUN>10) + (1|STUDY), family = 'binomial', data = sm_data_analysis %>% filter(Age<15))
# summary(mod_bun)
fixef(mod_bun)

mod_creat = lme4::glmer(Died_28D ~ Age_category + Coma + as.numeric(Creatinine_fold_change_Pottel_CG>log2(3)) + (1|STUDY), family = 'binomial', data = sm_data_analysis%>% filter(Age<15))
# summary(mod_creat)
fixef(mod_creat)

mod_lactate = lme4::glmer(Died_28D ~ Age_category + Coma + as.numeric(Lactate>5) + (1|STUDY), family = 'binomial', data = sm_data_analysis%>% filter(Age<15))
# summary(mod_lactate)
fixef(mod_lactate)

mod_lactate2 = lme4::glmer(Died_28D ~ Age_category + Coma + as.numeric(Lactate>5)*as.numeric(Hb<5) + (1|STUDY), family = 'binomial', data = sm_data_analysis%>% filter(Age<15))
summary(mod_lactate2)
fixef(mod_lactate2)

mod_both = lme4::glmer(Died_28D ~ Age + Coma + as.numeric(Lactate>5) + as.numeric(BUN>10) + as.numeric(Creatinine_fold_change_Pottel_CG>log2(3))+ (1|STUDY), family = 'binomial', data = sm_data_analysis%>% filter(Age<5))
summary(mod_both)
fixef(mod_both)

mod_both_gam = mgcv::gam(Died_28D ~ s(Age) + Coma + as.numeric(Lactate>5) + as.numeric(BUN>10) + as.numeric(Creatinine_fold_change_Schwartz>log2(3))+ s(SITE_STUDY, bs='re'),
                     family = 'binomial', data = sm_data_analysis%>% filter(Age<5))
summary(mod_both_gam)
```

### Reported comparison of urea >10 mmol/L versus lactate >5 mmol/L

This is the model quoted in the Results: all ages, age as a continuous variable,
adjusted for coma status, with nested random intercepts for study and study site,
fitted to the patients with both a lactate and a urea measurement.

```{r lactate_vs_bun_reported}
sm_lactate_bun = sm_data_analysis %>% filter(!is.na(Lactate), !is.na(BUN))

writeLines(sprintf('n=%s patients had both lactate and urea measured (n=%s children <15 years)',
                   nrow(sm_lactate_bun), sum(sm_lactate_bun$Age < 15)))

mod_lactate_comp = lme4::glmer(
  Died_28D ~ Coma + as.numeric(Lactate>5) + Age_category +
    (1|STUDY/SITE_STUDY),
  family = 'binomial', data = sm_lactate_bun %>% filter(Age<15))

mod_bun_comp = lme4::glmer(
  Died_28D ~ Coma + as.numeric(BUN>10) +
    (1|STUDY/SITE_STUDY),
  family = 'binomial', data = sm_lactate_bun %>% filter(Age<15))

mod_bun_lactate_comp = lme4::glmer(
  Died_28D ~ Coma + as.numeric(Lactate>5) + as.numeric(BUN>10) +
    (1|STUDY/SITE_STUDY),
  family = 'binomial', data = sm_lactate_bun %>% filter(Age<15))

summary(mod_lactate_comp)
summary(mod_bun_comp)
summary(mod_bun_lactate_comp)

print(f_or_table(mod_lactate_comp,     terms = c('as.numeric(Lactate > 5)', 'ComaYes')))
print(f_or_table(mod_bun_comp,         terms = c('as.numeric(BUN > 10)', 'ComaYes')))
print(f_or_table(mod_bun_lactate_comp,
                 terms = c('as.numeric(BUN > 10)', 'as.numeric(Lactate > 5)', 'ComaYes')))
```


### Diagnostic accuracy of urea >10 mmol/L, lactate >5 mmol/L and creatinine >3x baseline

Sensitivity, specificity and discrimination for death by day 28 for the three
candidate severe malaria thresholds.

Panel A reports each marker on every patient in whom that marker was measured.
This uses all the available data, but the three rows then refer to different
patients, contributed by different studies and with different underlying
mortality, so **the rows are not directly comparable with each other**. The
predictive values in particular depend on the mortality of whichever population
contributed that marker. Panels B and C give the valid head-to-head comparisons,
restricted to the patients in whom the markers being compared were both (or all)
measured; a paired test of two AUCs requires the same patients. Confidence
intervals for differences in AUC come both from the DeLong method, which treats
patients as independent, and from resampling whole studies, which reflects the
clustering in these pooled data.

This chunk uses `Creatinine_fold_change_Pottel_CG`, created earlier in this
document.

```{r diagnostic_accuracy_markers}
library(pROC)

acc_data = sm_data_analysis %>% mutate(died = Died_28D == 1)

# Each marker: the continuous variable used for the ROC, and the binary rule.
# Note the creatinine fold change is already on the log2 scale, so the ">3x
# baseline" rule is expressed as > log2(3).
MARKERS = list(
  list(key = 'Urea',       label = 'Urea (BUN) >10 mmol/L',
       var = 'BUN',
       rule = function(d) d$BUN > 10),
  list(key = 'Lactate',    label = 'Lactate >5 mmol/L',
       var = 'Lactate',
       rule = function(d) d$Lactate > 5),
  list(key = 'Creatinine', label = 'Creatinine >3x baseline',
       var = 'Creatinine_fold_change_Pottel_CG',
       rule = function(d) d$Creatinine_fold_change_Pottel_CG > log2(3))
)

# Exact (Clopper-Pearson) intervals for each proportion
f_ci_prop = function(x, n) if(n == 0) c(NA, NA) else as.numeric(binom.test(x, n)$conf.int)
f_pc      = function(p, lo, hi) sprintf('%.1f (%.1f-%.1f)', 100*p, 100*lo, 100*hi)

f_marker_perf = function(pos, died, label){
  tp = sum(pos & died);  fp = sum(pos & !died)
  fn = sum(!pos & died); tn = sum(!pos & !died)
  sens = tp/(tp+fn); spec = tn/(tn+fp); ppv = tp/(tp+fp); npv = tn/(tn+fn)
  data.frame(
    Test        = label,
    n           = length(died),
    deaths      = sum(died),
    pct_died    = sprintf('%.1f', 100*mean(died)),
    pct_flagged = sprintf('%.1f', 100*mean(pos)),
    Sensitivity = do.call(f_pc, c(list(sens), as.list(f_ci_prop(tp, tp+fn)))),
    Specificity = do.call(f_pc, c(list(spec), as.list(f_ci_prop(tn, tn+fp)))),
    PPV         = do.call(f_pc, c(list(ppv),  as.list(f_ci_prop(tp, tp+fp)))),
    NPV         = do.call(f_pc, c(list(npv),  as.list(f_ci_prop(tn, tn+fn)))),
    Youden_J    = sprintf('%.3f', sens + spec - 1))
}

f_auc_str = function(died, marker){
  ci = as.numeric(ci.auc(roc(died, marker, quiet = TRUE), method = 'delong'))
  sprintf('%.3f (%.3f-%.3f)', ci[2], ci[1], ci[3])
}

## ---- A. each marker on all patients in whom it was measured ---------------
writeLines('===== A. Each marker, all patients in whom that marker was measured =====')
tab_A = bind_rows(lapply(MARKERS, function(m){
  s = acc_data[!is.na(acc_data[[m$var]]), ]
  p = f_marker_perf(m$rule(s), s$died, m$label)
  p$AUC = f_auc_str(s$died, s[[m$var]])
  p
}))
print(as.data.frame(tab_A), row.names = FALSE)

writeLines('\nStudies contributing to each population:')
for(m in MARKERS){
  s = acc_data[!is.na(acc_data[[m$var]]), ]
  writeLines(sprintf('  %-11s %2d studies: %s', m$key, length(unique(s$STUDY)),
                     paste(sort(unique(as.character(s$STUDY))), collapse = ', ')))
}

## ---- B. pairwise head-to-head comparisons ---------------------------------
writeLines('\n===== B. Pairwise comparisons (patients with both markers of the pair) =====')
set.seed(20250826)
B_boot = 2000
for(pr in list(c(1,2), c(1,3), c(2,3))){
  m1 = MARKERS[[pr[1]]]; m2 = MARKERS[[pr[2]]]
  s  = acc_data[!is.na(acc_data[[m1$var]]) & !is.na(acc_data[[m2$var]]), ]
  r1 = roc(s$died, s[[m1$var]], quiet = TRUE)
  r2 = roc(s$died, s[[m2$var]], quiet = TRUE)
  dl = roc.test(r1, r2, method = 'delong', paired = TRUE)
  auc_diff = as.numeric(auc(r1)) - as.numeric(auc(r2))
  studies_ = unique(s$STUDY)
  boot_diff = replicate(B_boot, {
    bs = bind_rows(lapply(sample(studies_, length(studies_), replace = TRUE),
                          function(x) s[s$STUDY == x, ]))
    if(length(unique(bs$died)) < 2) return(NA_real_)
    tryCatch(as.numeric(auc(roc(bs$died, bs[[m1$var]], quiet = TRUE))) -
             as.numeric(auc(roc(bs$died, bs[[m2$var]], quiet = TRUE))),
             error = function(e) NA_real_)
  })
  boot_diff = boot_diff[!is.na(boot_diff)]
  writeLines(sprintf('\n  %s vs %s (n=%s, %s deaths)', m1$key, m2$key,
                     format(nrow(s), big.mark = ','), sum(s$died)))
  writeLines(sprintf('    AUC %.3f vs %.3f | difference %+.3f | DeLong p = %.3g | study-clustered 95%% CI %.3f to %.3f',
                     as.numeric(auc(r1)), as.numeric(auc(r2)), auc_diff, dl$p.value,
                     quantile(boot_diff, .025), quantile(boot_diff, .975)))
}

## ---- C. three-way comparison, all three measured --------------------------
writeLines('\n===== C. Three-way comparison (patients with all three measured) =====')
acc3 = acc_data %>% filter(!is.na(BUN), !is.na(Lactate),
                           !is.na(Creatinine_fold_change_Pottel_CG))
writeLines(sprintf('  n = %s (%s deaths, %.1f%%)', format(nrow(acc3), big.mark = ','),
                   sum(acc3$died), 100*mean(acc3$died)))
tab_C = bind_rows(lapply(MARKERS, function(m){
  p = f_marker_perf(m$rule(acc3), acc3$died, m$label)
  p$AUC = f_auc_str(acc3$died, acc3[[m$var]])
  p
}))
print(as.data.frame(tab_C)[, c('Test','pct_flagged','Sensitivity','Specificity',
                               'PPV','NPV','Youden_J','AUC')], row.names = FALSE)

## ---- D. by age group, each marker on its own population -------------------
writeLines('\n===== D. By age group, each marker on its own population =====')
for(nm in c('children <15 y', 'adults >=15 y')){
  in_age = if(nm == 'children <15 y') acc_data$Age < 15 else acc_data$Age >= 15
  writeLines(sprintf('\n-- %s --', nm))
  tab_D = bind_rows(lapply(MARKERS, function(m){
    s = acc_data[in_age & !is.na(acc_data[[m$var]]), ]
    p = f_marker_perf(m$rule(s), s$died, m$label)
    p$AUC = f_auc_str(s$died, s[[m$var]])
    p
  }))
  print(as.data.frame(tab_D)[, c('Test','n','deaths','pct_flagged','Sensitivity',
                                 'Specificity','PPV','NPV','AUC')], row.names = FALSE)
}
```


### KEMRI temporal trends

```{r kemri_temporal_trend}
KEMRI = sm_data_analysis %>% 
  filter(!is.na(Creatinine_umol_L), !is.na(date_admission), 
         STUDY=='Kilifi cohort')
KEMRI %>% ggplot(aes(x=year_admission, y = Creatinine_fold_change_Pottel_CG)) +
  geom_point()+
  geom_smooth()+theme_bw()+
  xlab('Date of admission')+
  ylab('Creatinine fold change from predicted baseline')
```




### Interactions?

```{r}
# mod=lme4::glmer(Died_28D ~ Coma*(BUN>10) + (1|STUDY), family='binomial',
#                 data= sm_data_analysis %>% filter(Continent=='AFRICA'))
# summary(mod)
# 
# mod=lme4::glmer(Died_28D ~ Coma*(Creatinine_fold_change>1) + (1|STUDY), family='binomial',
#                 data= sm_data_analysis %>% filter(Continent=='AFRICA'))
# summary(mod)
# 
# namazzi = sm_data %>% filter(STUDY=='Namazzi 2022')
# mod=glm(Died_28D ~ Coma*(BUN>10), family='binomial', data= namazzi)
# summary(mod)
# 
# mod=glm(Died_28D ~ Coma*Creatinine_umol_L, family='binomial', data= namazzi)
# summary(mod)
```



## BUN/Creatinine vs Potassium 

```{r bun_Cr_potassium, fig.width=10, fig.height=8}
xx = sm_data_analysis %>% filter(Potassium_mmol_L<=10, !is.na(Potassium_mmol_L)) %>%
  mutate(Hyperkalaemia=Potassium_mmol_L>6.5,
         Uraemia = BUN>10,
         Acidotic = ifelse(!is.na(Bicarbonate_mEq_L), Bicarbonate_mEq_L<10, `Base Excess_mmol_L`< -15))
table(xx$Acidotic)

table(xx$Age_category[xx$Uraemia & xx$Acidotic & xx$Hyperkalaemia])
table(xx$Age_category[!xx$Uraemia & xx$Acidotic & xx$Hyperkalaemia])
table(xx$Age_category[xx$Uraemia & !xx$Acidotic & xx$Hyperkalaemia])

table(Hyperkalaemia=xx$Hyperkalaemia, xx$Age<15)
aggregate(Hyperkalaemia ~ Age_category, data = xx, FUN = mean)

hyperkalaemia_dat = xx %>% filter(Potassium_mmol_L>6.5, Age<15)
table(hyperkalaemia_dat$Died_28D)
p0 = xx %>%
  ggplot(aes(x=Age, y = Potassium_mmol_L, colour = Age_Categories3))+
  geom_point(alpha=0.3, shape=4)+
  ylab('Potassium (mmol/L)')+xlab('Age (years)')+
  scale_x_sqrt()+
  theme_bw(base_size = 16)+geom_hline(yintercept = c(3.5, 6.5),color='black')+
  geom_smooth(se = F, aes(group = NA, colour = 'black'))+
  scale_colour_manual(
    name = "Age (years)",
    values = c("<5" = "orange", "[5,15]" = "darkred", ">15" = 'purple') 
  )+theme(legend.position = 'inside',legend.position.inside=c(0.85,.85),
          legend.text = element_text(size = 5),
          legend.title = element_text(size = 6))

p1= xx %>%
  ggplot(aes(x=BUN, y = Potassium_mmol_L, colour = Age_Categories3))+
  geom_point(alpha=0.3, shape=4)+
  ylab('Potassium (mmol/L)')+xlab('BUN (mmol/L)')+
  scale_x_log10(breaks=breaks, minor_breaks=minor_breaks)+
  theme_bw(base_size = 16)+geom_hline(yintercept = c(3.5, 6.5),color='black')+
  geom_smooth(se = F)+
  scale_colour_manual(
    name = "Age (years)",
    values = c("<5" = "orange", "[5,15]" = "darkred", ">15" = 'purple'))

p2=xx %>% 
  ggplot(aes(x=2^Creatinine_fold_change_Pottel_CG, y = Potassium_mmol_L, colour = Age_Categories3))+
  geom_point(alpha=0.3, shape=4)+geom_vline(xintercept = 3)+
  ylab('Potassium (mmol/L)')+xlab('Creatinine fold change')+
  scale_x_log10(breaks=breaks, minor_breaks=minor_breaks)+
  theme_bw(base_size = 16)+
  geom_hline(yintercept = c(3.5, 6.5),color='black')+
  geom_smooth(se = F)+
  scale_colour_manual(
    name = "Age (years)",
    values = c("<5" = "orange", "[5,15]" = "darkred", ">15" = 'purple') 
  )

p3= xx %>% filter(Bicarbonate_mEq_L<32) %>%
  ggplot(aes(x=Bicarbonate_mEq_L, y = Potassium_mmol_L, colour = Age_Categories3))+
  geom_point(alpha=0.3, shape=4)+
  xlab('Bicarbonate (mEq/L)')+ylab('Potassium (mmol/L)')+
  theme_bw(base_size = 16)+geom_hline(yintercept = c(3.5, 6.5),color='black')+
  geom_smooth(se = F)+
  scale_colour_manual(
    name = "Age (years)",
    values = c("<5" = "orange", "[5,15]" = "darkred", ">15" = 'purple') 
  )

p4= xx %>%
  ggplot(aes(x=Lactate, y = Potassium_mmol_L, colour = Age_Categories3))+
  geom_point(alpha=0.3, shape=4)+
  xlab('Lactate (mmol/L)')+ylab('Potassium (mmol/L)')+
  theme_bw(base_size = 16)+geom_hline(yintercept = c(3.5, 6.5),color='black')+
  geom_smooth(se = F)+
  scale_colour_manual(
    name = "Age (years)",
    values = c("<5" = "orange", "[5,15]" = "darkred", ">15" = 'purple') 
  )


xx_complete = xx %>% filter(!is.na(Acidotic), !is.na(Uraemia)) 
tab <- xx_complete %>%
  group_by(Acidotic, Uraemia) %>%                 
  summarise(
    n     = n(),                  
    propz = mean(Hyperkalaemia, na.rm = TRUE),
    .groups = "drop"
  )

p5=ggplot(tab, aes(x = Acidotic, y = Uraemia, fill = propz)) +
  geom_tile(color = "grey80") +
  geom_text(
    aes(label = sprintf("%1.1f%%\n(n = %d)", 100 * propz, n)),
    size = 3
  ) +
  scale_fill_gradient(
    name   = "Hyperkalaemia",
    low    = "white",
    high   = "red",
    limits = c(0, .2),     
    labels = percent_format(accuracy = 1)
  ) +
  coord_equal() +
  labs(x = "Bicarbonate < 10 mEq/L", y = "BUN>10 mmol/L") +
  theme_bw(base_size = 16)


grid.arrange(p0+theme(legend.position = 'none'),
             p1+theme(legend.position = 'none'),
             p2+theme(legend.position = 'none'),
             p3+theme(legend.position = 'none'),
             p4+theme(legend.position = 'none'),
             p5)

mod_hyperkalaemia1 = lme4::glmer(Hyperkalaemia ~ Age_Categories3+(Creatinine_fold_change_Pottel_CG>log2(3))+(Bicarbonate_mEq_L<10)+(1|SITEID), family = 'binomial', data = xx %>% filter(Age<15))
summary(mod_hyperkalaemia1)

mod_hyperkalaemia2 = lme4::glmer(Hyperkalaemia ~ Age_Categories3+(BUN>10)+(Bicarbonate_mEq_L<10)+(1|SITEID), family = 'binomial', data = xx %>% filter(Age<15))
summary(mod_hyperkalaemia2)

mod_hyperkalaemia_death = lme4::glmer(Died_28D ~ Hyperkalaemia+ (Creatinine_fold_change_Pottel_CG>log2(3))+(Age<2)+(Bicarbonate_imputed<10)+(1|STUDY/SITE_STUDY), family = 'binomial', data = xx %>% filter(Age<15, Bicarbonate_imputed<30))
summary(mod_hyperkalaemia_death)

grid.arrange(p0, p1,p2)

aggregate(Hyperkalaemia ~ Age_category + (BUN>10) + (Bicarbonate_imputed<10), data = xx, function(x) round(100*mean(x),1))


p0= xx %>%
  ggplot(aes(x=Age, y = Potassium_mmol_L))+
  geom_point(alpha=0.3, shape=4)+
  ylab('Potassium (mmol/L)')+xlab('Age (years)')+
  scale_x_sqrt()+
  theme_bw(base_size = 16)+geom_hline(yintercept = c(3.5, 6.5),color='red')+
  geom_smooth(se = F)
p1= xx %>%
  ggplot(aes(x=BUN, y = Potassium_mmol_L))+
  geom_point(alpha=0.3, shape=4)+
  ylab('Potassium (mmol/L)')+xlab('BUN (mmol/L)')+
  scale_x_log10(breaks=breaks, minor_breaks=minor_breaks)+
  theme_bw(base_size = 18)+geom_hline(yintercept = c(3.5, 6.5),color='red')+
  geom_smooth(se = F)
p2= xx %>%
  ggplot(aes(x=Bicarbonate_imputed, y = Potassium_mmol_L))+
  geom_point(alpha=0.3, shape=4)+
  xlab('Bicarbonate (mEq/L)')+ylab('Potassium (mmol/L)')+
  theme_bw(base_size = 18)+geom_hline(yintercept = c(3.5, 6.5),color='red')+
  geom_smooth(se = F)
p3= xx %>%
  ggplot(aes(x=Lactate_imputed, y = Potassium_mmol_L))+
  geom_point(alpha=0.3, shape=4)+
  xlab('Lactate (mmol/L)')+ylab('Potassium (mmol/L)')+
  theme_bw(base_size = 18)+geom_hline(yintercept = c(3.5, 6.5),color='red')+
  geom_smooth(se = F)
grid.arrange(p0,p1,p2,p3)
```



# Age based thresholds

```{r age_based_thresholds}

# Midpoints and labels for each plateau
labels_df <- data.frame(
  Age       = c(1, 3, 8, 13, 20),
  Creatinine_umol_L = c(60, 80, 100, 130, 180),
  label     = c("60", "80", "100","130","180")
)

sm_data_creat = sm_data_creat %>% 
  mutate(creatinine_threshold = case_when(
    Age<2 ~ 60,
    Age>=2 & Age<5 ~ 80,
    Age>=5 & Age<10 ~ 100,
    Age>=10 & Age<15 ~ 130,
    Age>=15 ~ 180
  ))
sm_data_creat %>% 
  ggplot(aes(x=Age, y = Creatinine_umol_L, 
             colour = Creatinine_fold_change_Pottel_CG>log2(3)))+
  scale_colour_manual(values = c("FALSE" = "steelblue", "TRUE" = "firebrick")) +
  geom_point(alpha=0.35)+
  geom_vline(xintercept = c(2, 5, 10, 15), linetype = "dashed",
             color = "gray50", linewidth = 0.6) + 
  geom_step(color = "#185FA5", linewidth = 1, 
            aes(x=Age, y = creatinine_threshold)) +
  scale_y_log10()+scale_x_sqrt()+theme_minimal()+
  xlab('Age (years)')+ylab('Creatinine (umol/L)')+theme(legend.position = 'none')+
  geom_label(
    data = labels_df,
    aes(label = label),
    color = "#185FA5", fill = "white",
    label.size = 0.3, size = 3.5, fontface = "bold",
    vjust = -0.4)+   
  annotate("text", x = c(1, 3, 8, 13, 20), y = 1400,
           label = c("<2", "2-5", "5-10", "10-15",">15"), size=5)
```



