---
title: "Renal Analysis Severe Malaria"
author: "James Watson"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = T, cache.comments = FALSE,
                      echo = F, include = TRUE, 
                      fig.width = 8, fig.height = 8,
                      fig.pos = 'H',dev = 'png', dpi = 300)

library(brms)
library(tidybayes)
library(tidyverse)
library(grid)
library(gridExtra)
library(rstanarm)
library(scales)
```



Plotting params

```{r}
source('setup_parameters.R')
```


## Load analysis data

Analysis dataset
```{r}
load('RData/Analysis_dataset.RData')
```


# Calculate expected Cr based on age, weight

## Baseline creatinines

We look at the data from the TRACT study, Malawi site where we have follow-up creatinines and BUN in patients who survived and were not lost to follow-up. We compare these values to the expected values from the Pottel equation.

```{r baseline_creatinine_BUN}
baseline_data = sm_data %>% filter(!is.na(Baseline_Creatinine) | !is.na(`Baseline_Urea Nitrogen`)) 
p1=sm_data %>% filter(!is.na(Baseline_Creatinine)) %>%
  ggplot(aes(x=Weight_kg, y = Baseline_Creatinine))+geom_point()+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks)+theme_bw()+
  xlab('Weight (kg)')+ylab('Normal creatinine (umol/L)')+geom_smooth(method=MASS::rlm)
summary(lm(Baseline_Creatinine~Weight_kg, data = sm_data))
eGFR=120 # Using Pottel equation
p2=sm_data %>% filter(!is.na(Baseline_Creatinine)) %>%
  ggplot(aes(x=Age, y = Baseline_Creatinine))+geom_point()+
  geom_abline(slope = (107.3/eGFR)*0.0270*88.4,
              intercept = (107.3/eGFR)*0.2329*88.4)+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks)+theme_bw()+
  xlab('Age (years)')+ylab('Normal creatinine (umol/L)')+geom_smooth(method=MASS::rlm)
summary(lm(Baseline_Creatinine~Age, data = sm_data))
p3=sm_data %>% filter(!is.na(Baseline_Creatinine)) %>%
  ggplot(aes(x=Height_cm, y = Baseline_Creatinine))+geom_point()+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks)+theme_bw()+
  xlab('Height (cm)')+ylab('Normal creatinine (umol/L)')+geom_smooth(method=MASS::rlm)
summary(lm(Baseline_Creatinine~Height_cm, data = sm_data))
p4=sm_data %>% filter(!is.na(Baseline_Creatinine)) %>%
  ggplot(aes(x=wfaz, y = Baseline_Creatinine))+geom_point()+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks)+theme_bw()+
  xlab('Weight-for-age z-score')+ylab('Normal creatinine (umol/L)')+geom_smooth(method=MASS::rlm)
summary(lm(Baseline_Creatinine~wfaz, data = sm_data))
grid.arrange(p1,p2,p3,p4)

p1=sm_data %>% filter(!is.na(`Baseline_Urea Nitrogen`)) %>%
  ggplot(aes(x=Weight_kg, y = `Baseline_Urea Nitrogen`, colour = Age_category))+geom_point()+scale_x_sqrt()+
  scale_y_log10()+theme_bw()+
  xlab('Weight (kg)')+ylab('Normal BUN (mmol/L)')

p2=sm_data %>% filter(!is.na(`Baseline_Urea Nitrogen`)) %>%
  ggplot(aes(x=Age, y = `Baseline_Urea Nitrogen`))+geom_point()+
  scale_y_log10()+theme_bw()+scale_x_sqrt()+
  xlab('Age (years)')+ylab('Normal BUN (mmol/L)')

grid.arrange(p1,p2)

sm_data %>% filter(!is.na(`Baseline_Urea Nitrogen`)) %>%
  ggplot(aes(x=Baseline_Creatinine, y = `Baseline_Urea Nitrogen`))+geom_point()+
  scale_y_log10()+
  scale_x_log10()+
  theme_bw()+
  xlab('Normal Creatinine (umol/L)')+ylab('Normal BUN (mmol/L)')

sm_data %>% filter(!is.na(`Baseline_Urea Nitrogen`)) %>%
  ggplot(aes(x=Baseline_Creatinine / `Baseline_Urea Nitrogen`))+geom_histogram()+
  scale_x_log10()+
  theme_bw()+
  xlab('Normal Creatinine (umol/L)')+ylab('Normal BUN (mmol/L)')


xx = sm_data %>% filter(!is.na(Baseline_Creatinine) | !is.na(`Baseline_Urea Nitrogen`))
xx$MUAC = xx$`Mid-Upper Arm Circumference_cm`
xx$Baseline_BUN = xx$`Baseline_Urea Nitrogen`
mod_creat_baseline=stan_lm(Baseline_Creatinine ~ Age+Height_cm+Weight_kg,
                           data = xx, prior = R2(0.5))
plot(mod_creat_baseline,pars = c('Age','Height_cm','Weight_kg'))
summary(mod_creat_baseline)

mod_bun_baseline=stan_lm(Baseline_BUN ~ scale(Age) + scale(Weight_kg)+scale(Height_cm)+SEX+scale(MUAC),
                         data = xx, prior = R2(0.5))
plot(mod_bun_baseline,pars = c('scale(Age)','scale(Weight_kg)','scale(Height_cm)','SEXM','scale(MUAC)'))
summary(mod_bun_baseline)
```

Pottel age-based equation, where eGFR=107.3/(SCr/Q)
assuming a normal GFR of 120 mL/min per 1.73 m2 (height-independent). 
Q=0.0270*age+0.2329.

We estimate the Pottel baseline creatinine (age only for children)
We estimate the adult baseline creatine using Cockcroft-Gault equation
We also estimate across the age spectrum using the Full Age equation (Pottel 2016)


```{r}
sm_data_analysis = sm_data_analysis %>%
  mutate(
    Height_m = Height_imputed/100,
    # Height_m = ifelse(Height_m<0.8, 0.8, Height_m),
    Q2=3.94-13.4*Height_m+17.6*Height_m^2-9.84*Height_m^3+2.04*Height_m^4,
    bSrc_adults = 88.4*((140-Age)*Weight_kg)/(120*72),
    Baseline_Cr_Age = ifelse(Age<15, (107.3/120)*( 0.0270*Age + 0.2329)*88.4, bSrc_adults),
    Baseline_Cr_Height = ifelse(Age<15, (107.3/120)*Q2*88.4, bSrc_adults))

sm_data_analysis %>% ggplot(aes(x=Age, y = Baseline_Cr_Age))+geom_point()+
  geom_point(aes(x=Age, y=Baseline_Cr_Height, colour = 'red',alpha=.5))+
  geom_smooth()+theme_bw()

# # Make the Fold Change values for each patients
sm_data_analysis = sm_data_analysis %>%
  mutate(Creatinine_fold_change = log2(Creatinine_umol_L/Baseline_Cr_Age),
         Creatinine_fold_change2 = log2(Creatinine_umol_L/Baseline_Cr_Height))

p1=sm_data_analysis %>% ggplot(aes(x=Age, y=Baseline_Cr_Age))+geom_point()+
  theme_bw()+ylim(0,120)+ylab('Age predicted baseline Creatinine (umol/L)')
p2=sm_data_analysis %>% ggplot(aes(x=Age, y=Baseline_Cr_Height))+geom_point()+
  theme_bw()+ylim(0,120)+ylab('Height predicted baseline Creatinine (umol/L)')

grid.arrange(p1,p2)
```


TRACT predicted baseline creatinine for children

```{r tract_baseline_Cr}
p1=sm_data_analysis %>% filter(!is.na(Baseline_Creatinine)) %>%
  ggplot(aes(x=Baseline_Cr_Age, Baseline_Creatinine))+geom_point()+
  theme_bw()+geom_abline(intercept = 0, slope = 1)

p2=sm_data_analysis %>% filter(!is.na(Baseline_Creatinine)) %>%
  ggplot(aes(x=Baseline_Cr_Height, Baseline_Creatinine))+geom_point()+
  theme_bw()+geom_abline(intercept = 0, slope = 1)
grid.arrange(p1,p2)
```


is there heterogeneity by age/weight?

```{r creatinine_fold_change}
sm_data_creat = sm_data_analysis %>% filter(!is.na(Creatinine_umol_L))
sm_data_creat %>% ggplot(aes(x=Creatinine_fold_change, y = Creatinine_fold_change2, colour = Age_category))+
  geom_point()+theme_bw()

sm_data_analysis %>% ggplot(aes(x=Creatinine_fold_change))+geom_histogram()

p1=sm_data_creat %>% ggplot(aes(x=Age, y = Creatinine_fold_change, colour = STUDY))+
  geom_point()+theme_bw(base_size = 14)+
  geom_smooth(aes(group=NA))+ylab('Creatinine fold change (Potel or CG)')
p2=sm_data_creat %>% ggplot(aes(x=Age, y = Creatinine_fold_change2, colour = STUDY))+
  geom_point()+theme_bw(base_size = 14)+
  geom_smooth(aes(group=NA))+ylab('Creatinine fold change (Potel or CG)')

grid.arrange(p1,p2)
```


```{r age_vs_Cr_foldchange, fig.width=10, fig.height=7}
sm_data_creat %>% 
  ggplot(aes(x=Age, y = 2^Creatinine_fold_change))+
   geom_point()+ scale_y_log10()+
  geom_smooth()+theme_bw(base_size = 12)+
  ylab('Creatinine fold change (Pottel/CG)')+
  facet_wrap(~STUDY,nrow=2)

# sm_data_creat %>% 
#   ggplot(aes(x=Age, y = Creatinine_fold_change2))+
#   scale_y_log10() + geom_point()+
#   ylab('Creatinine fold change (TRACT/CG)')+
#   scale_x_sqrt()+geom_smooth()+theme_bw(base_size = 12)+facet_wrap(~STUDY,nrow=2)
```

```{r mgcv_tensor_product_creatinine}
par(las=1)
# Fit GAM with age-varying spline for fold-change creatinine
m = mgcv::gam(
  Died_28D ~ s(Age, k = 6) + 
       te(Creatinine_fold_change, Age, bs = c("cr", "cr"), k = c(10, 6))+
    Coma + s(SITE_STUDY,bs='re'),
  data   = sm_data_creat,
  family = binomial(link = "logit"),
  method = "REML",
  select = TRUE
)
mgcv::vis.gam(m, view = c("Creatinine_fold_change", "Age"),
              plot.type = "contour", color = "topo", too.far = 0.05,
              main='Predicted baseline using Age/Weight/Height')

m2 = mgcv::gam(
  Died_28D ~ s(Age, k = 6) + 
       te(Creatinine_fold_change2, Age, bs = c("cr", "cr"), k = c(10, 6))+
    Coma + s(SITE_STUDY,bs='re'),
  data   = sm_data_creat,
  family = binomial(link = "logit"),
  method = "REML",
  select = TRUE
)
mgcv::vis.gam(m2, view = c("Creatinine_fold_change", "Age"),
              plot.type = "contour", color = "topo", too.far = 0.05,
              main='Predicted baseline from TRACT')

# Fit GAM with age-varying spline for fold-change creatinine - without TRACT
m_rmTRACT = mgcv::gam(
  Died_28D ~ s(Age, k = 6) + 
       te(Creatinine_fold_change, Age, bs = c("cr", "cr"), k = c(10, 6))+
    Coma + s(SITE_STUDY,bs='re'),
  data   = sm_data_creat %>% filter(STUDY != 'TRACT'),
  family = binomial(link = "logit"),
  method = "REML",
  select = TRUE
)
mgcv::vis.gam(m_rmTRACT, view = c("Creatinine_fold_change", "Age"),
              plot.type = "contour", color = "topo", too.far = 0.05,
              main='Predicted baseline using Age/Weight/Height\nwithout TRACT')


m2_rmTRACT = mgcv::gam(
  Died_28D ~ s(Age, k = 6) + 
       te(Creatinine_fold_change2, Age, bs = c("cr", "cr"), k = c(10, 6))+
    Coma + s(SITE_STUDY,bs='re'),
  data   = sm_data_creat %>% filter(STUDY != 'TRACT'),
  family = binomial(link = "logit"),
  method = "REML",
  select = TRUE
)
mgcv::vis.gam(m2_rmTRACT, view = c("Creatinine_fold_change2", "Age"),
              plot.type = "contour", color = "topo", too.far = 0.05,
              main='Predicted baseline using Age/Weight/Height\nwithout TRACT')
```


```{r}
#tensor product models
Creat_tensor_prod = as.formula(bf(Died_28D ~ t2(Creatinine_fold_change,Age,bs = c("cr", "cr"))+Coma+(1|STUDY/SITE_STUDY)))
#smooth spline models with categorical age
Creat_fs_mod = as.formula(bf(Died_28D ~ s(Creatinine_fold_change, bs = "cr", k = 10) + s(Creatinine_fold_change,Age_category,bs = "fs", m = 1, k = 10)+Age_category+Coma+(1|STUDY/SITE_STUDY)))

f_path = 'RData/creatinine_models.RData'
RUN_STAN_MODELS=F
if(!file.exists(f_path) | RUN_STAN_MODELS){
  Creat_tensor_prod_fit = 
    brm(formula = Creat_tensor_prod,
        data = sm_data_creat ,
        family = bernoulli(), cores = 4, seed = 17,
        iter = 2000, warmup = 1000, thin = 1, refresh = 250)
  
  Creat_fs_mod_fit = 
    brm(formula = Creat_fs_mod,
        data = sm_data_creat ,
        family = bernoulli(), cores = 4, seed = 17,
        iter = 2000, warmup = 1000, thin = 1, refresh = 250)
  
  save(Creat_tensor_prod_fit, Creat_fs_mod_fit, file = f_path)
} else {
  load(f_path)
}
```


Make some test data

```{r}
k_vals=40
test_data = expand_grid(Creatinine_fold_change = seq(0,3,length.out=k_vals),
                        Age = c(1, 2, 5, 10, 30),
                        Coma=unique(sm_data_creat$Coma),
                        STUDY=unique(sm_data_creat$STUDY), 
                        SITE_STUDY = -1) %>% arrange(Creatinine_fold_change) %>%
  mutate(Creatinine_fold_change2 = Creatinine_fold_change,
         Age_category=cut(Age,breaks = c(0, 2, 5, 15, 100)),
         Age_category = recode(Age_category,
                               "(0,2]" = "≤2",
                               "(2,5]" = "(2,5]",
                               "(5,15]" = "(5,15]",
                               "(15,100]" = ">15"))

# remove non plausible combinations
# test_data = test_data %>%
#   filter(!(Age_category==">15" & Weight_category=='(0,25]'),
#          !(Age_category %in% c("(2,5]","≤2") & Weight_category=='(25,100]')) %>%
#   mutate(Age_category=factor(Age_category, levels = c("≤2","(2,5]","(5,15]",">15")))

# get model predictions
# Full model
preds = posterior_epred(Creat_tensor_prod_fit, newdata = test_data, re_formula = NA)
test_data$Creat_tensor_prod_pred = colMeans(preds)
test_data$Creat_tensor_prod_threshold = apply(preds, 2, f_threshold,mortality_threshold=mortality_threshold)

preds = posterior_epred(Creat_fs_mod_fit, newdata = test_data, re_formula = NA)
test_data$Creat_fs_pred = colMeans(preds)
test_data$Creat_fs_threshold = apply(preds, 2, f_threshold,mortality_threshold=mortality_threshold)
```


Model using Pottel equation (children) and CG (adults)

```{r creatinine_fc_baseline_mortality_Pottel}
p1c=test_data %>% mutate(Age = as.factor(Age))%>%
  ggplot(aes(x = 2^Creatinine_fold_change, y = Creat_tensor_prod_pred, 
             colour =Age,linetype = Coma,
             group = interaction(Age,Coma)))+
  geom_line()+ ylab('Predicted mortality')+
  scale_x_log10() +
  annotation_logticks(sides='b')+theme_bw(base_size = 13)+
  xlab('Creatinine fold change')+
  geom_hline(yintercept = mortality_threshold)+
  labs(colour='Age')+
  theme(legend.position = c(0.02, 0.98),  # (x, y) in [0,1] relative to plot area
        legend.justification = c(0, 1))   # align legend's bottom-left to that point


p1c_threshold=test_data %>%mutate(Age = as.factor(Age))%>%
  ggplot(aes(x = 2^Creatinine_fold_change, y = Creat_tensor_prod_threshold, 
             colour = Age,linetype = Coma,
             group = interaction(Age,Coma)))+
  geom_line()+ylab('Probability mortality > 5%')+
  scale_x_log10() + theme_bw(base_size = 13)+
  annotation_logticks(sides='b')+
  xlab('Creatinine fold change')+
  geom_hline(yintercept = .9)+geom_vline(xintercept = 3)+
  theme(legend.position = 'none')
grid.arrange(p1c, p1c_threshold)


sm_data_analysis %>% filter(!is.na(Creatinine_umol_L)) %>%
  ggplot(aes(x=Age, y = Creatinine_umol_L, colour = Creatinine_fold_change>3))+
  geom_point()+scale_y_log10(breaks = breaks, minor_breaks = minor_breaks)+
  geom_hline(yintercept = 265,aes(linetype = 'dashed'))+
  scale_x_sqrt()+theme_bw(base_size = 14)+
  geom_vline(xintercept = c(2, 5,15))
```


Combined figure with BUN for paper

```{r combined_fig, fig.width=12, fig.height=10}
load('RData/bun_figures.RData')
# combined fig 3
grid.arrange(p1c, p1+theme(legend.position = 'none'), p1c_threshold, p1_threshold)
```




Model using age as catgeorical variable

```{r creatinine_fs_mod}
p1=test_data %>%
  ggplot(aes(x = 2^Creatinine_fold_change, y = Creat_fs_pred, 
             colour = Age_category,linetype = Coma,
             group = interaction(Age_category,Coma)))+
  geom_line()+ ylab('Predicted mortality')+
  scale_x_log10() +
  annotation_logticks(sides='b')+theme_bw(base_size = 13)+
  xlab('Creatinine fold change')+
  geom_hline(yintercept = mortality_threshold)+
  labs(colour='Age category')

p1_threshold=test_data %>%
  ggplot(aes(x = 2^Creatinine_fold_change, y = Creat_fs_threshold, 
             colour = Age_category,linetype = Coma,
             group = interaction(Age_category,Coma)))+
  geom_line()+ylab('Probability mortality > 5%')+
  scale_x_log10() +
    geom_hline(yintercept = .9)+geom_vline(xintercept = 3)+
  annotation_logticks(sides='b')+theme_bw(base_size = 13)+
  xlab('Creatinine fold change')
grid.arrange(p1, p1_threshold)
```



### Compare creatinine fold change with BUN

Compare BUN versus Creatinine when both measured

```{r}
sm_both = sm_data_analysis %>% filter(!is.na(Creatinine_umol_L), !is.na(BUN))

# children under 5
mod_both_U5 = lme4::glmer(Died_28D~scale(log10(BUN))+scale(log10(Creatinine_fold_change))+Coma+(1|SITE_STUDY),
                          family='binomial', data = sm_both %>% filter(Age<=5))
summary(mod_both_U5)

# children 5-15
mod_both_5_15 = lme4::glmer(Died_28D~scale(log10(BUN))+scale(log10(Creatinine_fold_change))+Coma+(1|SITE_STUDY),
                            family='binomial', data = sm_both %>% filter(Age>5, Age<15))
summary(mod_both_5_15)

# Adults
mod_both_adults = lme4::glmer(Died_28D~scale(log10(BUN))+scale(log10(Creatinine_fold_change))+Coma+(1|SITE_STUDY),
                              family='binomial', data = sm_both %>% filter(Age>15))
summary(mod_both_adults)

### Fit interaction model
sm_both = sm_both %>% mutate(BUN_log2 = log2(BUN),
                             Creatinine_FC_log2 = log2(Creatinine_fold_change))

f_brm_mod_both = as.formula(bf(Died_28D ~ BUN_log2*Age_category +
                                 Creatinine_FC_log2*Age_category+
                                 Coma+(1|STUDY/SITE_STUDY)))
f_path = 'RData/Model_both.RData'
if(RUN_STAN_MODELS | !file.exists(f_path)){
  mod_both_stan = brm(formula = f_brm_mod_both,
                      data = sm_both,
                      family = bernoulli(), cores = 4, seed = 17,
                      iter = 2000, warmup = 1000, thin = 1, refresh = 250)
  save(mod_both_stan, file = f_path)
} else{
  load(f_path)
}
brms::mcmc_plot(mod_both_stan)
```


```{r marginal_effects_children_adults, fig.width=10, fig.height=6}
posterior <- brms::as_draws_df(mod_both_stan)
marginal_effects <- posterior %>%
  mutate(
    BUN_adults = b_BUN_log2+`b_BUN_log2:Age_category>15`,
    BUN_U2 = b_BUN_log2,
    BUN_25 = b_BUN_log2+`b_BUN_log2:Age_category25`,
    BUN_515 = b_BUN_log2+`b_BUN_log2:Age_category515`,
    Creat_adults = b_BUN_log2+`b_Age_category>15:Creatinine_FC_log2`,
    Creat_U2 = b_Creatinine_FC_log2,
    Creat_25 = b_BUN_log2+`b_Age_category25:Creatinine_FC_log2`,
    Creat_515 = b_BUN_log2+`b_Age_category515:Creatinine_FC_log2`
  ) %>%
  select(BUN_adults, BUN_U2, BUN_25, BUN_515,
         Creat_adults, Creat_U2, Creat_25, Creat_515) %>%
  pivot_longer(cols = everything(),
               names_to = c("Biomarker", "Age_category"),
               names_pattern = "(.*)_(.*)",
               values_to = "estimate")

summary_df <- marginal_effects %>%
  group_by(Biomarker, Age_category) %>%
  summarise(
    OR = median(exp(estimate)),
    lower = quantile(exp(estimate), 0.025),
    upper = quantile(exp(estimate), 0.975),
    .groups = "drop"
  ) %>%
  mutate(
    Age_category2 = 
      case_when(
        Age_category=='25' ~ '[2-5]',
        Age_category=='515'~ '(5-15]',
        Age_category=='U2' ~ '<2',
        Age_category=='adults' ~ '>15'), 
    Biomarker = case_when(
      Biomarker=='BUN' ~ 'BUN\n(absolute value)',
      Biomarker=='Creat' ~ 'Creatinine\n(Fold change from baseline)'),
    Age_category = factor(Age_category2, levels = c('<2','[2-5]','(5-15]','>15'))
  ) %>% arrange(Biomarker, Age_category)

ggplot(summary_df, aes(x = Age_category2, y = OR, ymin = lower, ymax = upper, color = Biomarker)) +
  geom_pointrange(position = position_dodge(width = 0.5)) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_y_log10()+
  coord_flip() +
  labs(
    y = "Odds ratio for death per doubling",
    x='',
    title = "Multivariable model of death",
    color = "Biomarker"
  ) +
  theme_bw(base_size = 16)
```







Plasma HRP2
```{r PfHRP2_BUN_Creatinine, fig.width=12, fig.height=9}
library(quantreg)
p1=sm_data_analysis %>% filter(!is.na(PfHRP2_ng_ml), !is.na(Creatinine_umol_L)) %>%
  ggplot(aes(x=PfHRP2_ng_ml+1, y = Creatinine_fold_change, colour = STUDY))+
  geom_jitter(width = .05,height = .01,alpha=0.5)+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks,
                labels = trans_format("log10", math_format(10^.x)))+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks,
                labels = trans_format("log10", math_format(10^.x)))+
  theme_bw(base_size = 14)+geom_smooth(se = F, aes(group=NA))+
  geom_smooth(method = "rq", method.args = list(tau = 0.9), se = FALSE, color = "black",linetype = 2)+
  xlab('Plasma PfHRP2 (ng/ml)')+ylab('Creatinine fold change')+
   theme(legend.position = c(0.02, 0.98),  # (x, y) in [0,1] relative to plot area
        legend.justification = c(0, 1))   # align legend's bottom-left to that point

p2=sm_data_analysis %>% filter(!is.na(PfHRP2_ng_ml), !is.na(BUN)) %>%
  ggplot(aes(x=PfHRP2_ng_ml+1, y = BUN, colour = STUDY))+
  geom_jitter(width = .05,height = .01,alpha=0.5)+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks,
                labels = trans_format("log10", math_format(10^.x)))+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks,
                labels = trans_format("log10", math_format(10^.x)))+
  theme_bw(base_size = 14)+geom_smooth(se = F, aes(group=NA))+
  geom_smooth(method = "rq", method.args = list(tau = 0.9), se = FALSE, color = "black",linetype = 2)+
  xlab('Plasma PfHRP2 (ng/ml)')+ylab('BUN (mmol/L)')+
   theme(legend.position = c(0.02, 0.98),  # (x, y) in [0,1] relative to plot area
        legend.justification = c(0, 1))   # align legend's bottom-left to that point

grid.arrange(p1,p2,nrow=1)

writeLines(sprintf('There are %s patients with plasma hrp2 and creatinine measured',
                   nrow( sm_data_analysis %>% filter(!is.na(PfHRP2_ng_ml), !is.na(Creatinine_umol_L)) )))

writeLines(sprintf('There are %s patients with plasma hrp2 and BUN measured',
                   nrow( sm_data_analysis %>% filter(!is.na(PfHRP2_ng_ml), !is.na(BUN)) )))
```



## Comparison with a lactate threshold of 5 mmol/L

```{r lactate}
mod_bun = lme4::glmer(Died ~ Age_category + Coma + as.numeric(BUN>10) + (1|STUDY), family = 'binomial', data = sm_data_analysis %>% filter(Age<15))
# summary(mod_bun)
fixef(mod_bun)

mod_creat = lme4::glmer(Died ~ Age_category + Coma + as.numeric(Creatinine_fold_change>3) + (1|STUDY), family = 'binomial', data = sm_data_analysis%>% filter(Age<15))
# summary(mod_creat)
fixef(mod_creat)

mod_lactate = lme4::glmer(Died ~ Age_category + Coma + as.numeric(Lactate>5) + (1|STUDY), family = 'binomial', data = sm_data_analysis%>% filter(Age<15))
# summary(mod_lactate)
fixef(mod_lactate)

mod_lactate2 = lme4::glmer(Died ~ Age_category + Coma + as.numeric(Lactate>5)*as.numeric(Hb<5) + (1|STUDY), family = 'binomial', data = sm_data_analysis%>% filter(Age<15))
summary(mod_lactate2)
fixef(mod_lactate2)

mod_both = lme4::glmer(Died ~ Age_category + Coma + as.numeric(Lactate>5) + as.numeric(BUN>10) + (1|STUDY), family = 'binomial', data = sm_data_analysis%>% filter(Age<15))
summary(mod_both)
fixef(mod_both)
```


### KEMRI temporal trends

```{r kemri_temporal_trend}
KEMRI = sm_data_analysis %>% filter(!is.na(Creatinine_umol_L), !is.na(date_admission), STUDY=='KEMRI')
KEMRI %>% ggplot(aes(x=date_admission, y = Creatinine_fold_change)) +
  geom_point()+scale_y_log10(breaks=breaks, minor_breaks=minor_breaks)+
  geom_smooth()+theme_bw()+
  xlab('Date of admission')+
  ylab('Creatinine fold change from predicted baseline')
```




### Interactions?

```{r}
mod=lme4::glmer(Died_28D ~ Coma*(BUN>10) + (1|STUDY), family='binomial',
          data= sm_data_analysis %>% filter(Continent=='AFRICA'))
summary(mod)

sm_data_creat$Creatinine_fold_change = sm_data_analysis %>% mutate(Creatine_FC = Creatinine_umol_L/Baseline_Creatinine)
mod=lme4::glmer(Died_28D ~ Coma*(Creatinine_fold_change>1.5) + (1|STUDY), family='binomial',
          data= sm_data_analysis %>% filter(Continent=='AFRICA'))
summary(mod)

namazzi = sm_data %>% filter(STUDY=='Namazzi 2022')
mod=glm(Died_28D ~ Coma*(BUN>10), family='binomial', data= namazzi)
summary(mod)

mod=glm(Died_28D ~ Coma*Creatinine_umol_L, family='binomial', data= namazzi)
summary(mod)
```