---
title: "Renal Analysis Severe Malaria"
author: "James Watson"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = T, cache.comments = FALSE,
                      echo = F, include = TRUE, 
                      fig.width = 8, fig.height = 8,
                      fig.pos = 'H',dev = 'png', dpi = 300)


library(tidygam)
library(mgcv)
library(brms)
library(conflicted)
library(tidyverse)
library(readr)
library(rtables)
library(RColorBrewer)
library(grid)
library(gridExtra)
library(naniar)
#library(worlddatr)

conflicts_prefer(dplyr::filter)
conflicts_prefer(brms::ar)
conflicts_prefer(dplyr::collapse)
conflicts_prefer(dplyr::combine)
conflicts_prefer(tidyr::extract)
conflicts_prefer(dplyr::lag)
conflicts_prefer(brms::s)
conflicts_prefer(purrr::set_names)
conflicts_prefer(brms::t2)
conflicts_prefer(dplyr::group_rows)

binomial_smooth <- function(...) {
  geom_smooth(method = "gam", method.args = list(family = "binomial"), ...)
}
```



Plotting params

```{r}
sqrt_breaks = c(1,4,9,16,25,36,50,64,81)
breaks <- 10^(-10:10)
minor_breaks <- rep(1:9, 21)*(10^rep(-10:10, each=9))

SEsmooth=F
my_alpha=0.2
bun_limits = c(1,100)
crea_limits = c(10,1000)

mortality_threshold = 0.05
```


## Load data


```{r load_data}
sm_data = read_csv('Data/Renal_Analysis_Data.csv')
sm_data = sm_data%>% arrange(STUDY) %>% 
  mutate(Died_28D = ifelse(`Died within 28 days`=='Yes',1,0))
table(sm_data$STUDY)
```

Visualise missing data

```{r missing_data}
cols_check = c("RFSTDTC","SITEID","ARM","COUNTRY","SEX","Age","Died",
               "Time_to_death","Capillary Refill Time_sec","Head Circumference_cm",
               "Heart Rate_beats_min","Height_cm","Mid-Upper Arm Circumference_cm",
               "Respiratory Rate_breaths_min",
               "Systolic Blood Pressure_mmHg","Temperature_C",
               "Weight_kg","Oxygen Saturation_%","Diastolic Blood Pressure_mmHg",
               "Coma_Final","Respiratory distress","Shock",
               "Hypoglycaemia","Seizure","Jaundice",
               "Blood in urine","Anemia","Hyperparasitaemia",
               "Anuria","GCS_tot","BCS_tot","para_ul","Malaria_RDT","PfHRP2_ng_ml",
               "Malaria_Positive","Glucose_mmol_L","Hemoglobin_g_L",
               "Lactic Acid_mmol_L","Base Excess_mmol_L","Creatinine_umol_L",
               "Hematocrit_%","Leukocytes_10_9_L","Platelets_10_9_L",
               "Urea Nitrogen_mmol_L")
vis_miss(sm_data %>% select(all_of(cols_check)),warn_large_data = F)
```



```{r}
sm_data %>% group_by(STUDY) %>% 
  summarise(median(Time_to_death, na.rm = T))

sm_data$SITEID=as.factor(sm_data$SITEID)
sm_data$COUNTRY=as.factor(sm_data$COUNTRY)
sm_data$STUDY=as.factor(sm_data$STUDY)

sm_data$Lactate = sm_data$`Lactic Acid_mmol_L`
sm_data = sm_data%>% select(-`Lactic Acid_mmol_L`)
sm_data$Lactate = ifelse(sm_data$Lactate>50,NA,sm_data$Lactate)
```

## Variables in dataset

### High creatinines in TRACT

```{r }
tract = sm_data %>% filter(STUDY=='TRACT')
table(as.character(tract$SITEID))

tract_creat_high = sm_data %>% filter(STUDY=='TRACT', !is.na(Creatinine_umol_L) ,Creatinine_umol_L>265)
table(as.character(tract_creat_high$SITEID))
ind_rm = which(apply(tract_creat_high,2,FUN = function(x) mean(is.na(x))==1))
tract_creat_high = tract_creat_high[, -ind_rm]

tract_creat_high %>% ggplot(aes(x=Creatinine_umol_L, y=BUN, colour = `Died within 28 days`))+geom_point()+ scale_y_log10()+scale_x_log10()+theme_minimal()
table(tract_creat_high$`Died within 28 days`)

sm_data %>% filter(STUDY=='TRACT') %>% ggplot(aes(x=Creatinine_umol_L, y=BUN, colour = `Died within 28 days`))+geom_point()+ scale_y_log10()+scale_x_log10()+theme_minimal()+geom_vline(xintercept = 265)
```



### Hypoglycaemia

```{r}
sm_data %>% ggplot(aes(x = Age, y = Glucose_mmol_L, colour = Hypoglycaemia))+
  geom_point()+theme_minimal()+
  facet_wrap(vars(STUDY))
```


### Hemoglobin-Hematocrit

```{r hb_hct}
source('hb_data_plotting.R')
sm_data=hb_hct_impute(sm_data = sm_data)
```



### Lactate and bicarbonate

```{r acidosis}
table(sm_data$STUDY, acidosis_measured=!is.na(sm_data$Lactate) | !is.na(sm_data$Bicarbonate_mEq_L))
sm_data %>% filter(!is.na(sm_data$Lactate), !is.na(sm_data$Bicarbonate_mEq_L)) %>%
  ggplot(aes(x = Lactate, y = Bicarbonate_mEq_L, colour = STUDY))+
  geom_point()+theme_minimal() + geom_smooth() + xlab('Lactate (mmol/L)')+
  ylab('Bicarbonate (mEq/L)')
ind = which(is.na(sm_data$Lactate) & !is.na(sm_data$Bicarbonate_mEq_L))
sum(ind)
mod_lactate_impute = lme4::lmer(Lactate ~ Bicarbonate_mEq_L + (1|STUDY), data = sm_data)
summary(mod_lactate_impute)
preds = predict(mod_lactate_impute, newdata = sm_data[ind, ],re.form=NA)
sm_data$Lactate[ind] = ifelse(preds>0, preds, 0)
hist(sm_data$Lactate, breaks = 100)
```



### Outlier detection for Creatinine and BUN

```{r}
sm_data %>%
  ggplot(aes(x=Creatinine_umol_L, colour = STUDY))+
  geom_histogram()+scale_x_log10()
```



How do the distributions of BUN and creatinine vary across studies in the children?

```{r boxplots}
sm_data %>% filter(Weight_kg<30) %>%
  ggplot(aes(x=STUDY, y=Creatinine_umol_L)) + geom_boxplot()+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
sm_data %>% filter(Weight_kg<30) %>%
  ggplot(aes(x=STUDY, y=BUN)) + geom_boxplot()+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks,limits=c(1,100))
```




```{r creatinine_v_bun, fig.width=12, fig.height=8}
p1=sm_data %>%
  ggplot(aes(x=Creatinine_umol_L, y = BUN, colour = STUDY))+
  geom_point(alpha=0.5)+geom_smooth(se = F)+
  geom_hline(yintercept = 20)+geom_vline(xintercept = 265)+
  xlab('Creatinine (umol/L)')+ylab('BUN (mmol/L)')+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks)+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks)+
  annotation_logticks() + theme_bw(base_size = 20)+
  theme(
    legend.position = c(0.97, 0.03), 
    legend.justification = c(1, 0), 
    legend.background = element_rect(fill = "white", color = "black", size = 0.5),
    legend.margin = margin(6, 6, 6, 6),  
    legend.text = element_text(size = 9),
    legend.title = element_text(size = 10))


p2=sm_data %>%
  ggplot(aes(x=(Creatinine_umol_L/BUN), colour = STUDY))+
  geom_density()+theme_bw(base_size = 20)+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks,limits=c(1,100))+
  annotation_logticks(sides='b')+
  xlab('Creatinine to BUN ratio')+ylab('')+
  scale_y_discrete(labels = NULL, breaks = NULL)+
  theme(legend.position = 'none')

grid.arrange(p1,p2,nrow=1)
```



BUN versus Creatinine to spot outliers

```{r}
sm_data %>% ggplot(aes(x=log2(Creatinine_umol_L) - log2(BUN)))+
  geom_histogram()+geom_vline(xintercept = c(0.5,6.5))+
  theme_minimal()

sm_data = sm_data %>%
  mutate(BUN_Creat_ratio = log2(Creatinine_umol_L)-log2(BUN),
         BUN_Creat_outlier = ifelse(is.na(BUN_Creat_ratio) | (BUN_Creat_ratio>0.5 & BUN_Creat_ratio<6.5), F, T))

sm_data %>% ggplot(aes(x=BUN, y = Creatinine_umol_L, color = BUN_Creat_outlier))+
  geom_point()+scale_x_log10()+scale_y_log10()+
  theme_minimal()
table(sm_data$BUN_Creat_outlier)
```


```{r weight_age, fig.width=12, fig.height=8}
p1=sm_data %>% filter(!is.na(Creatinine_umol_L)) %>%
  ggplot(aes(x=Age, y = Creatinine_umol_L, colour = STUDY))+
  geom_jitter(height = 0.05, width = .1, alpha=my_alpha)+
  ylab('Creatinine (umol/L)')+xlab('Age (years)')+
  geom_hline(yintercept = 265)+
  scale_y_continuous(transform  ='log10') + 
  scale_x_continuous(transform  ='sqrt',breaks = sqrt_breaks) +
  geom_smooth(se=SEsmooth)+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks) +
  annotation_logticks(sides='l') +
  theme_bw(base_size = 15)+
  theme(legend.position="none")

xx = sm_data %>% filter(!is.na(Creatinine_umol_L), !is.na(Weight_kg)) 
writeLines(sprintf('There are %s patients with weight and Creatinine jointly measured', nrow(xx)))
p2 = xx%>%
  ggplot(aes(x=Weight_kg, y = Creatinine_umol_L, colour = STUDY))+
  geom_jitter(height = 0.01, width = .1, alpha=my_alpha)+
  ylab('Creatinine (umol/L)')+xlab('Weight (kg)')+
  geom_hline(yintercept = 265)+
  scale_x_continuous(transform  ='sqrt',breaks = sqrt_breaks) + 
  geom_smooth(se=SEsmooth)+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks,limits=crea_limits) +
  annotation_logticks(sides='l') +
  theme_bw(base_size = 15)+
  theme(
    legend.position = c(0.97, 0.03), 
    legend.justification = c(1, 0),
    legend.background = element_rect(fill = "white", color = "black", size = 0.5),
    legend.margin = margin(6, 6, 6, 6),  
    legend.text = element_text(size = 9),
    legend.title = element_text(size = 10))

p3=sm_data %>% 
  ggplot(aes(x=Age, y = BUN, colour = STUDY))+
  geom_jitter(height = 0.05, width = .1, alpha=my_alpha)+
  ylab('BUN (mmol/L)')+xlab('Age (years)')+
  geom_hline(yintercept = 20)+
  scale_y_continuous(transform='log10') + 
  scale_x_continuous(transform='sqrt',breaks = sqrt_breaks) +
  geom_smooth(se=SEsmooth)+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks) +
  annotation_logticks(sides='l') +
  theme_bw(base_size = 20)#+theme(legend.position="none")

xx = sm_data %>% filter(!is.na(BUN), !is.na(Weight_kg)) 
writeLines(sprintf('There are %s patients with weight and BUN jointly measured', nrow(xx)))
p4=xx%>%ggplot(aes(x=Weight_kg, y = BUN, colour = STUDY))+
  geom_jitter(height = 0.03, width = .1, alpha=my_alpha)+
  ylab('BUN (mmol/L)')+xlab('Weight (kg)')+
  geom_hline(yintercept = 20)+
  scale_x_continuous(transform='sqrt',breaks = sqrt_breaks) +
  geom_smooth(se=SEsmooth)+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks,limits=bun_limits) +
  annotation_logticks(sides='l') + theme_bw(base_size = 15)+
  theme(
    legend.position = c(0.97, 0.03),        # Position the legend inside the plot (bottom-right corner)
    legend.justification = c(1, 0),       # Justify the legend to the bottom-right
    legend.background = element_rect(fill = "white", color = "black", size = 0.5),  # Optional: adds border to legend
    legend.margin = margin(6, 6, 6, 6),    # Adjust the margin around the legend if needed
    legend.text = element_text(size = 9),
    legend.title = element_text(size = 10)
  )

grid.arrange(p2,p4,nrow=1)

sm_data %>%
  filter(STUDY %in% c('AQUAMAT','SEAQUAMAT','FEAST','TRACT'), !is.na(BUN)) %>%
  ggplot(aes(x=Weight_kg, y = BUN, colour = STUDY))+
  geom_jitter(height = 0.02, width = .02, alpha=my_alpha)+
  ylab('BUN (mmol/L)')+xlab('Weight (kg)')+
  geom_hline(yintercept = 20)+
  scale_y_continuous(transform='log10') +
  scale_x_continuous(transform='sqrt',breaks = sqrt_breaks,limits=c(4,80)) + 
  geom_smooth()+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks) +
  annotation_logticks(sides='l') +
  theme_bw(base_size = 15)#+theme(legend.position="none")
```

```{r impute_age_weight}
table(Age=is.na(sm_data$Age),Weight=is.na(sm_data$Weight_kg))

mod_age = gam(Age ~ s(Weight_kg) + s(STUDY, bs='re'), data = sm_data)
ind1 = is.na(sm_data$Age) & !is.na(sm_data$Weight_kg)
sum(ind1)
sm_data$Age[ind1] = predict(mod_age,newdata = sm_data[ind1,])
```



## Make analysis dataset

```{r maka_analysis_data}
#Remove AQ Gambia from analysis data-set as unreliable creatinine values
sm_data = sm_data %>% group_by(USUBJID) %>%
  mutate(INCLUDE_CC_ANALYSIS = 
           !is.na(Weight_kg) &
           !is.na(Age) &
           !is.na(Malaria_Positive) & 
           !is.na(Died) & 
           (!is.na(Creatinine_umol_L) | !is.na(BUN)) &
           !is.na(Coma_Final) &
           Malaria_Positive &
           STUDY!='AQ Gambia')

table(sm_data$INCLUDE_CC_ANALYSIS,useNA = 'ifany')

table(sm_data$STUDY, sm_data$INCLUDE_CC_ANALYSIS)
table(sm_data$STUDY, sm_data$Malaria_Positive,useNA = 'ifany')

print(aggregate(Age ~ STUDY+INCLUDE_CC_ANALYSIS, data = sm_data, FUN = quantile))
print(aggregate(RFSTDTC ~ STUDY, data = sm_data, FUN = function(x) range(x,na.rm = T)))
print(aggregate(Died ~ STUDY, data = sm_data, FUN = function(x) round(100*mean(x),1)))
```



Analysis data

```{r analysis_data}
xx = sm_data %>% filter(INCLUDE_CC_ANALYSIS)
writeLines(sprintf('Analysis data: there are %s patients from %s studies with at least one baseline creatinine or BUN measurement. %s patients have creatinine; %s patients have BUN; %s patients have both.',
                   nrow(xx),
                   length(unique(xx$STUDYID)),
                   sum(!is.na(xx$Creatinine_umol_L)),
                   sum(!is.na(xx$BUN)),
                   sum(!is.na(xx$Creatinine_umol_L) & !is.na(xx$BUN))
))
```

reasons for exclusion

```{r}
sm_data = sm_data %>%
  mutate(Reason_exclude = case_when(
    INCLUDE_CC_ANALYSIS ~ NA,
    !INCLUDE_CC_ANALYSIS & is.na(Creatinine_umol_L) & is.na(BUN) ~ 'No BUN or Creatinine measurement',
    !INCLUDE_CC_ANALYSIS & (is.na(Malaria_Positive) | !Malaria_Positive) ~ 'No confirmed malaria',
    !INCLUDE_CC_ANALYSIS & is.na(Died) ~ 'No mortality outcome',
    !INCLUDE_CC_ANALYSIS & STUDY=='AQ Gambia' ~ 'AQ Gambia study',
    !INCLUDE_CC_ANALYSIS & is.na(Weight_kg) ~ 'No weight/age',
    !INCLUDE_CC_ANALYSIS & is.na(Coma_Final) ~ 'No coma information',
    T ~ 'weird!'
  ))
table(sm_data$Reason_exclude, sm_data$INCLUDE_CC_ANALYSIS, useNA = 'ifany')

table(sm_data$Reason_exclude)
```



Save dataset for multiple imputation

```{r}
readr::write_csv(x = sm_data, file = 'Data/Analysis_data.csv')
```








