---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file and re-knit. -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

# Defining renal failure in severe malaria

Analysis code for the individual patient data meta-analysis reevaluating the
acute kidney injury (AKI) thresholds used to define severe falciparum malaria in
children and adults.

The current World Health Organization definition of severe malaria AKI uses a
plasma or serum creatinine above 265 µmol/L or a blood urea nitrogen (BUN) above
20 mmol/L. Both thresholds were derived from studies in adults and were never
adapted for children, whose lower muscle mass gives substantially lower baseline
creatinine concentrations. This analysis pools individual patient data from 14
prospective studies of severe malaria (22,663 patients, of whom 16,441 enter the
primary analysis population) to estimate the age-specific relationship between
admission creatinine or urea and mortality by day 28, and to derive thresholds
that carry a comparable risk of death across the age range.

## Data availability

**No participant data are included in this repository.** The `Data/` directory
and all `.RData` files are excluded by `.gitignore`, so a fresh clone contains
the analysis code only.

Pseudonymised participant data used in this analysis are available for access via
the WWARN website (<https://www.iddo.org/wwarn/accessing-data>). Requests for
access will be reviewed by a data access committee to ensure that use of data
protects the interests of the participants and researchers according to the terms
of ethics approvals and the principles of equitable data sharing. Requests can be
submitted by email to <malariaDAC@iddo.org> via the data access form. WWARN is
registered with the Registry of Research Data Repositories.

Data are curated to the CDISC SDTM standard. `make_adam.R` expects one CSV per
SDTM domain (`TS`, `DM`, `DS`, `VS`, `SA`, `RS`, `LB`, `MB`) under
`Data/DATA <date>/`, named by the `f_path()` helper at the top of that script;
adjust the `folder` and `extra` arguments to match the release you have been
granted.

## What the analysis does

- Defines the primary analysis population: confirmed malaria infection on
  admission, a recorded (or imputed) age, coma status, a mortality outcome, and
  at least one admission creatinine or urea measurement.
- Estimates a predicted baseline (pre-illness) creatinine for every patient — the
  Pottel equation in children under 15 years and the Cockroft-Gault equation in
  adults, assuming a normal clearance of 120 mL/min — validated against
  post-illness measurements in 404 Malawian children from the TRACT trial.
- Fits Bayesian generalised additive mixed models (`brms`/Stan) for death by day
  28, using a tensor-product smooth between each exposure and age, adjusted for
  coma status and calendar year, with nested random intercepts for study and
  study site.
- Derives thresholds as the exposure values at which at least 90% of the
  posterior mass places expected mortality above 5%.
- Compares the prognostic value of the creatinine fold change, urea and venous
  lactate, and reports diagnostic accuracy of the candidate thresholds by age
  group.
- Estimates the proportions of patients presenting with reduced kidney function,
  cerebral malaria, severe anaemia and lactic acidosis, and their overlap.

## Repository layout

| File | Purpose |
|---|---|
| `make_adam.R` | Builds the analysis-ready dataset from the SDTM domain CSVs; applies outlier rules; writes `Data/adam_out.csv`. Sources `hb_data_imputation.R`. |
| `hb_data_imputation.R` | Imputes haemoglobin from haematocrit where only the latter was recorded. |
| `Make_Analysis_dataset.Rmd` | Applies the inclusion criteria, derives weight-for-age z-scores, runs the imputation, and saves the analysis datasets. |
| `multiple_imputation.R` | Chained-equations imputation with random forests (`mice`, `method = "rf"`, `m = 5`, `maxit = 15`); also draws the missingness heatmap. |
| `setup_parameters.R` | Shared constants: axis breaks, plotting limits, the 5% mortality threshold. Sourced by each analysis document. |
| `renal_analysis.Rmd` | Descriptive analyses; creatinine and urea by age and weight; non-parametric mortality by exposure category; potassium and acidosis. |
| `BUN_Analysis.Rmd` | Bayesian models for urea. **Run before `Creatinine_Analysis.Rmd`.** |
| `Creatinine_Analysis.Rmd` | Baseline creatinine equations; Bayesian models for absolute creatinine and creatinine fold change; study- and site-level random effects; the multivariable creatinine-versus-urea comparison; ROC and diagnostic accuracy; *Pf*HRP2 associations. |
| `Venn_Diagrams.Rmd` | Euler diagrams of overlapping severe malaria manifestations. |
| `Variance_components.R` | Posterior summaries of the nested random-effect standard deviations and median odds ratios for every fitted model. Run after steps 2 and 3, since it loads both sets of fitted models. |
| `descriptive_analysis_prep.Rmd`, `descriptive_analysis.Rmd` | Data checking and cohort description, run against `Data/adam_out.csv` directly. |
| `plot_missing_data.R`, `plots.R` | Missingness and exploratory plotting helpers. |
| `Non_Severe_Malaria.R` | Sensitivity analyses restricted by *Pf*HRP2 concentration. |
| `make_adam_lactate_clearance.R` | Serial lactate measurements (not used in the primary analysis). |

Intermediate objects are written to `RData/`, which you will need to create on a
fresh clone.

## Running the full analysis

Create the output directory, place the granted data release under `Data/`, then
run the steps in order. Each `.Rmd` is knitted (or run chunk by chunk in RStudio)
from the repository root.

```{r}
dir.create("RData", showWarnings = FALSE)
```

**1. Build the analysis dataset.** Uncomment the `source('make_adam.R')` call at
the top of `Make_Analysis_dataset.Rmd`, or run it directly, to regenerate
`Data/adam_out.csv` from the SDTM domains. Then knit:

```{r}
rmarkdown::render("Make_Analysis_dataset.Rmd")
```

This writes `RData/Analysis_dataset.RData` (the analysis population plus the full
pooled dataset), `RData/All_data_Severe_malaria.RData` and
`RData/imputed_analysis_datasets.RData` (the five completed datasets).

Set `RUN_IMPUTE_MODEL = TRUE` in that document to refit the imputation rather
than reuse a cached `RData/imputed_dataset.RData`. Expect this to take a long
time — random-forest chained equations over ~16,000 patients and 25 variables.

**2. Fit the urea models.**

```{r}
rmarkdown::render("BUN_Analysis.Rmd")
```

Writes `RData/bun_models.RData` and `RData/bun_figures.RData`. Both are loaded by
`Creatinine_Analysis.Rmd` — the model objects for the random-effect figures and
the marker comparisons, the figure objects for a combined panel — so this step
must precede step 3.

**3. Fit the creatinine models and the marker comparisons.**

```{r}
rmarkdown::render("Creatinine_Analysis.Rmd")
```

Writes `RData/creatinine_models.RData` and `RData/Model_both.RData`.

**4. Descriptive analyses and the remaining figures.**

```{r}
rmarkdown::render("renal_analysis.Rmd")
rmarkdown::render("Venn_Diagrams.Rmd")
```

**5. Random-effect summaries.**

```{r}
source("Variance_components.R")
```

### Refitting the Bayesian models

`BUN_Analysis.Rmd` and `Creatinine_Analysis.Rmd` both set `RUN_STAN_MODELS = F`
near the top. With that default the fitted objects are loaded from `RData/` if
present, and fitted from scratch if not — so on a fresh clone the models will be
refitted automatically. Set `RUN_STAN_MODELS = TRUE` to force refitting even when
a cache exists, for example after changing the data or a model formula.

Each model runs four chains of 2,000 iterations (1,000 warmup) on four cores.
Fitting all seven models takes on the order of hours. Seeds are set in the calls,
so the fits are reproducible for a given dataset and Stan version.

## Outputs

The `.Rmd` documents produce the figures and tables reported in the paper.
Figures are generated inline rather than written to file, with the exception of
the missingness and imputation diagnostics in `multiple_imputation.R`, which are
saved via `ggsave()` to hard-coded absolute paths — edit those two calls before
running it on another machine. `make_adam.R` also reads an external `.RData` file
from outside the repository for a supplementary comparison; that line can be
commented out without affecting the primary analysis.

## Dependencies

Developed under `r R.version.string`. Install the packages used across the
pipeline:

```{r}
install.packages(c(
  # data handling and plotting
  "tidyverse", "readr", "dplyr", "tidyr", "tibble", "ggplot2", "scales",
  "grid", "gridExtra", "RColorBrewer", "gtools", "conflicted",
  # modelling
  "brms", "rstan", "mgcv", "lme4", "splines", "posterior", "tidybayes",
  "broom", "broom.mixed", "quantreg", "pROC",
  # imputation
  "mice", "ranger", "missForest", "naniar", "doParallel", "foreach", "tictoc",
  # anthropometry, tables and diagrams
  "zscorer", "eulerr", "kableExtra", "rtables", "tidygam"
))
```

`brms` requires a working C++ toolchain and a Stan installation; see the
[RStan getting-started guide](https://mc-stan.org/rstan/). The `library(rstanarm)`
call in `Creatinine_Analysis.Rmd` supports commented-out exploratory code only
and can be removed.

## Citation and licence

Please cite the accompanying paper. Code in this repository is released under
CC0 1.0 Universal (see `LICENSE`); the participant data are governed separately
by the WWARN/IDDO data access terms described above.
