---
title: "Renal Analysis Severe Malaria"
author: "James Watson"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = T, cache.comments = FALSE,
                      echo = F, include = TRUE, 
                      fig.width = 8, fig.height = 8,
                      fig.pos = 'H',dev = 'png', dpi = 300)


library(tidygam)
library(mgcv)
library(brms)
library(conflicted)
library(tidyverse)
library(readr)
library(rtables)
library(RColorBrewer)
library(grid)
library(gridExtra)
library(naniar)
library(gtools)
#library(worlddatr)

conflicts_prefer(dplyr::filter)
conflicts_prefer(brms::ar)
conflicts_prefer(dplyr::collapse)
conflicts_prefer(dplyr::combine)
conflicts_prefer(tidyr::extract)
conflicts_prefer(dplyr::lag)
conflicts_prefer(brms::s)
conflicts_prefer(purrr::set_names)
conflicts_prefer(brms::t2)
conflicts_prefer(dplyr::group_rows)
conflicted::conflicts_prefer(brms::ddirichlet)
conflicted::conflicts_prefer(brms::rdirichlet)
conflicted::conflicts_prefer(mgcv::scat)

binomial_smooth <- function(...) {
  geom_smooth(method = "gam", method.args = list(family = "binomial"), ...)
}
```



Plotting params

```{r}
sqrt_breaks = c(1,4,9,16,25,36,50,64,81)
breaks <- 10^(-10:10)
minor_breaks <- rep(1:9, 21)*(10^rep(-10:10, each=9))

SEsmooth=F
my_alpha=0.2
bun_limits = c(1,100)
crea_limits = c(10,1000)

mortality_threshold = 0.05
```


## Load data

Analysis dataset
```{r}
sm_data_analysis = readr::read_csv(file = 'Data/Analysis_data.csv')
sm_data_analysis = sm_data_analysis %>% 
  mutate(Coma = ifelse(Coma_Final==1,'Yes','No')) %>%
  filter(!is.na(Died_28D), !is.na(Coma), !is.na(Weight_category))

sm_data_analysis = sm_data_analysis %>%
  mutate(
    Coma = as.factor(Coma),
    Weight_category = as.factor(Weight_category),
    COUNTRY=as.factor(COUNTRY),
    SITE_STUDY=paste(SITEID, STUDY,sep = '_'),
    SITE_STUDY = as.factor(SITE_STUDY))

cols_check = c("RFSTDTC","SITEID","ARM","COUNTRY","SEX","Age","Died",
               "Time_to_death","Capillary Refill Time_sec","Head Circumference_cm",
               "Heart Rate_beats_min","Height_cm","Mid-Upper Arm Circumference_cm",
               "Respiratory Rate_breaths_min",
               "Systolic Blood Pressure_mmHg","Temperature_C",
               "Weight_kg","Oxygen Saturation_%","Diastolic Blood Pressure_mmHg",
               "Coma_Final","Respiratory distress","Shock",
               "Hypoglycaemia","Seizure","Jaundice",
               "Blood in urine","Anemia","Hyperparasitaemia",
               "Anuria","GCS_tot","BCS_tot","para_ul","Malaria_RDT","PfHRP2_ng_ml",
               "Malaria_Positive","Glucose_mmol_L","Hb",
               "Lactate","Base Excess_mmol_L","Creatinine_umol_L",
               "HCT","Leukocytes_10_9_L","Platelets_10_9_L",
               "Urea Nitrogen_mmol_L")
vis_miss(sm_data_analysis %>% select(all_of(cols_check)),warn_large_data = F)
```


```{r BUN_Creatinine_ratio}
sm_both = sm_data_analysis %>% filter(!is.na(BUN), !is.na(Creatinine_umol_L))
sm_both %>% ggplot(aes(x=BUN, y=Creatinine_umol_L, colour = STUDY))+
  geom_point()+ scale_x_log10(breaks = breaks, minor_breaks = minor_breaks)+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks)+theme_minimal()+
  ylab('Creatinine (umol/L)')+xlab('Blood urea nitrogen (mmol/L)')

sm_both %>% ggplot(aes(x=BUN, y=Creatinine_umol_L/BUN, colour = STUDY))+
  geom_point()+ scale_x_log10(breaks = breaks, minor_breaks = minor_breaks)+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks)+theme_minimal()+
  xlab('Blood urea nitrogen (mmol/L)')+ylab('Creatinine/urea ratio')+
  geom_smooth(aes(group=NA))
```


Imputation of Creatinine and BUN from each other

```{r imputation_BUN_Creatinine}
mod_bun_impute = gam(log10(BUN) ~ s(log10(Creatinine_umol_L)) + s(SITE_STUDY, bs='re') + s(sqrt(Weight_kg)), data = sm_data_analysis)
summary(mod_bun_impute)
sm_data_analysis$BUN_pred = 10^predict(mod_bun_impute, newdata = sm_data_analysis)
sm_data_analysis$BUN_imputed = ifelse(is.na(sm_data_analysis$BUN), 
                                      sm_data_analysis$BUN_pred,
                                      sm_data_analysis$BUN)

mod_creat_impute = gam(log10(Creatinine_umol_L) ~ s(log10(BUN)) + s(SITE_STUDY, bs='re') + s(sqrt(Weight_kg)), data = sm_data_analysis)
summary(mod_creat_impute)
sm_data_analysis$Creatinine_pred = 10^predict(mod_creat_impute, newdata = sm_data_analysis)
sm_data_analysis$Creatinine_imputed = ifelse(is.na(sm_data_analysis$Creatinine_umol_L), 
                                             sm_data_analysis$Creatinine_pred,
                                             sm_data_analysis$Creatinine_umol_L)

table(is.na(sm_data_analysis$Creatinine_imputed), is.na(sm_data_analysis$BUN_imputed))
```


## Mortality models

### Bin BUN and Creatinine

```{r}
# Bin BUN into reasonable groups 
bun_cuts = c(.5, 5, 10, 20, 200)
sm_data_analysis <- sm_data_analysis %>%
  mutate(BUN_groups = cut(BUN, breaks = bun_cuts,include.lowest = T))  
levels(sm_data_analysis$BUN_groups)=c("<5","(5,10]","(10,20]",">20")
# Bin Creatinine into reaosnable groups 
creat_cuts = c(10,65, 130, 265, 2000)
sm_data_analysis <- sm_data_analysis %>%
  mutate(Creat_groups = cut(Creatinine_umol_L, breaks = creat_cuts,include.lowest = T))  
levels(sm_data_analysis$Creat_groups)=c("<65","(65,130]","(130,265]",">265")
```


### Non parametric

make analysis set with imputed values

```{r}
sm_data_analysis_imp = sm_data_analysis %>% 
  filter(!is.na(BUN_imputed), !is.na(Weight_category), !is.na(Coma),
         INCLUDE_CC_ANALYSIS)

# Bin BUN into 4 groups
bun_cuts = c(.5, 5, 10, 20, 200)
sm_data_analysis_imp <- sm_data_analysis_imp %>%
  mutate(BUN_groups = cut(BUN, breaks = bun_cuts,include.lowest = T))  
levels(sm_data_analysis_imp$BUN_groups)=c("<5","(5,10]","(10,20]",">20")


creat_cuts = c(10,65, 130, 265, 2000)
sm_data_analysis_imp <- sm_data_analysis_imp %>%
  mutate(Creat_groups = cut(Creatinine_umol_L, breaks = creat_cuts,include.lowest = T))  
levels(sm_data_analysis_imp$Creat_groups)=c("<65","(65,130]","(130,265]",">265")
```


Using only measured BUN data


```{r bun_groups_mortality}
sm_dat_bun = sm_data_analysis %>% 
  filter(!is.na(BUN), !is.na(Weight_category), 
         INCLUDE_CC_ANALYSIS) 

# Summarize by BUN quantile and Coma
summary_df <- sm_data_analysis_imp %>% filter(!is.na(BUN)) %>%
  group_by(Coma, BUN_groups, Weight_category) %>%
  summarise(
    mortality = mean(Died),
    n = n(),
    se = sqrt(mortality * (1 - mortality) / n),
    lower = 100*(mortality - 1.96 * se),
    upper = 100*(mortality + 1.96 * se),
    mortality = 100*mortality,
    .groups = "drop"
  )

# Plot
p1=summary_df %>% ggplot(aes(x = BUN_groups, y = mortality, group = Coma, colour = Coma)) +
  geom_point(aes(size = n), position = position_dodge(width = 0.4)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2,
                position = position_dodge(width = 0.4)) +
  labs(
    x = "Blood urea nitrogen (mmol/L)",
    y = "Mortality (%)",
    size = "Number of Patients",
    title = "Mortality by urea nitrogen level, coma status, and weight"
  ) + geom_hline(yintercept = 100*mortality_threshold)+
  theme_minimal()+facet_wrap(vars(Weight_category))
```



using the measured creatinine

```{r creat_groups_mortality}
sm_dat_creat = sm_data_analysis %>% 
  filter(!is.na(Creatinine_umol_L), !is.na(Weight_category),
         !is.na(Died_28D),!is.na(Coma),
         INCLUDE_CC_ANALYSIS)

# Summarize by Creatinine quantile and Coma
summary_df <- sm_data_analysis_imp %>% filter(!is.na(Creatinine_umol_L)) %>%
  group_by(Creat_groups, Coma, Weight_category) %>%
  summarise(
    mortality = mean(Died),
    deaths = sum(Died),
    n = n(),
    se = sqrt(mortality * (1 - mortality) / n),
    lower = 100*(mortality - 1.96 * se),
    upper = 100*(mortality + 1.96 * se),
    mortality = 100*mortality,
    .groups = "drop"
  )

# Plot
p2=summary_df %>% ggplot(aes(x = Creat_groups, y = mortality, group = Coma, colour = Coma)) +
  geom_point(aes(size = n), position = position_dodge(width = 0.4)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2,
                position = position_dodge(width = 0.4)) +
  labs(
    x = "Creatinine (umol/L)",
    y = "Mortality (%)",
    size = "Number of Patients",
    title = "Mortality by creatinine level, coma status, and weight"
  ) + geom_hline(yintercept = 100*mortality_threshold)+
  theme_minimal()+facet_wrap(vars(Weight_category))
sum(summary_df$n)
```





```{r non_para_mortality}
ggsave(plot = grid.arrange(p1,p2), filename = 'nonpara.pdf')
```

#### Individual studies

```{r indiv_studies}
sm_dat_creat %>%
  group_by(STUDY) %>%
  mutate(STUDY = paste0(STUDY, " (n = ", n(), ")")) %>%
  ggplot(aes(x=Creatinine_umol_L, y = Died_28D))+
  geom_point()+
  scale_x_log10(minor_breaks=minor_breaks, breaks=breaks)+theme_minimal()+
  geom_hline(yintercept = mortality_threshold)+ylab('Proportion dead by day 28')+
  xlab('Creatinine (umol/L)')+ geom_vline(xintercept = 265)+
  geom_smooth(method='gam',method.args = list(family='binomial'))+
  facet_wrap(~STUDY)

sm_dat_bun %>%
  group_by(STUDY) %>%
  mutate(STUDY = paste0(STUDY, " (n = ", n(), ")")) %>%
  ggplot(aes(x=BUN, y = Died_28D))+
  geom_point()+
  scale_x_log10(minor_breaks=minor_breaks, breaks=breaks)+theme_minimal()+
  geom_hline(yintercept = mortality_threshold)+ylab('Proportion dead by day 28')+
  xlab('Urea Nitrogen (mmol/L)')+ geom_vline(xintercept = 20)+
  geom_smooth(method='gam',method.args = list(family='binomial'))+
  facet_wrap(~STUDY)
```



```{r individual_study_bun_creat_mortality}
mod1 = brm(bf(Died_28D ~ s(log10(Creatinine_umol_L),by=Weight_category,k=3)+Coma+s(Age)+(1|STUDY/SITE_STUDY)),
          data = sm_dat_creat ,
          family = bernoulli(), cores = 4, seed = 17,
          iter = 2000, warmup = 1000, thin = 1, refresh = 250)
sm_dat_creat$mod1_preds=colMeans(fitted(mod1, summary = FALSE,re_formula = NA))
sm_dat_creat %>% 
  ggplot(aes(x=Creatinine_umol_L, y = mod1_preds, colour = Weight_category))+
  geom_point()+scale_x_log10()+theme_minimal()+
  geom_hline(yintercept = mortality_threshold)+
  geom_vline(xintercept = 265)

newdata = expand_grid(Creatinine_umol_L = 10^seq(1,3,length.out=20), Age=c(2,20), Weight_category=unique(sm_dat_creat$Weight_category), Coma=unique(sm_dat_creat$Coma),STUDY = -1, SITE_STUDY = -1 )
newdata = newdata %>% filter(!(Age==20 & Weight_category=='(0,25]'),
                             !(Age==2 & Weight_category=='(25,100]'))
newdata$pred = colMeans(posterior_predict(mod1, newdata = newdata, re_formula = NA))

newdata %>%
  ggplot(aes(x = Creatinine_umol_L, y = pred, 
             colour =Weight_category,linetype = Coma,
             group = interaction(Age,Weight_category,Coma)))+
   geom_line()+
   scale_x_log10(breaks = breaks, minor_breaks = minor_breaks,limits=crea_limits) +
   annotation_logticks(sides='b')+theme_bw(base_size = 15)+xlab('Creatinine (umol/L)')+
   geom_vline(xintercept = 265)+geom_hline(yintercept = mortality_threshold)



### ************** BUN model **************
mod2 = brm(bf(Died_28D ~ s(log10(BUN),by=Weight_category,k=3)+Coma+s(Age)+(1|STUDY/SITE_STUDY)),
          data = sm_dat_bun,
          family = bernoulli(), cores = 4, seed = 17,
          iter = 2000, warmup = 1000, thin = 1, refresh = 250)
sm_dat_bun$mod2_preds=colMeans(fitted(mod2, summary = FALSE,re_formula = NA))
sm_dat_bun %>% 
  ggplot(aes(x=BUN, y = mod2_preds, colour = Weight_category))+
  geom_point()+scale_x_log10()+theme_minimal()+
  geom_hline(yintercept = mortality_threshold)+
  geom_vline(xintercept = 265)

newdata = expand_grid(BUN = 10^seq(0,2,length.out=20), Age=c(2,20), Weight_category=unique(sm_dat_creat$Weight_category), Coma=unique(sm_dat_creat$Coma),STUDY = -1, SITE_STUDY = -1 )
newdata = newdata %>% filter(!(Age==20 & Weight_category=='(0,25]'),
                             !(Age==2 & Weight_category=='(25,100]'))
newdata$pred = colMeans(posterior_predict(mod2, newdata = newdata, re_formula = NA))

newdata %>%
  ggplot(aes(x = BUN, y = pred, 
             colour =Weight_category,linetype = Coma,
             group = interaction(Age,Weight_category,Coma)))+
   geom_line()+
   scale_x_log10(breaks = breaks, minor_breaks = minor_breaks,limits=crea_limits) +
   annotation_logticks(sides='b')+theme_bw(base_size = 15)+xlab('BUN (mmol/L)')+
   geom_vline(xintercept = 265)+geom_hline(yintercept = mortality_threshold)

```


#### All data

Creatinine models of mortality

```{r mortality_model_creatinine, fig.width=12, fig.height=8}
# # Using measured data
# my_f = as.formula(Died_28D ~ s(log10(Creatinine_umol_L),by=Weight_category,k=5)+
#                     Coma+Weight_category+(1|SITE_STUDY))
# mod_creat_1 <- brm(bf(formula = my_f),
#                    data = sm_data_analysis_imp ,
#                    family = bernoulli(), cores = 4, seed = 17,
#                    iter = 2000, warmup = 1000, thin = 10, refresh = 250)
# 
# my_f = as.formula(Died_28D ~ s(log10(Creatinine_imputed),by=Weight_category,k=5)+
#                     Coma+Weight_category+(1|SITE_STUDY))
# mod_creat_2 <- brm(bf(formula = my_f),
#                    data = sm_data_analysis_imp ,
#                    family = bernoulli(), cores = 4, seed = 17,
#                    iter = 2000, warmup = 1000, thin = 10, refresh = 250)
# 
# xx = colMeans(posterior_linpred(mod_creat_1,re_formula=NA))
# sm_data_analysis_imp$mod1_creat_preds=NA
# sm_data_analysis_imp$mod1_creat_preds[!is.na(sm_data_analysis_imp$Creatinine_umol_L)]=xx
# sm_data_analysis_imp$mod2_creat_preds= colMeans(posterior_linpred(mod_creat_2,re_formula=NA))
# 
# p1=sm_data_analysis_imp %>%
#   ggplot(aes(x=Creatinine_imputed, y = mod1_creat_preds, 
#              colour = Weight_category))+geom_point()+
#   scale_x_log10(breaks=breaks,minor_breaks=minor_breaks)+
#   xlab('Creatinine (umol/L)')+
#   ylab('Linear predictor for death')+ggtitle('Measured values only')+
#   geom_hline(yintercept = gtools::logit(mortality_threshold))+
#   geom_vline(xintercept = 265)+theme_minimal()
# p2=sm_data_analysis_imp %>%
#   ggplot(aes(x=Creatinine_imputed, y = mod2_creat_preds, 
#              colour = Weight_category))+geom_point()+
#   scale_x_log10(breaks=breaks,minor_breaks=minor_breaks)+
#   xlab('Creatinine (umol/L)')+
#   ylab('Linear predictor for death')+ggtitle('Measured values only')+
#   geom_hline(yintercept = gtools::logit(mortality_threshold))+
#   geom_vline(xintercept = 265)+theme_minimal()
# grid.arrange(p1,p2)
```



BUN models of mortality

```{r mortality_model_bun, fig.width=12, fig.height=8}
# # Using measured data
# my_f = as.formula(Died_28D ~ s(log10(BUN),by=Weight_category,k=5)+
#                     Coma+Weight_category+(1|SITE_STUDY))
# mod_bun_1 <- brm(bf(formula = my_f),
#                  data = sm_data_analysis_imp ,
#                  family = bernoulli(), cores = 4, seed = 17,
#                  iter = 2000, warmup = 1000, thin = 10, refresh = 250)
# 
# my_f = as.formula(Died_28D ~ s(log10(BUN_imputed),by=Weight_category,k=5)+
#                     Coma+Weight_category+(1|SITE_STUDY))
# mod_bun_2 <- brm(bf(formula = my_f),
#                  data = sm_data_analysis_imp ,
#                  family = bernoulli(), cores = 4, seed = 17,
#                  iter = 2000, warmup = 1000, thin = 10, refresh = 250)
# 
# xx = colMeans(posterior_linpred(mod_bun_1,re_formula=NA))
# sm_data_analysis_imp$mod1_bun_preds=NA
# sm_data_analysis_imp$mod1_bun_preds[!is.na(sm_data_analysis_imp$BUN)]=xx
# sm_data_analysis_imp$mod2_bun_preds= colMeans(posterior_linpred(mod_bun_2,re_formula=NA))
# 
# # Linear predictor scale
# p1=sm_data_analysis_imp %>%
#   ggplot(aes(x=BUN_imputed, y = mod1_bun_preds, 
#              colour = Weight_category))+geom_point()+
#   scale_x_log10(breaks=breaks,minor_breaks=minor_breaks)+
#   xlab('Blood urea nitrogen (mmol/L)')+
#   ylab('Linear predictor for death')+ggtitle('Measured values only')+
#   geom_hline(yintercept = gtools::logit(mortality_threshold))+
#   geom_vline(xintercept = 20)+theme_minimal()
# p2=sm_data_analysis_imp %>%
#   ggplot(aes(x=BUN_imputed, y = mod2_bun_preds, 
#              colour = Weight_category))+geom_point()+
#   scale_x_log10(breaks=breaks,minor_breaks=minor_breaks)+
#   xlab('Blood urea nitrogen (mmol/L)')+
#   ylab('Linear predictor for death')+ggtitle('Imputed values')+
#   geom_hline(yintercept = gtools::logit(mortality_threshold))+
#   geom_vline(xintercept = 20)+theme_minimal()
# grid.arrange(p1,p2)
# 
# # Outcome scale
# p1=sm_data_analysis_imp %>%
#   ggplot(aes(x=BUN_imputed, y = inv.logit(mod1_bun_preds), 
#              colour = Weight_category))+geom_point()+
#   scale_x_log10()+xlab('Blood urea nitrogen (mmol/L)')+
#   ylab('Probability of death')+ggtitle('Measured values only')+
#   geom_hline(yintercept = mortality_threshold)+
#   geom_vline(xintercept = 20)+theme_minimal()
# p2=sm_data_analysis_imp %>%
#   ggplot(aes(x=BUN_imputed, y = inv.logit(mod2_bun_preds), 
#              colour = Weight_category))+geom_point()+
#   scale_x_log10()+xlab('Blood urea nitrogen (mmol/L)')+
#   ylab('Linear predictor for death')+ggtitle('Imputed values')+
#   geom_hline(yintercept = mortality_threshold)+
#   geom_vline(xintercept = 20)+theme_minimal()
# grid.arrange(p1,p2)
```




Using both creatinine and BUN to predict outcomes

```{r}
my_f = as.formula(Died_28D ~ s(log10(BUN_imputed),by=Weight_category,k=5)+
                    s(log10(Creatinine_imputed),by=Weight_category,k=5)+
                    Coma+Weight_category+(1|SITE_STUDY))
mod_both <- brm(bf(formula = my_f),
                data = sm_data_analysis_imp ,
                family = bernoulli(), cores = 4, seed = 17,
                iter = 2000, warmup = 1000, thin = 10, refresh = 250)


mod_both2 <- lme4::glmer(Died_28D ~ log10(BUN_imputed)+
                           log10(Creatinine_imputed)+
                           Coma+Weight_category+(1|SITE_STUDY),
                         data = sm_data_analysis_imp, family = binomial())
summary(mod_both2)

mod_both2 =
  lme4::glmer(Died_28D ~ log10(BUN)+
                log10(Creatinine_umol_L)+
                Coma+(1|SITE_STUDY),
              data = sm_data_analysis_imp %>% filter(Weight_kg<25),
              family = binomial())
summary(mod_both2)

mod_both2 <- lme4::glmer(Died_28D ~ scale(log10(BUN))+
                           scale(log10(Creatinine_umol_L))+
                           Coma+(1|SITE_STUDY),
                         data = sm_data_analysis_imp %>% filter(Weight_kg>25),
                         family = binomial())
summary(mod_both2)

# Create grid of values for log10(BUN) and log10(Creatinine)
bun_seq <- seq(min(log10(sm_data_analysis_imp$BUN_imputed), na.rm = TRUE),
               max(log10(sm_data_analysis_imp$BUN_imputed), na.rm = TRUE), 
               length.out = 100)
creatinine_seq <- seq(min(log10(sm_data_analysis_imp$Creatinine_imputed), na.rm = TRUE),
                      max(log10(sm_data_analysis_imp$Creatinine_imputed), na.rm = TRUE),
                      length.out = 100)

# If Weight_category is a factor, get levels
weight_levels <- unique(sm_data_analysis_imp$Weight_category)

# Create prediction grid
new_data <- expand.grid(
  BUN_imputed = 10^bun_seq,
  Creatinine_imputed = 10^creatinine_seq,
  Weight_category = weight_levels,
  Coma = c('Yes','No')
)

# Get fitted values (mean predicted probability of death)
new_data$predicted  = colMeans(posterior_linpred(mod_both, newdata = new_data, scale = "response",re_formula = NA))

# Plot heatmap
ggplot(new_data, aes(x = BUN_imputed, y = Creatinine_imputed, fill = predicted)) +
  geom_raster() + scale_x_log10()+scale_y_log10()+
  scale_fill_viridis_c(name = "P(Death)") +
  facet_wrap(~ Weight_category) +
  labs(x = "log10(BUN)", y = "log10(Creatinine)", 
       title = "Predicted Probability of Death") +
  theme_minimal()
```



## Half thresholds??


```{r}
sm_data_analysis = sm_data_analysis %>%
  mutate(
    STUDY_Weight = paste(STUDY, Weight_category,sep = '_'),
    Above_BUN_threshold = 
      case_when(!is.na(BUN) & Weight_category=='(0,25]'&BUN>10 ~ 'YES',
                !is.na(BUN) & Weight_category=='(0,25]'&BUN<=10 ~ 'NO',
                !is.na(BUN) & Weight_category=='(25,100]'&BUN>20 ~ 'YES',
                !is.na(BUN) & Weight_category=='(25,100]'&BUN<20 ~ 'NO'),
    Above_Creat_threshold = 
      case_when(!is.na(Creatinine_umol_L) & Weight_category=='(0,25]'&Creatinine_umol_L>(265/2) ~ 'YES',
                !is.na(Creatinine_umol_L) & Weight_category=='(0,25]'&Creatinine_umol_L<=(265/2) ~ 'NO',
                !is.na(Creatinine_umol_L) & Weight_category=='(25,100]'&Creatinine_umol_L>265 ~ 'YES',
                !is.na(Creatinine_umol_L) & Weight_category=='(25,100]'&Creatinine_umol_L<=265 ~ 'NO'))%>%
  arrange(STUDY, Weight_category)


aggregate(Died ~ Above_BUN_threshold + STUDY, data = sm_data_analysis,
          FUN = function(x) c(round(100*mean(x)), sum(x),length(x)))

lyt <- basic_table() %>%
  split_cols_by('STUDY') %>%
  split_cols_by('Weight_category') %>%
  split_rows_by('Above_BUN_threshold')%>%
  summarize_row_groups() %>%
  analyze("Died", afun = function(x) c(sum(x),mean(x)), format = "xx (xx.x%)")
build_table(lyt = lyt, df = sm_data_analysis%>%filter(!is.na(BUN)))

lyt <- basic_table() %>%
  split_cols_by('STUDY') %>%
  split_cols_by('Weight_category') %>%
  split_rows_by('Above_Creat_threshold')%>%
  summarize_row_groups() %>%
  analyze("Died", afun = function(x) c(sum(x),mean(x)), format = "xx (xx.%)")
build_table(lyt = lyt, df = sm_data_analysis%>%filter(!is.na(Creatinine_umol_L)))

```


## Better prognosticator

```{r}
# sm_data_analysis_both$Lactate=ifelse(sm_data_analysis_both$`Lactic Acid_mmol_L`>50, NA, sm_data_analysis_both$`Lactic Acid_mmol_L`)
# 
# dim(sm_data_analysis_both)
# table(sm_data_analysis_both$STUDY)
# table(is.na(sm_data_analysis_both$`Lactic Acid_mmol_L`))
# 
# sm_data_analysis_both$STUDY=as.factor(sm_data_analysis_both$STUDY)
# 
# mod_both_BUN = gam(Died ~ scale(log10(BUN)) + s(COUNTRY,bs='re') , data = sm_bun , family = 'binomial')
# 
# mod_both_crea = gam(Died ~ scale(log10(Creatinine_umol_L)) + s(COUNTRY,bs='re') , data = sm_creat, family = 'binomial')
# 
# mod_both = gam(Died ~ scale(log10(Creatinine_umol_L)) + scale(log10(BUN))+ scale(Lactate) + s(STUDY,bs='re') , data = sm_data_analysis_both, family = 'binomial')
# 
# 
# exp(coef(mod_both_BUN)[2])
# exp(coef(mod_both_crea)[2])
# exp(coef(mod_both)[c(2,3,4)])

```





