---
title: "Renal Analysis Severe Malaria"
author: "James Watson"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = T, cache.comments = FALSE,
                      echo = F, include = TRUE, 
                      fig.width = 8, fig.height = 8,
                      fig.pos = 'H',dev = 'png', dpi = 300)


library(tidygam)
library(mgcv)
library(brms)
library(conflicted)
library(tidyverse)
library(readr)
library(rtables)
library(RColorBrewer)
library(grid)
library(gridExtra)
library(naniar)
library(gtools)
library(broom)
# library(tidybayes)
#library(worlddatr)

conflicts_prefer(dplyr::filter)
conflicts_prefer(brms::ar)
conflicts_prefer(dplyr::collapse)
conflicts_prefer(dplyr::combine)
conflicts_prefer(tidyr::extract)
conflicts_prefer(dplyr::lag)
conflicts_prefer(brms::s)
conflicts_prefer(purrr::set_names)
conflicts_prefer(brms::t2)
conflicts_prefer(dplyr::group_rows)
conflicted::conflicts_prefer(brms::ddirichlet)
conflicted::conflicts_prefer(brms::rdirichlet)
conflicted::conflicts_prefer(mgcv::scat)

binomial_smooth <- function(...) {
  geom_smooth(method = "gam", method.args = list(family = "binomial"), ...)
}
```



Plotting params

```{r}
sqrt_breaks = c(1,4,9,16,25,36,50,64,81)
breaks <- 10^(-10:10)
minor_breaks <- rep(1:9, 21)*(10^rep(-10:10, each=9))

SEsmooth=F
my_alpha=0.5
bun_limits = c(2,50)
crea_limits = c(20,500)

mortality_threshold = 0.05
```


## Load data

Analysis dataset
```{r}
sm_data = readr::read_csv(file = 'Data/Analysis_data.csv')
sm_data_analysis = sm_data %>% 
  filter(INCLUDE_CC_ANALYSIS) %>%
  mutate(Coma = ifelse(Coma_Final==1,'Yes','No'),
         Weight_category = factor(Weight_category, levels = c(">30","≤30")),
         Age_category = factor(Age_category, levels = c(">15","(5,15]","(2,5]","≤2")))

sm_data_analysis = sm_data_analysis %>%
  mutate(
    Coma = as.factor(Coma),
    COUNTRY=as.factor(COUNTRY),
    SITE_STUDY=paste(SITEID, STUDY,sep = '_'),
    SITE_STUDY = as.factor(SITE_STUDY),
    BUN_ANALYSIS = !is.na(BUN),
    CREAT_ANALYSIS = !is.na(Creatinine_umol_L),
    BUN_CREAT_ANALYSIS = BUN_ANALYSIS & CREAT_ANALYSIS)

cols_check = c("SITEID","ARM","COUNTRY","SEX","Age","Died",
               "Time_to_death","Capillary Refill Time_sec","Head Circumference_cm",
               "Heart Rate_beats_min","Height_cm","Mid-Upper Arm Circumference_cm",
               "Respiratory Rate_breaths_min",
               "Systolic Blood Pressure_mmHg","Temperature_C",
               "Weight_kg","Oxygen Saturation_%","Diastolic Blood Pressure_mmHg",
               "Coma_Final","Respiratory distress","Shock",
               "Hypoglycaemia","Seizure","Jaundice",
               "Blood in urine","Anemia","Hyperparasitaemia",
               "Anuria","GCS_tot","BCS_tot","para_ul","Malaria_RDT","PfHRP2_ng_ml",
               "Malaria_Positive","Glucose_mmol_L","Hb",
               "Lactate","Base Excess_mmol_L","Creatinine_umol_L",
               "HCT","Leukocytes_10_9_L","Platelets_10_9_L",
               "Urea Nitrogen_mmol_L")
vis_miss(sm_data_analysis %>% filter(INCLUDE_CC_ANALYSIS) %>% select(all_of(cols_check)),warn_large_data = F)
```


Some descriptive data on included patients

```{r}
print(aggregate(Age ~ STUDY, data = sm_data_analysis, FUN = quantile))
# print(aggregate(RFSTDTC ~ STUDY, data = sm_data_analysis, FUN = function(x) range(x,na.rm = T)))
print(aggregate(Died ~ STUDY, data = sm_data_analysis, FUN = function(x) round(100*mean(x),1)))
```




```{r BUN_Creatinine_ratio}
sm_both = sm_data_analysis %>% filter(!is.na(BUN), !is.na(Creatinine_umol_L))
sm_both %>% ggplot(aes(x=BUN, y=Creatinine_umol_L, colour = STUDY))+
  geom_point()+ scale_x_log10(breaks = breaks, minor_breaks = minor_breaks)+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks)+theme_minimal()+
  ylab('Creatinine (umol/L)')+xlab('Blood urea nitrogen (mmol/L)')

sm_both %>% ggplot(aes(x=BUN, y=Creatinine_umol_L/BUN, colour = STUDY))+
  geom_point()+ scale_x_log10(breaks = breaks, minor_breaks = minor_breaks)+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks)+theme_minimal()+
  xlab('Blood urea nitrogen (mmol/L)')+ylab('Creatinine/urea ratio')+
  geom_smooth(aes(group=NA))


sm_both %>% 
  ggplot(aes(x=Weight_kg, y=Creatinine_umol_L/BUN, colour = STUDY))+
  geom_point()+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks,limits=c(1,100))+
  theme_minimal()+
  xlab('Weight (kg)')+ylab('Creatinine/urea ratio')+
  geom_smooth(aes(group=NA))
```


Imputation of Creatinine and BUN from each other

```{r imputation_BUN_Creatinine}
mod_bun_impute = gam(log10(BUN) ~ s(log10(Creatinine_umol_L)) + s(SITE_STUDY, bs='re') + s(sqrt(Weight_kg)),
                     data = sm_data_analysis)
summary(mod_bun_impute)
sm_data_analysis$BUN_pred = 10^predict(mod_bun_impute, newdata = sm_data_analysis)
sm_data_analysis$BUN_imputed = ifelse(is.na(sm_data_analysis$BUN), 
                                      sm_data_analysis$BUN_pred,
                                      sm_data_analysis$BUN)

mod_creat_impute = gam(log10(Creatinine_umol_L) ~ s(log10(BUN)) + s(SITE_STUDY, bs='re') + s(sqrt(Weight_kg)), data = sm_data_analysis)
summary(mod_creat_impute)
sm_data_analysis$Creatinine_pred = 10^predict(mod_creat_impute, newdata = sm_data_analysis)
sm_data_analysis$Creatinine_imputed = ifelse(is.na(sm_data_analysis$Creatinine_umol_L), 
                                             sm_data_analysis$Creatinine_pred,
                                             sm_data_analysis$Creatinine_umol_L)

table(is.na(sm_data_analysis$Creatinine_imputed), is.na(sm_data_analysis$BUN_imputed))
```


```{r creatinine_v_bun, fig.width=12, fig.height=8}
p1=sm_data_analysis %>% filter(!is.na(Creatinine_umol_L), !is.na(BUN)) %>%
  ggplot(aes(x=Creatinine_umol_L, y = BUN, colour = STUDY))+
  geom_point(alpha=0.5)+geom_smooth(se = F)+
  geom_hline(yintercept = 20)+geom_vline(xintercept = 265)+
  xlab('Creatinine (umol/L)')+ylab('BUN (mmol/L)')+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks)+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks)+
  annotation_logticks() + theme_bw(base_size = 20)+
  theme(
    legend.position = c(0.97, 0.03), 
    legend.justification = c(1, 0), 
    legend.background = element_rect(fill = "white", color = "black", size = 0.5),
    legend.margin = margin(6, 6, 6, 6),  
    legend.text = element_text(size = 9),
    legend.title = element_text(size = 10))


p2=sm_data_analysis %>% filter(!is.na(Creatinine_umol_L), !is.na(BUN)) %>%
  ggplot(aes(x=(Creatinine_umol_L/BUN), colour = STUDY))+
  geom_density()+theme_bw(base_size = 20)+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks,limits=c(1,100))+
  annotation_logticks(sides='b')+
  xlab('Creatinine to BUN ratio')+ylab('')+
  scale_y_discrete(labels = NULL, breaks = NULL)+
  theme(legend.position = 'none')

grid.arrange(p1,p2,nrow=1)

p1=sm_data_analysis %>% filter(!is.na(Creatinine_umol_L), !is.na(BUN)) %>%
  ggplot(aes(x=Creatinine_umol_L, y = BUN, colour = Continent))+
  geom_point(alpha=0.5)+geom_smooth(se = F)+
  geom_hline(yintercept = 20)+geom_vline(xintercept = 265)+
  xlab('Creatinine (umol/L)')+ylab('BUN (mmol/L)')+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks)+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks)+
  annotation_logticks() + theme_bw(base_size = 20)+
  theme(
    legend.position = c(0.97, 0.03), 
    legend.justification = c(1, 0), 
    legend.background = element_rect(fill = "white", color = "black", size = 0.5),
    legend.margin = margin(6, 6, 6, 6),  
    legend.text = element_text(size = 9),
    legend.title = element_text(size = 10))
p2=sm_data_analysis %>% filter(!is.na(Creatinine_umol_L), !is.na(BUN)) %>%
  ggplot(aes(x=(Creatinine_umol_L/BUN), colour = Continent))+
  geom_density()+theme_bw(base_size = 20)+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks,limits=c(1,100))+
  annotation_logticks(sides='b')+
  xlab('Creatinine to BUN ratio')+ylab('')+
  scale_y_discrete(labels = NULL, breaks = NULL)+
  theme(legend.position = 'none')
grid.arrange(p1,p2,nrow=1)

# Show differences by weight
p1=sm_data_analysis %>% filter(!is.na(Creatinine_umol_L), !is.na(BUN)) %>%
  ggplot(aes(x=Creatinine_umol_L, y = BUN, colour = Weight_category))+
  geom_point(alpha=0.5)+geom_smooth(se = F)+
  geom_hline(yintercept = 20)+geom_vline(xintercept = 265)+
  xlab('Creatinine (umol/L)')+ylab('BUN (mmol/L)')+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks)+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks)+
  annotation_logticks() + theme_bw(base_size = 20)+
  theme(
    legend.position = c(0.97, 0.03), 
    legend.justification = c(1, 0), 
    legend.background = element_rect(fill = "white", color = "black", size = 0.5),
    legend.margin = margin(6, 6, 6, 6),  
    legend.text = element_text(size = 9),
    legend.title = element_text(size = 10))


p2=sm_data_analysis %>% filter(!is.na(Creatinine_umol_L), !is.na(BUN)) %>%
  ggplot(aes(x=(Creatinine_umol_L/BUN), colour = Weight_category))+
  geom_density()+theme_bw(base_size = 20)+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks,limits=c(1,100))+
  annotation_logticks(sides='b')+
  xlab('Creatinine to BUN ratio')+ylab('')+
  scale_y_discrete(labels = NULL, breaks = NULL)+
  theme(legend.position = 'none')

grid.arrange(p1,p2,nrow=1)

mod_bun_weight=lme4::lmer(log10(BUN) ~ Weight_kg + (1|STUDY), data = sm_data_analysis)
mod_creat_weight=lme4::lmer(log10(Creatinine_umol_L) ~ Weight_kg + (1|STUDY), data = sm_data_analysis)
summary(mod_bun_weight)
summary(mod_creat_weight)
10^(fixef(mod_bun_weight)['Weight_kg']*10)
10^(fixef(mod_creat_weight)['Weight_kg']*10)
```


```{r weight_age, fig.width=12, fig.height=8}
writeLines(sprintf('n=%s patients from %s studies have creatinine measured',sum(!is.na(sm_data_analysis$Creatinine_umol_L)), length(unique(sm_data_analysis$STUDY[!is.na(sm_data_analysis$Creatinine_umol_L)]))))

p1 = sm_data_analysis %>%
  ggplot(aes(x=Weight_kg, y = Creatinine_umol_L, colour = STUDY))+
  geom_jitter(height = 0.01, width = .1, alpha=my_alpha)+
  ylab('Creatinine (umol/L)')+xlab('Weight (kg)')+
  geom_hline(yintercept = 265)+
  scale_x_continuous(transform  ='sqrt',breaks = sqrt_breaks) + 
  geom_smooth(se=SEsmooth)+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks,limits=crea_limits) +
  annotation_logticks(sides='l') +
  theme_bw(base_size = 15)+
  theme(
    legend.position = c(0.97, 0.03), 
    legend.justification = c(1, 0),
    legend.background = element_rect(fill = "white", color = "black", size = 0.5),
    legend.margin = margin(6, 6, 6, 6),  
    legend.text = element_text(size = 7),
    legend.title = element_text(size = 8))

writeLines(sprintf('n=%s patients from %s studies have BUN measured',sum(!is.na(sm_data_analysis$BUN)), length(unique(sm_data_analysis$STUDY[!is.na(sm_data_analysis$BUN)]))))

p2=sm_data_analysis%>%
  ggplot(aes(x=Weight_kg, y = BUN, colour = STUDY))+
  geom_jitter(height = 0.03, width = .1, alpha=my_alpha)+
  ylab('BUN (mmol/L)')+xlab('Weight (kg)')+
  geom_hline(yintercept = 20)+
  scale_x_continuous(transform='sqrt',breaks = sqrt_breaks) +
  geom_smooth(se=SEsmooth)+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks,limits=bun_limits) +
  annotation_logticks(sides='l') + theme_bw(base_size = 15)+
  theme(
    legend.position = c(0.97, 0.03),        # Position the legend inside the plot (bottom-right corner)
    legend.justification = c(1, 0),       # Justify the legend to the bottom-right
    legend.background = element_rect(fill = "white", color = "black", size = 0.5),  # Optional: adds border to legend
    legend.margin = margin(6, 6, 6, 6),    # Adjust the margin around the legend if needed
    legend.text = element_text(size = 7),
    legend.title = element_text(size = 8))

grid.arrange(p1,p2,nrow=1)
```

Plot Creatinines by study

```{r creatinine_by_study, fig.width=12, fig.height=7}
sm_data_analysis %>% filter(!is.na(Creatinine_umol_L)) %>%
  ggplot(aes(x=Weight_kg, y = Creatinine_umol_L, colour = `Died within 28 days`))+
  geom_jitter(height = 0.01, width = .1, alpha=my_alpha)+
  ylab('Creatinine (umol/L)')+xlab('Weight (kg)')+
  geom_hline(yintercept = 265)+
  geom_smooth(se=SEsmooth, aes(group=NA))+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks) +
  theme_minimal()+
  annotation_logticks(sides='l') + facet_wrap(~STUDY, nrow=2)
```


Plot BUN by study

```{r bun_by_study, fig.width=12, fig.height=7}
sm_data_analysis %>% filter(!is.na(BUN)) %>%
  ggplot(aes(x=Weight_kg, y = BUN, colour = `Died within 28 days`))+
  geom_jitter(height = 0.01, width = .1, alpha=my_alpha)+
  ylab('BUN (mmol/L)')+xlab('Weight (kg)')+
  geom_hline(yintercept = 20)+
  geom_smooth(se=SEsmooth, aes(group=NA))+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks) +
  theme_minimal()+
  annotation_logticks(sides='l') + facet_wrap(~STUDY, nrow=2)
```


## Mortality models

### Bin BUN and Creatinine

```{r}
# Bin BUN & Creatinine into reasonable groups 
bun_cuts = c(.5, 5, 10, 20, 200)
creat_cuts = c(10,65, 130, 265, 2000)
sm_data_analysis = sm_data_analysis %>%
  mutate(BUN_groups = cut(BUN, breaks = bun_cuts,include.lowest = T),
         Creat_groups = cut(Creatinine_umol_L, breaks = creat_cuts,include.lowest = T))  
levels(sm_data_analysis$BUN_groups)=c("<5","(5,10]","(10,20]",">20")
levels(sm_data_analysis$Creat_groups)=c("<65","(65,130]","(130,265]",">265")
```


### Non parametric


Using measured BUN data

```{r bun_groups_mortality}
sm_dat_bun = sm_data_analysis %>% filter(BUN_ANALYSIS) 

# Summarize by BUN quantile and Coma
summary_df <- sm_dat_bun %>% filter(!is.na(BUN)) %>%
  group_by(Coma, BUN_groups, Weight_category) %>%
  summarise(
    mortality = mean(Died),
    n = n(),
    se = sqrt(mortality * (1 - mortality) / n),
    lower = 100*(mortality - 1.96 * se),
    upper = 100*(mortality + 1.96 * se),
    mortality = 100*mortality,
    .groups = "drop"
  )

# Plot
p1=summary_df %>% ggplot(aes(x = BUN_groups, y = mortality, group = Coma, colour = Coma)) +
  geom_point(aes(size = n), position = position_dodge(width = 0.4)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2,
                position = position_dodge(width = 0.4)) +
  labs(
    x = "Blood urea nitrogen (mmol/L)",
    y = "Mortality (%)",
    size = "Number of Patients",
    title = "Mortality by urea nitrogen level, coma status, and weight"
  ) + geom_hline(yintercept = 100*mortality_threshold)+
  theme_minimal()+facet_wrap(vars(Weight_category))
```



using the measured creatinine

```{r creat_groups_mortality}
sm_dat_creat = sm_data_analysis %>% 
  filter(CREAT_ANALYSIS)

# Summarize by Creatinine quantile and Coma
summary_df <- sm_dat_creat %>% filter(!is.na(Creatinine_umol_L)) %>%
  group_by(Creat_groups, Coma, Weight_category) %>%
  summarise(
    mortality = mean(Died),
    deaths = sum(Died),
    n = n(),
    se = sqrt(mortality * (1 - mortality) / n),
    lower = 100*(mortality - 1.96 * se),
    upper = 100*(mortality + 1.96 * se),
    mortality = 100*mortality,
    .groups = "drop"
  )

# Plot
p2=summary_df %>% ggplot(aes(x = Creat_groups, y = mortality, group = Coma, colour = Coma)) +
  geom_point(aes(size = n), position = position_dodge(width = 0.4)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2,
                position = position_dodge(width = 0.4)) +
  labs(
    x = "Creatinine (umol/L)",
    y = "Mortality (%)",
    size = "Number of Patients",
    title = "Mortality by creatinine level, coma status, and weight"
  ) + geom_hline(yintercept = 100*mortality_threshold)+
  theme_minimal()+facet_wrap(vars(Weight_category))
sum(summary_df$n)
```





```{r non_para_mortality}
# ggsave(plot = grid.arrange(p1,p2), filename = 'nonpara.pdf')
grid.arrange(p1,p2)
```

#### Individual studies

```{r indiv_studies, fig.width=12, fig.height=8}
sm_dat_creat %>%
  group_by(STUDY) %>%
  mutate(STUDY = paste0(STUDY, " (n = ", n(), ")")) %>%
  ggplot(aes(x=Creatinine_umol_L, y = Died_28D))+
  geom_point()+
  scale_x_log10(minor_breaks=minor_breaks, breaks=breaks)+theme_minimal()+
  geom_hline(yintercept = mortality_threshold)+ylab('Proportion dead by day 28')+
  xlab('Creatinine (umol/L)')+ geom_vline(xintercept = 265)+
  geom_smooth(method='gam',method.args = list(family='binomial'))+
  facet_wrap(~STUDY, nrow=2)

sm_dat_bun %>%
  group_by(STUDY) %>%
  mutate(STUDY = paste0(STUDY, " (n = ", n(), ")")) %>%
  ggplot(aes(x=BUN, y = Died_28D))+
  geom_point()+
  scale_x_log10(minor_breaks=minor_breaks, breaks=breaks)+theme_minimal()+
  geom_hline(yintercept = mortality_threshold)+ylab('Proportion dead by day 28')+
  xlab('Urea Nitrogen (mmol/L)')+ geom_vline(xintercept = 20)+
  geom_smooth(method='gam',method.args = list(family='binomial'))+
  facet_wrap(~STUDY,nrow=3)

#Fits by age category
sm_dat_bun %>% filter(Coma=='No', STUDY!='TRACT')%>%
  ggplot(aes(x=BUN, y = Died_28D))+
  geom_point()+
  scale_x_log10(minor_breaks=minor_breaks, breaks=breaks)+theme_minimal()+
  geom_hline(yintercept = mortality_threshold)+ylab('Proportion dead by day 28')+
  xlab('Urea Nitrogen (mmol/L)')+ geom_vline(xintercept = 20)+
  geom_smooth(method='gam',method.args = list(family='binomial'))+
  facet_wrap(~Age_category)
```


```{r brm_formulas}
f_brm_mod_creat = as.formula(bf(Died_28D ~ s(log10(Creatinine_umol_L),by=Weight_category)+Coma+(1|STUDY/SITE_STUDY)))
f_brm_mod_bun = as.formula(bf(Died_28D ~ s(log10(BUN),by=Weight_category)+Coma+(1|STUDY/SITE_STUDY)))

f_brm_mod_creat_age = as.formula(bf(Died_28D ~ s(log10(Creatinine_umol_L),by=Age_category)+Coma+(1|STUDY/SITE_STUDY)))
f_brm_mod_bun_age = as.formula(bf(Died_28D ~ s(log10(BUN),by=Age_category)+Coma+(1|STUDY/SITE_STUDY)))

f_brm_mod_creat_study = as.formula(bf(Died_28D ~ s(log10(Creatinine_umol_L),by=STUDY)+Coma+(1|STUDY)))
f_brm_mod_bun_study = as.formula(bf(Died_28D ~ s(log10(BUN),by=STUDY)+Coma+(1|STUDY)))
```



```{r fit_full_models}
RUN_STAN_MODELS=F
f_path='RData/full_models.RData'
if(!file.exists(f_path) | RUN_STAN_MODELS){
  mod_creat_full = 
    brm(formula = f_brm_mod_creat,
        data = sm_dat_creat ,
        family = bernoulli(), cores = 4, seed = 17,
        iter = 2000, warmup = 1000, thin = 1, refresh = 250)
  
  mod_bun_full = 
    brm(formula = f_brm_mod_bun,
        data = sm_dat_bun,
        family = bernoulli(), cores = 4, seed = 17,
        iter = 2000, warmup = 1000, thin = 1, refresh = 250)
  save(mod_creat_full, mod_bun_full, file = f_path)
} else {
  load(f_path)
}
```

Sensitivity analyses creatinine and BUN

```{r sensitivity_models}
RUN_STAN_MODELS=F
f_path='RData/sensitivity_models.RData'
if(!file.exists(f_path) | RUN_STAN_MODELS){
  ## fit to data removing TRACT
  mod_creat_remove_tract = 
    brm(formula = f_brm_mod_creat,
        data = sm_dat_creat %>% filter(STUDY != 'TRACT'),
        family = bernoulli(), cores = 4, seed = 17,
        iter = 2000, warmup = 1000, thin = 1, refresh = 250)
  mod_bun_remove_tract = 
    brm(f_brm_mod_bun,
        data = sm_dat_bun %>% filter(STUDY != 'TRACT'),
        family = bernoulli(), cores = 4, seed = 17,
        iter = 2000, warmup = 1000, thin = 1, refresh = 250)
  
  ## stratify by age category
  mod_creat_age = 
    brm(formula = f_brm_mod_creat_age,
        data = sm_dat_creat,
        family = bernoulli(), cores = 4, seed = 17,
        iter = 2000, warmup = 1000, thin = 1, refresh = 250)
  mod_bun_age = 
    brm(f_brm_mod_bun_age,
        data = sm_dat_bun,
        family = bernoulli(), cores = 4, seed = 17,
        iter = 2000, warmup = 1000, thin = 1, refresh = 250)
  
  ## stratify by STUDY
  mod_creat_study = 
    brm(formula = f_brm_mod_creat_study,
        data = sm_dat_creat,
        family = bernoulli(), cores = 4, seed = 17,
        iter = 2000, warmup = 1000, thin = 1, refresh = 250)
  mod_bun_study = 
    brm(f_brm_mod_bun_study,
        data = sm_dat_bun,
        family = bernoulli(), cores = 4, seed = 17,
        iter = 2000, warmup = 1000, thin = 1, refresh = 250)
  
  ## save output
  save( mod_creat_age, mod_bun_age,mod_creat_study,mod_bun_study,
        mod_creat_remove_tract, mod_bun_remove_tract, file = f_path)
} else {
  load(f_path)
}
```

Heterogeneity of effect by age/weight?

```{r bun_and_creatinine_heterogeneity}
p1=glm(Died_28D~log10(Creatinine_umol_L)*Weight_category+Coma+STUDY,
       family='binomial', data = sm_dat_creat) %>%
  broom::tidy(conf.int = TRUE, exponentiate = F) %>%
  filter(!grepl("STUDY", term),term!='(Intercept)') %>%
  ggplot(aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_pointrange() + coord_flip() + geom_hline(yintercept = 0, linetype = "dashed") +
  labs(y = "log odds", x = NULL) + theme_minimal()


p2=glm(Died_28D~log10(Creatinine_umol_L)*Age_category+Coma+STUDY,
       family='binomial', data = sm_dat_creat) %>%
  broom::tidy(conf.int = TRUE, exponentiate = F) %>%
  filter(!grepl("STUDY", term),term!='(Intercept)') %>%
  ggplot(aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_pointrange() + coord_flip() + geom_hline(yintercept = 0, linetype = "dashed") +
  labs(y = "log odds", x = NULL) + theme_minimal()

p3=glm(Died_28D~log10(BUN)*Weight_category+Coma+STUDY,
       family='binomial', data = sm_dat_bun) %>%
  broom::tidy(conf.int = TRUE, exponentiate = F) %>%
  filter(!grepl("STUDY", term),term!='(Intercept)') %>%
  ggplot(aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_pointrange() + coord_flip() + geom_hline(yintercept = 0, linetype = "dashed") +
  labs(y = "log odds", x = NULL) + theme_minimal()


p4=glm(Died_28D~log10(BUN)*Age_category+Coma+STUDY,
       family='binomial', data = sm_dat_bun) %>%
  broom::tidy(conf.int = TRUE, exponentiate = F) %>%
  filter(!grepl("STUDY", term),term!='(Intercept)') %>%
  ggplot(aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_pointrange() + coord_flip() + geom_hline(yintercept = 0, linetype = "dashed") +
  labs(y = "log odds", x = NULL) + theme_minimal()
grid.arrange(p1,p2,p3,p4)
```


Compare BUN versus Creatinine when both measured

```{r}
table(sm_both$STUDY)
mod_both1 = lme4::glmer(Died_28D~scale(log10(BUN))+scale(log10(Creatinine_umol_L))+Coma+(1|SITE_STUDY),
                        family='binomial', data = sm_both%>%filter(Weight_kg<=30, STUDY!='TRACT'))
summary(mod_both1)
mod_both12 = lme4::glmer(Died_28D~scale(log10(BUN))+scale(log10(Creatinine_umol_L))+Coma+(1|SITE_STUDY),
                         family='binomial', data = sm_both%>%filter(Weight_kg<=30))
summary(mod_both12)

mod_both2 = lme4::glmer(Died_28D~scale(log10(BUN))+scale(log10(Creatinine_umol_L))+Coma+(1|SITE_STUDY),
                        family='binomial', data = sm_both%>%filter(Weight_kg>30, STUDY!='TRACT'))
summary(mod_both2)
mod_both22 = lme4::glmer(Died_28D~scale(log10(BUN))+scale(log10(Creatinine_umol_L))+Coma+(1|SITE_STUDY),
                         family='binomial', data = sm_both%>%filter(Weight_kg>30))
summary(mod_both22)

mod_both3 = lme4::glmer(Died_28D~scale(log10(BUN))*Weight_category+
                          scale(log10(Creatinine_umol_L))*Weight_category+Coma+(1|SITE_STUDY),
                        family='binomial', data = sm_both%>%filter(STUDY!='TRACT'))
summary(mod_both3)
mod_both32 = lme4::glmer(Died_28D~scale(log10(BUN))*Weight_category+
                           scale(log10(Creatinine_umol_L))*Weight_category+Coma+(1|SITE_STUDY),
                         family='binomial', data = sm_both)
summary(mod_both32)

mod_both_rm_TRACT = lme4::glmer(Died_28D ~ scale(log10(BUN)) + scale(log10(Creatinine_umol_L))+Coma+(1|SITEID),
                                family='binomial', data = sm_both %>% filter(INCLUDE_CC_ANALYSIS, STUDY!='TRACT'))
summary(mod_both_rm_TRACT)

sm_both = sm_both %>% mutate(BUN_log10 = log10(BUN),
                             Creatinine_log10 = log10(Creatinine_umol_L))

f_brm_mod_both = as.formula(bf(Died_28D ~ BUN_log10*Weight_category + Creatinine_log10*Weight_category+Coma+(1|STUDY/SITE_STUDY)))
mod_both_stan = brm(formula = f_brm_mod_both,
                    data = sm_both,
                    family = bernoulli(), cores = 4, seed = 17,
                    iter = 2000, warmup = 1000, thin = 1, refresh = 250)
brms::mcmc_plot(mod_both_stan)
```


```{r marginal_effects_children_adults, fig.width=10, fig.height=6}
posterior <- as_draws_df(mod_both_stan)
marginal_effects <- posterior %>%
  tidybayes::spread_draws(
    b_BUN_log10,
    `b_BUN_log10:Weight_category≤30`,
    b_Creatinine_log10,
    `b_Weight_category≤30:Creatinine_log10`
  ) %>%
  mutate(
    BUN_adults = b_BUN_log10,
    BUN_children = b_BUN_log10 + `b_BUN_log10:Weight_category≤30`,
    
    Cr_adults = b_Creatinine_log10,
    Cr_children = b_Creatinine_log10 + `b_Weight_category≤30:Creatinine_log10`
  ) %>%
  select(BUN_adults, BUN_children, Cr_adults, Cr_children) %>%
  pivot_longer(cols = everything(),
               names_to = c("Biomarker", "Weight_category"),
               names_pattern = "(.*)_(.*)",
               values_to = "estimate")

summary_df <- marginal_effects %>%
  group_by(Biomarker, Weight_category) %>%
  summarise(
    OR = median(exp(estimate)),
    lower = quantile(exp(estimate), 0.025),
    upper = quantile(exp(estimate), 0.975),
    .groups = "drop"
  )

ggplot(summary_df, aes(x = Weight_category, y = OR, ymin = lower, ymax = upper, color = Biomarker)) +
  geom_pointrange(position = position_dodge(width = 0.5)) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_y_log10()+
  coord_flip() +
  labs(
    y = "Odds Ratio for death for 10x increase",
    x='',
    title = "Multivariable model",
    color = "Biomarker"
  ) +
  theme_minimal(base_size = 16)
```

separately by study
```{r}
# Fit logistic regressions per group in z
results <- sm_both %>% filter(STUDY != "Dong Nai") %>%
  group_by(STUDY) %>%
  do(tidy(glm(Died_28D~BUN_log10+Creatinine_log10+Coma, data = ., family = binomial))) %>%
  filter(term %in% c("BUN_log10", "Creatinine_log10")) %>%  # Only keep the slope (effect of y)
  mutate(
    OR = (estimate),                    # Odds ratio
    conf.low = (estimate - 1.96 * std.error), # 95% CI lower bound
    conf.high = (estimate + 1.96 * std.error) # 95% CI upper bound
  )

results %>% ggplot(aes(x = OR, y = STUDY)) +
  geom_pointrange(aes(xmin = conf.low, xmax = conf.high)) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(
    x = "Odds ratio for death for 10x increase"
  ) +
  theme_minimal()+facet_wrap(~term)
```


Only TRACT study


```{r}
preds = colMeans(posterior_epred(mod_creat_study))
sm_dat_creat$preds_study = preds
p1=sm_dat_creat %>% filter(Coma=='No') %>%
  arrange(STUDY, Creatinine_umol_L)%>%
  ggplot(aes(x=Creatinine_umol_L, y=preds_study, colour = STUDY,linetype = Continent))+
  geom_line()+theme_minimal()+
  ylab('Predicted mortality')+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks,limits=crea_limits) +
  annotation_logticks(sides='b')+theme_bw(base_size = 15)+xlab('Creatinine (umol/L)')+
  geom_vline(xintercept = 265)+geom_hline(yintercept = mortality_threshold)

preds = colMeans(posterior_epred(mod_bun_study))
sm_dat_bun$preds_study = preds
p2=sm_dat_bun %>% filter(Coma=='No') %>%
  arrange(STUDY, BUN)%>%
  ggplot(aes(x=BUN, y=preds_study, colour = STUDY,linetype = Continent))+
  geom_line()+theme_minimal()+
  ylab('Predicted mortality')+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks,limits=bun_limits) +
  annotation_logticks(sides='b')+theme_bw(base_size = 15)+xlab('BUN (mmol/L)')+
  geom_vline(xintercept = 20)+geom_hline(yintercept = mortality_threshold)
grid.arrange(p1,p2)
```


make a test dataset on which to evaluate model predictions

```{r test_data}
k_vals=40
test_data = expand_grid(Creatinine_umol_L = 10^seq(1,3,length.out=k_vals),
                        BUN = 10^seq(0,2,length.out=k_vals),
                        Age_category=c(">15","(5,15]","(2,5]","≤2"), 
                        Weight_category=unique(sm_dat_creat$Weight_category),
                        Coma=unique(sm_dat_creat$Coma),
                        STUDY = -1, SITE_STUDY = -1) %>% arrange(Creatinine_umol_L)
# test_data$BUN = sort(rep(10^seq(0,2,length.out=k_vals), 16))

bun_qn_adults = (quantile(sm_data_analysis$BUN[sm_data_analysis$Weight_category=='>30'], na.rm = T, probs = c(0.025,.975)))
bun_qn_child = (quantile(sm_data_analysis$BUN[sm_data_analysis$Weight_category=='≤30'], na.rm = T, probs = c(0.025,.975)))

creat_qn_adults = (quantile(sm_data_analysis$Creatinine_umol_L[sm_data_analysis$Weight_category=='>30'], na.rm = T, probs = c(0.025,.975)))
creat_qn_child = (quantile(sm_data_analysis$Creatinine_umol_L[sm_data_analysis$Weight_category=='≤30'], na.rm = T, probs = c(0.025,.975)))

# remove non plausible combinations
test_data = test_data %>%
  filter(!(BUN<bun_qn_adults[1] & Weight_category=='>30'), 
         !(BUN>bun_qn_adults[2] & Weight_category=='>30'),
         !(BUN<bun_qn_child[1] & Weight_category=='≤30'), 
         !(BUN>bun_qn_child[2] & Weight_category=='≤30'),
         !(Creatinine_umol_L<creat_qn_adults[1] & Weight_category=='>30'), 
         !(Creatinine_umol_L>creat_qn_adults[2] & Weight_category=='>30'),
         !(Creatinine_umol_L<creat_qn_child[1] & Weight_category=='≤30'), 
         !(Creatinine_umol_L>creat_qn_child[2] & Weight_category=='≤30'),
         !(Age_category==">15" & Weight_category=='(0,25]'),
         !(Age_category %in% c("(2,5]","≤2") & Weight_category=='(25,100]')) %>%
  mutate(Age_category=factor(Age_category, levels = c("≤2","(2,5]","(5,15]",">15")))

# get model predictions
f_threshold = function(x, mortality_threshold) mean(x>mortality_threshold) 
# Full model
preds = posterior_epred(mod_creat_full, newdata = test_data, re_formula = NA)
test_data$pred_mod_creat_full = colMeans(preds)
test_data$creat_full_threshold = apply(preds, 2, f_threshold,mortality_threshold=mortality_threshold)

preds = posterior_epred(mod_bun_full, newdata = test_data, re_formula = NA)
test_data$pred_mod_bun_full = colMeans(preds)
test_data$bun_full_threshold = apply(preds, 2, f_threshold,mortality_threshold=mortality_threshold)

# Remove TRACT
preds = posterior_epred(mod_creat_remove_tract, newdata = test_data, re_formula = NA)
test_data$pred_mod_creat_remove_tract = colMeans(preds)
test_data$creat_full_threshold_tractrm = apply(preds, 2, f_threshold,mortality_threshold=mortality_threshold)

preds=posterior_epred(mod_bun_remove_tract, newdata = test_data, re_formula = NA)
test_data$pred_mod_bun_remove_tract = colMeans(preds)
test_data$bun_full_threshold_tractrm = apply(preds, 2, f_threshold,mortality_threshold=mortality_threshold)

# Age stratification model
preds = posterior_epred(mod_bun_age, newdata = test_data, re_formula = NA)
test_data$pred_mod_bun_age = colMeans(preds)
test_data$bun_age_threshold = apply(preds, 2, f_threshold,mortality_threshold=mortality_threshold)

preds = posterior_epred(mod_creat_age, newdata = test_data, re_formula = NA)
test_data$pred_mod_creat_age = colMeans(preds)
test_data$creat_age_threshold = apply(preds, 2, f_threshold,mortality_threshold=mortality_threshold)
```

Plot model predictions

```{r creat_bun_full_model}
p1=test_data %>%
  ggplot(aes(x = Creatinine_umol_L, y = pred_mod_creat_full, 
             colour =Weight_category,linetype = Coma,
             group = interaction(Weight_category,Coma)))+
  geom_line()+ ylab('Predicted mortality')+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks,limits=crea_limits) +
  annotation_logticks(sides='b')+theme_bw(base_size = 13)+xlab('Creatinine (umol/L)')+
  geom_vline(xintercept = 265)+geom_hline(yintercept = mortality_threshold)+
  theme(legend.position = 'none')
p1_threshold=test_data %>%
  ggplot(aes(x = Creatinine_umol_L, y = creat_full_threshold, 
             colour = Weight_category,linetype = Coma,
             group = interaction(Weight_category,Coma)))+
  geom_line()+ylab('Probability mortality > 5%')+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks,limits=crea_limits) +
  annotation_logticks(sides='b')+theme_bw(base_size = 13)+xlab('Creatinine (umol/L)')+
  geom_vline(xintercept = 265)+geom_hline(yintercept = mortality_threshold)+
  theme(legend.position = 'none')

p2=test_data %>%
  ggplot(aes(x = BUN, y = pred_mod_bun_full, 
             colour =Weight_category,linetype = Coma,
             group = interaction(Weight_category,Coma)))+
  geom_line()+ylab('Predicted mortality')+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks,limits=bun_limits) +
  annotation_logticks(sides='b')+theme_bw(base_size = 13)+xlab('BUN (mmol/L)')+
  geom_vline(xintercept = 20)+geom_hline(yintercept = mortality_threshold)+
  theme(legend.position = 'none')
p2_threshold=test_data %>%
  ggplot(aes(x = BUN, y = bun_full_threshold, 
             colour = Weight_category, linetype = Coma,
             group = interaction(Weight_category,Coma)))+
  geom_line()+ylab('Probability mortality > 5%')+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks,limits=bun_limits) +
  annotation_logticks(sides='b')+theme_bw(base_size = 13)+xlab('BUN (mmol/L)')+
  geom_vline(xintercept = 20)+geom_hline(yintercept = mortality_threshold)+
  theme(legend.position = 'none')
```


```{r Remove_TRACT_creat_bun_model}
p3=test_data %>%
  ggplot(aes(x = Creatinine_umol_L, y = pred_mod_creat_remove_tract, 
             colour =Weight_category,linetype = Coma,
             group = interaction(Weight_category,Coma)))+
  geom_line()+ylab('Predicted mortality')+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks,limits=crea_limits) +
  annotation_logticks(sides='b')+theme_bw(base_size = 15)+xlab('Creatinine (umol/L)')+
  geom_vline(xintercept = 265)+geom_hline(yintercept = mortality_threshold)+
  ggtitle('Creatinine - without TRACT')+theme(legend.position = 'none')
p3_threshold=test_data %>%
  ggplot(aes(x = Creatinine_umol_L, y = creat_full_threshold_tractrm, 
             colour = Weight_category,linetype = Coma,
             group = interaction(Weight_category,Coma)))+
  geom_line()+ylab('Probability mortality > 5%')+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks,limits=crea_limits) +
  annotation_logticks(sides='b')+theme_bw(base_size = 15)+xlab('Creatinine (umol/L)')+
  geom_vline(xintercept = 265)+geom_hline(yintercept = mortality_threshold)+
  theme(legend.position = 'none')

p4=test_data %>%
  ggplot(aes(x = BUN, y = pred_mod_bun_remove_tract, 
             colour =Weight_category,linetype = Coma,
             group = interaction(Weight_category,Coma)))+
  geom_line()+ylab('Predicted mortality')+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks,limits=bun_limits) +
  annotation_logticks(sides='b')+theme_bw(base_size = 15)+xlab('BUN (mmol/L)')+
  geom_vline(xintercept = 20)+geom_hline(yintercept = mortality_threshold)+
  theme(legend.position = 'none')
p4_threshold=test_data %>%
  ggplot(aes(x = BUN, y = bun_full_threshold_tractrm, 
             colour = Weight_category, linetype = Coma,
             group = interaction(Weight_category,Coma)))+
  geom_line()+ylab('Probability mortality > 5%')+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks,limits=bun_limits) +
  annotation_logticks(sides='b')+theme_bw(base_size = 15)+xlab('BUN (mmol/L)')+
  geom_vline(xintercept = 20)+geom_hline(yintercept = mortality_threshold)+
  theme(legend.position = 'none')
```


```{r}
p5=test_data %>%
  ggplot(aes(x = BUN, y = pred_mod_bun_age, 
             colour = Age_category, linetype = Coma,
             group = interaction(Age_category,Coma)))+
  geom_line()+ylab('Predicted mortality')+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks,limits=bun_limits) +
  annotation_logticks(sides='b')+theme_bw(base_size = 15)+xlab('BUN (mmol/L)')+
  geom_vline(xintercept = 20)+geom_hline(yintercept = mortality_threshold)+
  theme(legend.position = 'none')
p5_threshold=test_data %>%
  ggplot(aes(x = BUN, y = bun_age_threshold, 
             colour = Age_category, linetype = Coma,
             group = interaction(Age_category,Coma)))+
  geom_line()+ylab('Probability mortality > 5%')+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks,limits=bun_limits) +
  annotation_logticks(sides='b')+theme_bw(base_size = 15)+xlab('BUN (mmol/L)')+
  geom_vline(xintercept = 20)+geom_hline(yintercept = mortality_threshold)+
  theme(legend.position = 'none')

p6=test_data %>%
  ggplot(aes(x = Creatinine_umol_L, y = pred_mod_creat_age, 
             colour = Age_category,linetype = Coma,
             group = interaction(Age_category,Coma)))+
  geom_line()+ylab('Predicted mortality')+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks,limits=crea_limits) +
  annotation_logticks(sides='b')+theme_bw(base_size = 15)+xlab('Creatinine (umol/L)')+
  geom_vline(xintercept = 265)+geom_hline(yintercept = mortality_threshold)+
  theme(legend.position = 'none')
p6_threshold=test_data %>%
  ggplot(aes(x = Creatinine_umol_L, y = creat_age_threshold, 
             colour = Age_category,linetype = Coma,
             group = interaction(Age_category,Coma)))+
  geom_line()+ylab('Probability mortality > 5%')+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks,limits=crea_limits) +
  annotation_logticks(sides='b')+theme_bw(base_size = 15)+xlab('Creatinine (umol/L)')+
  geom_vline(xintercept = 265)+geom_hline(yintercept = mortality_threshold)

```



```{r model_mortality_creat_bun}
grid.arrange(p1,p1_threshold,p2,p2_threshold)
grid.arrange(p3,p3_threshold,p4,p4_threshold)
grid.arrange(p6,p6_threshold+theme(legend.position = 'none'),
             p5,p5_threshold)
p6_threshold
```


```{r bun_creatinine_imputed}
# test_data %>%
#   ggplot(aes(x = BUN_imputed, y = pred_mod_bun_imputed, 
#              colour =Weight_category,linetype = Coma,
#              group = interaction(Age,Weight_category,Coma)))+
#   geom_line()+ylab('Predicted mortality')+
#   scale_x_log10(breaks = breaks, minor_breaks = minor_breaks,limits=bun_limits) +
#   annotation_logticks(sides='b')+theme_bw(base_size = 15)+xlab('BUN (mmol/L)')+
#   geom_vline(xintercept = 20)+geom_hline(yintercept = mortality_threshold)+
#   ggtitle('BUN imputed analysis')
# 
# 
# test_data %>%
#   ggplot(aes(x = Creatinine_imputed, y = pred_mod_creat_imputed, 
#              colour =Weight_category,linetype = Coma,
#              group = interaction(Age,Weight_category,Coma)))+
#   geom_line()+ylab('Predicted mortality')+
#   scale_x_log10(breaks = breaks, minor_breaks = minor_breaks,limits=crea_limits) +
#   annotation_logticks(sides='b')+theme_bw(base_size = 15)+xlab('Creatinine (umol/L)')+
#   geom_vline(xintercept = 265)+geom_hline(yintercept = mortality_threshold)+
#   ggtitle('Creatinine imputed analysis')
```


## Adjusting for bicarbonate

From Nick:
Incidentally can you do a bicarbonate vs lactate plot separating adults and children. It would be really interesting to look at urea or creatinine versus bicarbonate adjusting for lactate (i.e. two discrete sources of acid) in relation to age (e.g. younger kids vs older kids +adults). That might be a neat way of showing if kids are being affected by their renal impairment.

Multiple imputation of bicarbonate/base excess
```{r}
sm_data_analysis%>%ggplot(aes(x=Bicarbonate_mEq_L, y = `Base Excess_mmol_L`))+geom_point()
sm_data_analysis$Base_Excess_imp = sm_data_analysis$Bicarbonate_imp = NA
sm_data_analysis$BE = sm_data_analysis$`Base Excess_mmol_L`
mod_be = gam(BE ~ s(Bicarbonate_mEq_L), data = sm_data_analysis)
mod_bicarb = gam(Bicarbonate_mEq_L ~ s(BE), data = sm_data_analysis)
ind = !is.na(sm_data_analysis$Bicarbonate_mEq_L)
sm_data_analysis$Base_Excess_imp[ind]=predict(mod_be, newdata = sm_data_analysis[ind,])
ind = !is.na(sm_data_analysis$BE)
sm_data_analysis$Bicarbonate_imp[ind]=predict(mod_bicarb, newdata = sm_data_analysis[ind,])
sm_data_analysis$Base_Excess_imp = ifelse(is.na(sm_data_analysis$BE), sm_data_analysis$Base_Excess_imp, sm_data_analysis$BE)
sm_data_analysis$Bicarbonate_imp = ifelse(is.na(sm_data_analysis$Bicarbonate_mEq_L), sm_data_analysis$Bicarbonate_imp, sm_data_analysis$Bicarbonate_mEq_L)
sm_data_analysis%>%ggplot(aes(x=Bicarbonate_imp, y = Base_Excess_imp))+geom_point()
```


```{r lactate_bicarbonate}
sm_data_analysis$Lactate_category = cut(sm_data_analysis$Lactate,breaks = c(0,5,10,20),include.lowest = T)

sm_data_analysis %>% filter(Bicarbonate_mEq_L<30, !is.naLactate)%>%
  ggplot(aes(x=Bicarbonate_imp, y=BUN_imputed,colour = Lactate_category))+
  geom_point(aes(alpha=.5))+geom_smooth(method = lm)+
  scale_y_log10()+
  facet_wrap(~Age_category)+theme_minimal()+
  xlab('Bicarbonate (mEq/L)')+ylab('BUN (mmol/L)')

#Children - BUN
mod1=lm(Bicarbonate_imp ~ Lactate + (log2(BUN_imputed)),
           data = sm_data_analysis %>% filter(Age<=5))
summary(mod1)$coefficients

#Children - Cr
mod12=lm(Bicarbonate_imp ~ Lactate + (log2(Creatinine_imputed)),
           data = sm_data_analysis %>% filter(Age<=5))
summary(mod12)$coefficients

#Adults - BUN
mod2=lm(Bicarbonate_imp ~ Lactate + (log2(BUN_imputed)),
           data = sm_data_analysis %>% filter(Age>15))
summary(mod2)$coefficients

#Adults - Cr
mod22=lm(Bicarbonate_imp ~ Lactate + (log2(Creatinine_imputed)),
           data = sm_data_analysis %>% filter(Age>15))
summary(mod22)$coefficients
```



## Half thresholds??


```{r}
sm_data_analysis = sm_data_analysis %>%
  mutate(
    STUDY_Weight = paste(STUDY, Weight_category,sep = '_'),
    Above_BUN_threshold = 
      case_when(!is.na(BUN) & Weight_category=='(0,25]'&BUN>10 ~ 'YES',
                !is.na(BUN) & Weight_category=='(0,25]'&BUN<=10 ~ 'NO',
                !is.na(BUN) & Weight_category=='(25,100]'&BUN>20 ~ 'YES',
                !is.na(BUN) & Weight_category=='(25,100]'&BUN<20 ~ 'NO'),
    Above_Creat_threshold = 
      case_when(!is.na(Creatinine_umol_L) & Weight_category=='(0,25]'&Creatinine_umol_L>(265/2) ~ 'YES',
                !is.na(Creatinine_umol_L) & Weight_category=='(0,25]'&Creatinine_umol_L<=(265/2) ~ 'NO',
                !is.na(Creatinine_umol_L) & Weight_category=='(25,100]'&Creatinine_umol_L>265 ~ 'YES',
                !is.na(Creatinine_umol_L) & Weight_category=='(25,100]'&Creatinine_umol_L<=265 ~ 'NO'))%>%
  arrange(STUDY, Weight_category)


aggregate(Died ~ Above_BUN_threshold + STUDY, data = sm_data_analysis,
          FUN = function(x) c(round(100*mean(x)), sum(x),length(x)))

lyt <- basic_table() %>%
  split_cols_by('STUDY') %>%
  split_cols_by('Weight_category') %>%
  split_rows_by('Above_BUN_threshold')%>%
  summarize_row_groups() %>%
  analyze("Died", afun = function(x) c(sum(x),mean(x)), format = "xx (xx.x%)")
build_table(lyt = lyt, df = sm_data_analysis%>%filter(!is.na(BUN)))

lyt <- basic_table() %>%
  split_cols_by('STUDY') %>%
  split_cols_by('Weight_category') %>%
  split_rows_by('Above_Creat_threshold')%>%
  summarize_row_groups() %>%
  analyze("Died", afun = function(x) c(sum(x),mean(x)), format = "xx (xx.%)")
build_table(lyt = lyt, df = sm_data_analysis%>%filter(!is.na(Creatinine_umol_L)))

```


## Better prognosticator

```{r}
# sm_data_analysis_both$Lactate=ifelse(sm_data_analysis_both$`Lactic Acid_mmol_L`>50, NA, sm_data_analysis_both$`Lactic Acid_mmol_L`)
# 
# dim(sm_data_analysis_both)
# table(sm_data_analysis_both$STUDY)
# table(is.na(sm_data_analysis_both$`Lactic Acid_mmol_L`))
# 
# sm_data_analysis_both$STUDY=as.factor(sm_data_analysis_both$STUDY)
# 
# mod_both_BUN = gam(Died ~ scale(log10(BUN)) + s(COUNTRY,bs='re') , data = sm_bun , family = 'binomial')
# 
# mod_both_crea = gam(Died ~ scale(log10(Creatinine_umol_L)) + s(COUNTRY,bs='re') , data = sm_creat, family = 'binomial')
# 
# mod_both = gam(Died ~ scale(log10(Creatinine_umol_L)) + scale(log10(BUN))+ scaleLactate + s(STUDY,bs='re') , data = sm_data_analysis_both, family = 'binomial')
# 
# 
# exp(coef(mod_both_BUN)[2])
# exp(coef(mod_both_crea)[2])
# exp(coef(mod_both)[c(2,3,4)])

```





