---
title: "Renal Analysis Severe Malaria"
author: "James Watson"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = T, cache.comments = FALSE,
                      echo = F, include = TRUE, 
                      fig.width = 8, fig.height = 8,
                      fig.pos = 'H',dev = 'png', dpi = 300)


library(tidygam)
library(mgcv)
library(conflicted)
library(tidyverse)
library(readr)
library(rtables)
library(RColorBrewer)
library(grid)
library(gridExtra)
library(worlddatr)

conflicted::conflicts_prefer(dplyr::filter)


binomial_smooth <- function(...) {
  geom_smooth(method = "gam", method.args = list(family = "binomial"), ...)
}
```

## Load data


```{r load_data}
sm_data = read_csv('Data/Renal_Analysis_Data.csv')
sm_data = sm_data%>%arrange(STUDY)
table(sm_data$STUDY)

# Remove late deaths in TRACT trial
sm_data$Died[sm_data$STUDY=='TRACT' & sm_data$Time_to_death>28]=0

# impute weight based on age
table(Age_missing=is.na(sm_data$Age), Weight_missing=is.na(sm_data$Weight_kg))
sm_data$SITEID=as.factor(sm_data$SITEID)
sm_data$COUNTRY=as.factor(sm_data$COUNTRY)

sm_data$Lactate = sm_data$`Lactic Acid_mmol_L`
sm_data = sm_data%>% select(-`Lactic Acid_mmol_L`)
sm_data$Lactate = ifelse(sm_data$Lactate>50,NA,sm_data$Lactate)
```

## Variables in dataset

### Hemoglobin-Hematocrit

```{r hb_hct}
source('hb_data_plotting.R')
sm_data=hb_hct_impute(sm_data = sm_data)
```




## Make analysis dataset
```{r}
sm_data_analysis = sm_data %>% 
  filter(Malaria_Positive, !is.na(Died), !is.na(Creatinine_umol_L) | !is.na(BUN))
table(sm_data_analysis$STUDY)
table(!is.na(sm_data_analysis$Creatinine_umol_L))
table(!is.na(sm_data_analysis$BUN))

sm_data_analysis_both = sm_data_analysis %>% 
  filter(!is.na(Creatinine_umol_L), !is.na(BUN)) 

# rm(sm_data)

aggregate(Age ~ STUDY, data = sm_data_analysis, FUN = quantile)
aggregate(RFSTDTC ~ STUDY, data = sm_data_analysis, FUN = function(x) range(x,na.rm = T))
aggregate(Died ~ STUDY, data = sm_data_analysis, FUN = function(x) round(100*mean(x),1))
```
plotting params

```{r}
sqrt_breaks = c(1,4,9,16,25,36,50,64,81)
breaks <- 10^(-10:10)
minor_breaks <- rep(1:9, 21)*(10^rep(-10:10, each=9))

SEsmooth=F
my_alpha=0.2
bun_limits = c(1,100)
crea_limits = c(10,1000)

mortality_threshold = 0.05
```


map
```{r patient_map}
# create_participant_map(DATA_DM= sm_data_analysis, include_ATA = FALSE)
```



Analysis data

```{r analysis_data}
writeLines(sprintf('Analysis data: there are %s patients from %s studies with at least one baseline creatinine or BUN measurement. %s patients have creatinine; %s patients have BUN; %s patients have both.',
                   nrow(sm_data_analysis),
                   length(unique(sm_data_analysis$STUDYID)),
                   sum(!is.na(sm_data_analysis$Creatinine_umol_L)),
                   sum(!is.na(sm_data_analysis$BUN)),
                   sum(!is.na(sm_data_analysis$Creatinine_umol_L) & !is.na(sm_data_analysis$BUN))
))


lyt <- basic_table() %>%
  split_cols_by("STUDY") %>%
  summarize_row_groups() %>%
  analyze("Age", afun = function(x) quantile(x, probs = c(0.25,.5,.75)), format = "xx.x (xx.x - xx.x)")%>%
  analyze(c('Creatinine_umol_L',"BUN"), afun = function(x) sum(!is.na(x)), format = "xx.x")
build_table(lyt = lyt, df = sm_data_analysis)
```



How do the distributions of BUN and creatinine vary across studies in the children?

```{r boxplots}
sm_data_analysis %>% filter(Weight_kg<30) %>%
  ggplot(aes(x=STUDY, y=Creatinine_umol_L)) + geom_boxplot()+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks,limits=c(1,100))

sm_data_analysis %>% filter(Weight_kg<30) %>%
  ggplot(aes(x=STUDY, y=BUN)) + geom_boxplot()+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks,limits=c(1,100))
```

# xx=sm_data_analysis %>% group_by(SITEID, STUDY) %>%
#   mutate(group_mortality = mean(Died),
#          group_n=n(),
#          group_median_creatinine=median(Creatinine_umol_L,na.rm=T),
#          group_median_bun=median(BUN,na.rm=T)) %>%
#   distinct(SITEID, STUDY, .keep_all = T) %>%
#   select(SITEID, STUDY, group_mortality, group_n, group_median_creatinine, group_median_bun)
# xx %>% ggplot(aes(x=group_median_creatinine, y=group_mortality))+geom_point()+geom_smooth(method = lm)
# xx %>% ggplot(aes(x=group_median_bun, y=group_mortality))+geom_point()+geom_smooth(method = lm)





```{r creatinine_v_bun, fig.width=12, fig.height=8}
p1=sm_data_analysis_both %>%
  ggplot(aes(x=Creatinine_umol_L, y = BUN, colour = STUDY))+
  geom_point(alpha=0.2)+geom_smooth(se = F)+
  scale_y_continuous(transform  ='log10') +
  geom_hline(yintercept = 20)+geom_vline(xintercept = 265)+
  xlab('Creatinine (umol/L)')+ylab('BUN (mmol/L)')+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks,limits=crea_limits) +
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks,limits=bun_limits) +
  annotation_logticks() + theme_bw(base_size = 20)+
  theme(
    legend.position = c(0.97, 0.03), 
    legend.justification = c(1, 0), 
    legend.background = element_rect(fill = "white", color = "black", size = 0.5),
    legend.margin = margin(6, 6, 6, 6),  
    legend.text = element_text(size = 9),
    legend.title = element_text(size = 10))


p2=sm_data_analysis_both %>%
  ggplot(aes(x=(Creatinine_umol_L/BUN), colour = STUDY))+
  geom_density()+theme_bw(base_size = 20)+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks,limits=c(1,100))+
  annotation_logticks(sides='b')+
  xlab('Creatinine to BUN ratio')+ylab('')+
  scale_y_discrete(labels = NULL, breaks = NULL)+
  theme(legend.position = 'none')

grid.arrange(p1,p2,nrow=1)
```


```{r weight_age, fig.width=12, fig.height=8}
p1=sm_data_analysis %>% filter(!is.na(Creatinine_umol_L)) %>%
  ggplot(aes(x=Age, y = Creatinine_umol_L, colour = STUDY))+
  geom_jitter(height = 0.05, width = .1, alpha=my_alpha)+
  ylab('Creatinine (umol/L)')+xlab('Age (years)')+
  geom_hline(yintercept = 265)+
  scale_y_continuous(transform  ='log10') + 
  scale_x_continuous(transform  ='sqrt',breaks = sqrt_breaks) +
  geom_smooth(se=SEsmooth)+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks) +
  annotation_logticks(sides='l') +
  theme_bw(base_size = 15)+
  theme(legend.position="none")

xx = sm_data_analysis %>% filter(!is.na(Creatinine_umol_L), !is.na(Weight_kg)) 
writeLines(sprintf('There are %s patients with weight and Creatinine jointly measured', nrow(xx)))
p2 = xx%>%
  ggplot(aes(x=Weight_kg, y = Creatinine_umol_L, colour = STUDY))+
  geom_jitter(height = 0.01, width = .1, alpha=my_alpha)+
  ylab('Creatinine (umol/L)')+xlab('Weight (kg)')+
  geom_hline(yintercept = 265)+
  scale_x_continuous(transform  ='sqrt',breaks = sqrt_breaks) + 
  geom_smooth(se=SEsmooth)+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks,limits=crea_limits) +
  annotation_logticks(sides='l') +
  theme_bw(base_size = 15)+
  theme(
    legend.position = c(0.97, 0.03), 
    legend.justification = c(1, 0),
    legend.background = element_rect(fill = "white", color = "black", size = 0.5),
    legend.margin = margin(6, 6, 6, 6),  
    legend.text = element_text(size = 9),
    legend.title = element_text(size = 10))

p3=sm_data_analysis %>% 
  ggplot(aes(x=Age, y = BUN, colour = STUDY))+
  geom_jitter(height = 0.05, width = .1, alpha=my_alpha)+
  ylab('BUN (mmol/L)')+xlab('Age (years)')+
  geom_hline(yintercept = 20)+
  scale_y_continuous(transform='log10') + 
  scale_x_continuous(transform='sqrt',breaks = sqrt_breaks) +
  geom_smooth(se=SEsmooth)+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks) +
  annotation_logticks(sides='l') +
  theme_bw(base_size = 20)#+theme(legend.position="none")

xx = sm_data_analysis %>% filter(!is.na(BUN), !is.na(Weight_kg)) 
writeLines(sprintf('There are %s patients with weight and BUN jointly measured', nrow(xx)))
p4=xx%>%ggplot(aes(x=Weight_kg, y = BUN, colour = STUDY))+
  geom_jitter(height = 0.03, width = .1, alpha=my_alpha)+
  ylab('BUN (mmol/L)')+xlab('Weight (kg)')+
  geom_hline(yintercept = 20)+
  scale_x_continuous(transform='sqrt',breaks = sqrt_breaks) +
  geom_smooth(se=SEsmooth)+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks,limits=bun_limits) +
  annotation_logticks(sides='l') + theme_bw(base_size = 15)+
  theme(
    legend.position = c(0.97, 0.03),        # Position the legend inside the plot (bottom-right corner)
    legend.justification = c(1, 0),       # Justify the legend to the bottom-right
    legend.background = element_rect(fill = "white", color = "black", size = 0.5),  # Optional: adds border to legend
    legend.margin = margin(6, 6, 6, 6),    # Adjust the margin around the legend if needed
    legend.text = element_text(size = 9),
    legend.title = element_text(size = 10)
  )

grid.arrange(p2,p4,nrow=1)

# 
# sm_data_analysis %>% ggplot(aes(x=para_ul+50, y = BUN, colour = STUDY))+
#   geom_point(alpha=0.05)+xlab('Parasitaemia (/uL)')+
#   scale_y_log10(breaks = breaks, minor_breaks = minor_breaks) +
#   scale_x_log10(breaks = breaks, minor_breaks = minor_breaks) +
#   annotation_logticks(sides='lb') +
#   geom_smooth(aes())+theme_bw()+geom_hline(yintercept = 20)

xx = sm_data_analysis %>%
  filter(STUDY %in% c('AQUAMAT','SEAQUAMAT','FEAST','TRACT'), !is.na(BUN)) 
dim(xx)
xx %>% ggplot(aes(x=Weight_kg, y = BUN, colour = STUDY))+
  geom_jitter(height = 0.02, width = .02, alpha=my_alpha)+
  ylab('BUN (mmol/L)')+xlab('Weight (kg)')+
  geom_hline(yintercept = 20)+
  scale_y_continuous(transform='log10') +
  scale_x_continuous(transform='sqrt',breaks = sqrt_breaks,limits=bun_limits) + 
  geom_smooth()+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks) +
  annotation_logticks(sides='l') +
  theme_bw(base_size = 15)#+theme(legend.position="none")
```


Remove AQ Gambia from analysis dataset as unreliable creatinines
```{r}
sm_data_analysis = sm_data_analysis %>% filter(!STUDY=='AQ Gambia')
```


## Mortality models


```{r}
xx=sm_data_analysis%>%filter(Coma_Final==0, !is.na(Creatinine_umol_L))
xx %>% ggplot(aes(x=BUN,y=Died, color=Weight_category))+
  scale_x_continuous(transform='log10') +
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks) +
  annotation_logticks(sides='b') +
  binomial_smooth()+theme_minimal()+
  #theme(plot.margin = unit(c(0.5, 0.5, 0, 0.5), "cm"), legend.position = "none")+
  ylab('Probability of death')+xlab('BUN (mmol/L)')+
  geom_hline(yintercept = mortality_threshold)+
  coord_cartesian(ylim=c(0, .4), xlim = c(1,100))


xx=sm_data_analysis%>%filter(Coma_Final==0, !is.na(BUN))
xx %>% ggplot(aes(x=BUN,y=Died, color=Weight_category))+
  scale_x_continuous(transform='log10') +
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks) +
  annotation_logticks(sides='b') +
  binomial_smooth()+theme_minimal()+
  #theme(plot.margin = unit(c(0.5, 0.5, 0, 0.5), "cm"), legend.position = "none")+
  ylab('Probability of death')+xlab('BUN (mmol/L)')+
  geom_hline(yintercept = mortality_threshold)+
  coord_cartesian(ylim=c(0, .4), xlim = c(1,40))

```



```{r mortality_models, fig.width=12, fig.height=8}
sm_data_analysis$Weight_category = as.factor(sm_data_analysis$Weight_category)
sm_data_analysis$COUNTRY=as.factor(sm_data_analysis$COUNTRY)

### Creatinine model
sm_creat = sm_data_analysis %>%
  filter(!is.na(Creatinine_umol_L), !is.na(Weight_category),
         Malaria_Positive, Coma_Final==0
  ) 
table(sm_creat$STUDY)
dim(sm_creat)
mean(sm_creat$Died)
table(sm_creat$Weight_category)

mod1 = gam(Died ~ s(log10(Creatinine_umol_L),k=3)+s(SITEID, bs='re'), 
           family = 'binomial',data = sm_creat %>% filter(Weight_category=='(0,25]'))
mod2 = gam(Died ~ s(log10(Creatinine_umol_L),k=3)+s(SITEID, bs='re'), 
           family = 'binomial',data = sm_creat %>% filter(Weight_category!='(0,25]'))
pred1=predict_gam(mod1,tran_fun = gtools::inv.logit,series = "Creatinine_umol_L",
                  values = list(SITEID= -1, 
                                Creatinine_umol_L=10^seq(1.3,2.8,length.out=20))) 
pred1$Weight_category='(0,25]'
pred2=predict_gam(mod2,tran_fun = gtools::inv.logit,series = "Creatinine_umol_L",
                  values = list(SITEID= -1, 
                                Creatinine_umol_L=10^seq(1.3,2.8,length.out=20))) 
pred2$Weight_category='(25,100]'
preds = rbind(pred1, pred2)


h1=sm_creat%>%ggplot(aes(x = Creatinine_umol_L,colour = Weight_category))+
  scale_x_log10(limits=crea_limits) + geom_density()+theme_minimal()+theme_bw(base_size = 15)+
  theme(axis.title.x = element_blank(), axis.text.x = element_blank(), 
        legend.position="none",
        plot.margin = unit(c(0, 0.5, 0, 0.5), "cm"))+
  ggtitle(label = paste('No coma (n=',sum(xx$Coma_Final==0,na.rm = T),')',sep = ''))
p1=preds %>%
  plot(comparison='Weight_category')+ylim(0,.25)+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks,limits=crea_limits) +
  annotation_logticks(sides='b')+theme_bw(base_size = 15)+xlab('Creatinine (umol/L)')+
  geom_vline(xintercept = 265)+geom_hline(yintercept = mortality_threshold)+
  theme(legend.position="none",
        plot.margin = unit(c(0.5, 0.5, 0, 0.5), "cm"))


### BUN model
sm_bun = sm_data_analysis %>%
  filter(!is.na(BUN), !is.na(Weight_category),
         Malaria_Positive, Coma_Final==0
  ) 
table(sm_bun$STUDY)
dim(sm_bun)
mean(sm_bun$Died)
table(sm_bun$Weight_category)

mod_all = gam(Died ~ s(log10(BUN),by = Weight_category)+s(COUNTRY, bs='re'), 
              family = 'binomial',data = sm_creat)
pred1=predict_gam(mod_all,tran_fun = gtools::inv.logit,series = "BUN",
                  values = list(COUNTRY= -1, 
                                Weight_category='(0,25]',
                                BUN=10^seq(0,1.45,length.out=20) )) 
pred1$Weight_category='(0,25]'
pred2=predict_gam(mod_all,tran_fun = gtools::inv.logit,series = "BUN",
                  values = list(COUNTRY= -1, 
                                Weight_category='(25,100]',
                                BUN=10^seq(0,2,length.out=20))) 
pred2$Weight_category='(25,100]'
preds = rbind(pred1, pred2)

sm_bun$BUN_deciles = 
  cut(sm_bun$BUN, 
      breaks = round(quantile(sm_bun$BUN,probs=seq(0,1,by=0.1)),1),
      include.lowest = T)

sm_bun_summary=sm_bun %>% group_by(BUN_deciles, Weight_category) %>%
  mutate(Bun_decile_mortality=mean(Died),
         n_deaths = sum(Died),
         n = n(),
         lower_ci = qbeta(0.025, 1+n_deaths, 1+(n-n_deaths)),
         upper_ci = qbeta(0.975, 1+n_deaths, 1+(n-n_deaths)) ) %>%
  distinct(BUN_deciles, Weight_category, .keep_all = T) %>%
  select(Weight_category,BUN_deciles,
         Bun_decile_mortality,n,n_deaths,lower_ci, upper_ci)%>% 
  arrange(Weight_category, BUN_deciles)


sm_bun_summary%>%
  ggplot(aes(x=BUN_deciles, y=Bun_decile_mortality,
                            colour = Weight_category))+
  geom_point()+
  geom_linerange(aes(ymin = lower_ci,ymax = upper_ci))+
  theme_minimal()+
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

plot(sm_bun_summary$BUN_deciles, sm_bun_summary$Bun_decile_mortality)
h2=sm_bun%>%ggplot(aes(x = BUN,colour = Weight_category))+
  scale_x_log10(limits=bun_limits) + geom_density()+theme_minimal()+theme_bw(base_size = 15)+
  theme(axis.title.x = element_blank(), axis.text.x = element_blank(), 
        legend.position="none",
        plot.margin = unit(c(0, 0.5, 0, 0.5), "cm"))+
  ggtitle(label = paste('No coma (n=',sum(xx$Coma_Final==0,na.rm = T),')',sep = ''))
p2=preds %>%
  ggplot(aes(x=BUN,y = Died, color = Weight_category))+
  geom_ribbon(aes(ymin=lower_ci, ymax = upper_ci,alpha=0.2,fill = Weight_category))+
  geom_line()+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks,limits=bun_limits) +
  annotation_logticks(sides='b')+theme_bw(base_size = 15)+xlab('BUN (mmol/L)')+
  geom_vline(xintercept = 265)+geom_hline(yintercept = mortality_threshold)+
  coord_cartesian(ylim = c(0,0.25))

plot(comparison='Weight_category')+ylim(0,.25)+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks,limits=bun_limits) +
  annotation_logticks(sides='b')+theme_bw(base_size = 15)+xlab('BUN (mmol/L)')+
  geom_vline(xintercept = 265)+geom_hline(yintercept = mortality_threshold)+
  theme(legend.position="none",
        plot.margin = unit(c(0.5, 0.5, 0, 0.5), "cm"))


grid.arrange(h1, h2, p1,p2,nrow=2, ncol = 2, heights = c(1, 1.2))


sm_bun = sm_data_analysis %>%
  filter(!is.na(BUN), 
         Malaria_Positive,
         Coma_Final==0
  ) 
nrow(sm_bun)
table(sm_bun$Weight_category)
mean(sm_bun$Died)

aggregate(Died ~ Weight_category + (BUN>20), data = sm_bun, FUN = mean)
aggregate(Died ~ STUDY + (BUN>10), data = sm_bun, FUN = function(x) c(round(100*mean(x)), length(x)))
aggregate(Died ~ Weight_category + (BUN>10), data = sm_bun, FUN = function(x) c(round(100*mean(x)), length(x)))

mod_all = gam(Died ~ s(log10(BUN), by=Weight_category)+s(COUNTRY, bs='re'), 
              family = 'binomial',data = sm_bun)
p2=predict_gam(mod_all,tran_fun = gtools::inv.logit,series = "BUN",
               values = list(COUNTRY= -1, 
                             BUN=10^seq(0,2,length.out=20),
                             Weight_category = unique(sm_bun$Weight_category))) %>%
  plot(comparison='Weight_category')+ylim(0,.25)+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks,limits=c(2,40)) +
  annotation_logticks(sides='b')+theme_bw(base_size = 20)+xlab('Urea Nitrogen (mmol/L)')+
  geom_vline(xintercept = 20)+geom_hline(yintercept = mortality_threshold)+
  theme(legend.position="none")



```




## Half thresholds??


```{r}
sm_data_analysis = sm_data_analysis %>%
  mutate(
    STUDY_Weight = paste(STUDY, Weight_category,sep = '_'),
    Above_BUN_threshold = 
      case_when(!is.na(BUN) & Weight_category=='(0,25]'&BUN>10 ~ 'YES',
                !is.na(BUN) & Weight_category=='(0,25]'&BUN<=10 ~ 'NO',
                !is.na(BUN) & Weight_category=='(25,100]'&BUN>20 ~ 'YES',
                !is.na(BUN) & Weight_category=='(25,100]'&BUN<20 ~ 'NO'),
    Above_Creat_threshold = 
      case_when(!is.na(Creatinine_umol_L) & Weight_category=='(0,25]'&Creatinine_umol_L>(265/2) ~ 'YES',
                !is.na(Creatinine_umol_L) & Weight_category=='(0,25]'&Creatinine_umol_L<=(265/2) ~ 'NO',
                !is.na(Creatinine_umol_L) & Weight_category=='(25,100]'&Creatinine_umol_L>265 ~ 'YES',
                !is.na(Creatinine_umol_L) & Weight_category=='(25,100]'&Creatinine_umol_L<=265 ~ 'NO'))%>%
  arrange(STUDY, Weight_category)


aggregate(Died ~ Above_BUN_threshold + STUDY, data = sm_data_analysis,
          FUN = function(x) c(round(100*mean(x)), sum(x),length(x)))

lyt <- basic_table() %>%
  split_cols_by('STUDY') %>%
  split_cols_by('Weight_category') %>%
  split_rows_by('Above_BUN_threshold')%>%
  summarize_row_groups() %>%
  analyze("Died", afun = function(x) c(sum(x),mean(x)), format = "xx (xx.x%)")
build_table(lyt = lyt, df = sm_data_analysis%>%filter(!is.na(BUN)))

lyt <- basic_table() %>%
  split_cols_by('STUDY') %>%
  split_cols_by('Weight_category') %>%
  split_rows_by('Above_Creat_threshold')%>%
  summarize_row_groups() %>%
  analyze("Died", afun = function(x) c(sum(x),mean(x)), format = "xx (xx.%)")
build_table(lyt = lyt, df = sm_data_analysis%>%filter(!is.na(Creatinine_umol_L)))

```


## Better prognosticator

```{r}
# sm_data_analysis_both$Lactate=ifelse(sm_data_analysis_both$`Lactic Acid_mmol_L`>50, NA, sm_data_analysis_both$`Lactic Acid_mmol_L`)
# 
# dim(sm_data_analysis_both)
# table(sm_data_analysis_both$STUDY)
# table(is.na(sm_data_analysis_both$`Lactic Acid_mmol_L`))
# 
# sm_data_analysis_both$STUDY=as.factor(sm_data_analysis_both$STUDY)
# 
# mod_both_BUN = gam(Died ~ scale(log10(BUN)) + s(COUNTRY,bs='re') , data = sm_bun , family = 'binomial')
# 
# mod_both_crea = gam(Died ~ scale(log10(Creatinine_umol_L)) + s(COUNTRY,bs='re') , data = sm_creat, family = 'binomial')
# 
# mod_both = gam(Died ~ scale(log10(Creatinine_umol_L)) + scale(log10(BUN))+ scale(Lactate) + s(STUDY,bs='re') , data = sm_data_analysis_both, family = 'binomial')
# 
# 
# exp(coef(mod_both_BUN)[2])
# exp(coef(mod_both_crea)[2])
# exp(coef(mod_both)[c(2,3,4)])

```




