---
title: "Renal Analysis Severe Malaria"
author: "James Watson"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = T, cache.comments = FALSE,
                      echo = F, include = TRUE, 
                      fig.width = 8, fig.height = 8,
                      fig.pos = 'H',dev = 'png', dpi = 300)


library(tidygam)
library(mgcv)
library(brms)
library(conflicted)
library(tidyverse)
library(readr)
library(rtables)
library(RColorBrewer)
library(grid)
library(gridExtra)
library(naniar)
library(gtools)
library(broom)
library(zscorer)
library(mice)
library(rstanarm)
# library(tidybayes)
#library(worlddatr)

conflicts_prefer(dplyr::filter)
conflicts_prefer(brms::ar)
conflicts_prefer(dplyr::collapse)
conflicts_prefer(dplyr::combine)
conflicts_prefer(tidyr::extract)
conflicts_prefer(dplyr::lag)
conflicts_prefer(brms::s)
conflicts_prefer(purrr::set_names)
conflicts_prefer(brms::t2)
conflicts_prefer(dplyr::group_rows)
conflicted::conflicts_prefer(brms::ddirichlet)
conflicted::conflicts_prefer(brms::rdirichlet)
conflicted::conflicts_prefer(mgcv::scat)

binomial_smooth <- function(...) {
  geom_smooth(method = "gam", method.args = list(family = "binomial"), ...)
}
```



Plotting params

```{r}
sqrt_breaks = c(1,4,9,16,25,36,50,64,81)
breaks <- 10^(-10:10)
minor_breaks <- rep(1:9, 21)*(10^rep(-10:10, each=9))

SEsmooth=F
my_alpha=0.5
bun_limits = c(2,50)
crea_limits = c(20,500)

mortality_threshold = 0.05
```


## Load data

Analysis dataset
```{r}
sm_data = readr::read_csv(file = 'Data/Analysis_data.csv')
```


## Weight for age z-scores

```{r}
sm_data = sm_data  %>%
  mutate(sex = ifelse(SEX=='M',1,2),
         age_days = ifelse(Age<10, Age*365, NA))
sm_data =
  addWGSR(data = sm_data, sex = 'sex', 
          firstPart = 'Weight_kg', 
          secondPart = 'age_days',index = 'wfa')
sm_data$wfaz = ifelse(sm_data$wfaz>5 | sm_data$wfaz< -6, NA, sm_data$wfaz)
sm_data %>% filter(!is.na(wfaz), STUDY != 'Chittagong') %>% 
  ggplot(aes(x= Age, y = wfaz, color=SEX))+geom_point()+geom_hline(yintercept = 0) +
  theme_minimal() + xlab('Age (years)')+ylab('Weight for age z-score')+
  facet_wrap(~STUDY)
```


make analysis dataset

```{r}
sm_data_analysis = sm_data %>% 
  filter(INCLUDE_CC_ANALYSIS) %>%
  mutate(Coma = ifelse(Coma_Final==1,'Yes','No'),
         Weight_category = factor(Weight_category, levels = c(">30","≤30")),
         Age_category = factor(Age_category, levels = c(">15","(5,15]","(2,5]","≤2")))

sm_data_analysis = sm_data_analysis %>%
  mutate(
    Coma = as.factor(Coma),
    COUNTRY=as.factor(COUNTRY),
    SITE_STUDY=paste(SITEID, STUDY,sep = '_'),
    SITE_STUDY = as.factor(SITE_STUDY),
    BUN_ANALYSIS = !is.na(BUN),
    CREAT_ANALYSIS = !is.na(Creatinine_umol_L),
    BUN_CREAT_ANALYSIS = BUN_ANALYSIS & CREAT_ANALYSIS)

load('RData/imputed_dataset.RData')
imp_data = list()
for(i in 1:imp_list$m){
  xx = complete(imp_list, action = i)
  xx = xx[sm_data$INCLUDE_CC_ANALYSIS, ]
  imp_data[[i]] = xx
}
rm(imp_list)
```


# Calculate expected Cr and BUN based on age, weight

## Baseline creatinines

```{r}
p1=sm_data %>% filter(!is.na(Baseline_Creatinine)) %>%
  ggplot(aes(x=Weight_kg, y = Baseline_Creatinine))+geom_point()+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks)+theme_minimal()+
  xlab('Weight (kg)')+ylab('Normal creatinine (umol/L)')+geom_smooth(method=MASS::rlm)
summary(lm(Baseline_Creatinine~Weight_kg, data = sm_data))
eGFR=120 # Using Pottel equation
p2=sm_data %>% filter(!is.na(Baseline_Creatinine)) %>%
  ggplot(aes(x=Age, y = Baseline_Creatinine))+geom_point()+
  geom_abline(slope = (107.3/eGFR)*0.0270*88.4,
              intercept = (107.3/eGFR)*0.2329*88.4)+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks)+theme_minimal()+
  xlab('Age (years)')+ylab('Normal creatinine (umol/L)')+geom_smooth(method=MASS::rlm)
summary(lm(Baseline_Creatinine~Age, data = sm_data))
p3=sm_data %>% filter(!is.na(Baseline_Creatinine)) %>%
  ggplot(aes(x=Height_cm, y = Baseline_Creatinine))+geom_point()+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks)+theme_minimal()+
  xlab('Height (cm)')+ylab('Normal creatinine (umol/L)')+geom_smooth(method=MASS::rlm)
summary(lm(Baseline_Creatinine~Height_cm, data = sm_data))
p4=sm_data %>% filter(!is.na(Baseline_Creatinine)) %>%
  ggplot(aes(x=wfaz, y = Baseline_Creatinine))+geom_point()+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks)+theme_minimal()+
  xlab('Weight-for-age z-score')+ylab('Normal creatinine (umol/L)')+geom_smooth(method=MASS::rlm)
summary(lm(Baseline_Creatinine~wfaz, data = sm_data))
grid.arrange(p1,p2,p3,p4)

p1=sm_data %>% filter(!is.na(`Baseline_Urea Nitrogen`)) %>%
  ggplot(aes(x=Weight_kg, y = `Baseline_Urea Nitrogen`, colour = Age_category))+geom_point()+scale_x_sqrt()+
  scale_y_log10()+theme_minimal()+
  xlab('Weight (kg)')+ylab('Normal BUN (mmol/L)')

p2=sm_data %>% filter(!is.na(`Baseline_Urea Nitrogen`)) %>%
  ggplot(aes(x=Age, y = `Baseline_Urea Nitrogen`))+geom_point()+
  scale_y_log10()+theme_minimal()+scale_x_sqrt()+
  xlab('Age (years)')+ylab('Normal BUN (mmol/L)')

grid.arrange(p1,p2)

sm_data %>% filter(!is.na(`Baseline_Urea Nitrogen`)) %>%
  ggplot(aes(x=Baseline_Creatinine, y = `Baseline_Urea Nitrogen`))+geom_point()+
  scale_y_log10()+
  scale_x_log10()+
  theme_minimal()+
  xlab('Normal Creatinine (umol/L)')+ylab('Normal BUN (mmol/L)')

sm_data %>% filter(!is.na(`Baseline_Urea Nitrogen`)) %>%
  ggplot(aes(x=Baseline_Creatinine / `Baseline_Urea Nitrogen`))+geom_histogram()+
  scale_x_log10()+
  theme_minimal()+
  xlab('Normal Creatinine (umol/L)')+ylab('Normal BUN (mmol/L)')


xx = sm_data %>% filter(!is.na(Baseline_Creatinine) | !is.na(`Baseline_Urea Nitrogen`))
xx$MUAC = xx$`Mid-Upper Arm Circumference_cm`
xx$Baseline_BUN = xx$`Baseline_Urea Nitrogen`
mod_creat_baseline=stan_lm(Baseline_Creatinine ~ Age+Height_cm+`Mid-Upper Arm Circumference_cm`,
            data = xx, prior = R2(0.5))
# plot(mod_creat_baseline,pars = c('Age','Height_cm','SEXM',`Mid-Upper Arm Circumference_cm`))
summary(mod_creat_baseline)

mod_bun_baseline=stan_lm(Baseline_BUN ~ scale(Age) + scale(Weight_kg)+scale(Height_cm)+SEX+scale(MUAC),
            data = xx, prior = R2(0.5))
plot(mod_bun_baseline,pars = c('scale(Age)','scale(Weight_kg)','scale(Height_cm)','SEXM','scale(MUAC)'))
summary(mod_bun_baseline)
```

Pottel age-based equation, where eGFR=107.3/(SCr/Q)
assuming a normal GFR of 120 mL/min per 1.73 m2 (height-independent). 
Q=0.0270*age+0.2329.

```{r}
sm_data_analysis = sm_data_analysis %>%
  mutate(
    Q = 0.0270*Age + 0.2329,
    bSrc_adults = 88.4*(0.6+0.01*(Age)+0.002*(Weight_kg) - 0.02*as.numeric(SEX=='F')),
    bSrc_children = (107.3/120)*Q*88.4,
    bSrc = ifelse(Age<15, bSrc_children, bSrc_adults),
    Creatinine_fold_change = Creatinine_umol_L/bSrc)
plot(sm_data_analysis$Age, sm_data_analysis$bSrc_adults,
     ylim=c(0,150),xlab='Age (years)', 
     ylab = 'Estimated baseline Sr (umol/L)',
     panel.first=grid())
lines(sm_data_analysis$Age, sm_data_analysis$bSrc_children,col='red',lwd=2)
```


TRACT predicted baseline creatinine for children
```{r}
preds_baseline_creat = list()
for(i in 1:length(imp_data)){
  preds_baseline_creat[[i]] = colMeans(posterior_predict(object = mod_creat_baseline, newdata = imp_data[[i]]))
}
plot(preds_baseline_creat[[1]], preds_baseline_creat[[2]])
preds_baseline_creat = colMeans(do.call(base::rbind, preds_baseline_creat))
sm_data_analysis$bSrc_children_tract = preds_baseline_creat
sm_data_analysis %>% filter(Age<15) %>%
  ggplot(aes(x=bSrc_children, y = bSrc_children_tract))+geom_point()
```


is there heterogeneity by age/weight?

```{r mod_age_het}
sm_data_creat = sm_data_analysis %>% filter(!is.na(Creatinine_umol_L))

f_brm_mod_creat = as.formula(bf(Died_28D ~ s(log10(Creatinine_fold_change),by=Age_category)+Coma+(1|STUDY/SITE_STUDY)))

f_brm_mod_creat2 = as.formula(bf(Died_28D ~ s(log10(Creatinine_fold_change),by=STUDY)+Coma+(1|STUDY/SITE_STUDY)))

fpath = 'RData/expected_change_renal_models.RData'
RUN_STAN_MODELS=F
if(!file.exists(f_path) | RUN_STAN_MODELS){
  mod_creat_age_het = 
    brm(formula = f_brm_mod_creat,
        data = sm_data_creat ,
        family = bernoulli(), cores = 4, seed = 17,
        iter = 2000, warmup = 1000, thin = 1, refresh = 250)
  
  mod_creat_study_het = 
    brm(formula = f_brm_mod_creat2,
        data = sm_data_creat ,
        family = bernoulli(), cores = 4, seed = 17,
        iter = 2000, warmup = 1000, thin = 1, refresh = 250)
  save(mod_creat_age_het, mod_creat_study_het,file = fpath)
} else {
  load(f_path)
}
```


Test data

```{r}
k_vals=40
test_data = expand_grid(Creatinine_fold_change = 10^seq(0,1,length.out=k_vals),
                        # BUN = 10^seq(0,2,length.out=k_vals),
                        Age_category=c(">15","(5,15]","(2,5]","≤2"),
                        Weight_category=unique(sm_data_creat$Weight_category),
                        Coma=unique(sm_data_creat$Coma),
                        STUDY = unique(sm_data_creat$STUDY), 
                        SITE_STUDY = -1) %>% arrange(Creatinine_fold_change)

# remove non plausible combinations
test_data = test_data %>%
  filter(!(Age_category==">15" & Weight_category=='(0,25]'),
         !(Age_category %in% c("(2,5]","≤2") & Weight_category=='(25,100]')) %>%
  mutate(Age_category=factor(Age_category, levels = c("≤2","(2,5]","(5,15]",">15")))

# get model predictions
f_threshold = function(x, mortality_threshold) mean(x>mortality_threshold) 
# Full model
preds = posterior_epred(mod_creat_age_het, newdata = test_data, re_formula = NA)
test_data$pred_mod_creat_age_het = colMeans(preds)
test_data$creat_age_het_threshold = apply(preds, 2, f_threshold,mortality_threshold=mortality_threshold)

# Study het model
preds = posterior_epred(mod_creat_study_het, newdata = test_data, re_formula = NA)
test_data$pred_mod_creat_study_het = colMeans(preds)
test_data$creat_study_het_threshold = apply(preds, 2, f_threshold,mortality_threshold=mortality_threshold)
```



```{r}
p1=test_data %>%
  ggplot(aes(x = Creatinine_fold_change, y = pred_mod_creat_age_het, 
             colour =Age_category,linetype = Coma,
             group = interaction(Age_category,Coma)))+
  geom_line()+ ylab('Predicted mortality')+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks) +
  annotation_logticks(sides='b')+theme_bw(base_size = 13)+xlab('Creatinine: fold change from expected baseline')+
  geom_hline(yintercept = mortality_threshold)
# theme(legend.position = 'none')
p1_threshold=test_data %>%
  ggplot(aes(x = Creatinine_fold_change, y = creat_age_het_threshold, 
             colour = Age_category,linetype = Coma,
             group = interaction(Age_category,Coma)))+
  geom_line()+ylab('Probability mortality > 5%')+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks) +
  annotation_logticks(sides='b')+theme_bw(base_size = 13)+
  xlab('Creatinine: fold change from expected baseline')+
  geom_hline(yintercept = mortality_threshold)
# theme(legend.position = 'none')
grid.arrange(p1, p1_threshold)

sm_data_analysis %>% ggplot(aes(x=Creatinine_fold_change))+geom_histogram()+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks) 
sm_data_analysis %>% 
  ggplot(aes(x=Age, y = Creatinine_fold_change, colour = `Blood in urine`))+
  scale_y_log10(breaks = breaks, minor_breaks = minor_breaks) + geom_point()+
  scale_x_sqrt()+facet_wrap(~STUDY)

sm_data_analysis %>% filter(!is.na(Creatinine_umol_L)) %>%
  ggplot(aes(x=Age, y = Creatinine_umol_L, colour = Creatinine_fold_change>2.7))+
  geom_point()+scale_y_log10(breaks = breaks, minor_breaks = minor_breaks)+
  geom_hline(yintercept = 265)+scale_x_sqrt()+theme_minimal()

```



study het

```{r}
p1=test_data %>%
  ggplot(aes(x = Creatinine_fold_change, y = pred_mod_creat_study_het, 
             colour =STUDY,linetype = Coma,
             group = interaction(STUDY,Coma)))+
  geom_line()+ ylab('Predicted mortality')+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks) +
  annotation_logticks(sides='b')+theme_bw(base_size = 13)+xlab('Creatinine: fold change from expected baseline')+
  geom_hline(yintercept = mortality_threshold)+
  theme(legend.position = 'none')
p1_threshold=test_data %>%
  ggplot(aes(x = Creatinine_fold_change, y = creat_study_het_threshold, 
             colour = STUDY,linetype = Coma,
             group = interaction(STUDY,Coma)))+
  geom_line()+ylab('Probability mortality > 5%')+
  scale_x_log10(breaks = breaks, minor_breaks = minor_breaks) +
  annotation_logticks(sides='b')+theme_bw(base_size = 13)+
  xlab('Creatinine: fold change from expected baseline')+
  geom_hline(yintercept = mortality_threshold)+
  theme(legend.position = 'none')
grid.arrange(p1, p1_threshold)
```

